### 2014-08-25 paper.tex
Started an outline and intro paragraph.

\documentclass{article}

\begin{document}

\section{Introduction}

Emergig powered prothetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an abled bodied person. A modern gait lab is able to
collect a variety of kinmetic, kinetic, and physiological data from humans
during gait. This data can be used to drive the design of the controller. With
a rich enough data set, one may be able to identify control mechanism used
during a human's natural gait and recovery from perturbations. We have
collected one such data set that we believe is richer than previous gait data
sets and may be rich enough for control identification. The data can also be
used for verifciation purposes for controllers that have been designed in other
manners.

We collected over five and half hours of perturbed gait data from 11 subjects
sampled at 100 hz. The data has been organized and made available for other
research uses.

\end{document}

### 2014-08-29 paper.tex
Added abstract and start of raw data section.

@@
 \begin{document}
+
+\begin{abstract}
+
+  Herein we share a data set of typical gait data for 11 subjects walking at a
+  three speeds on an instrumented treadmill while being longitudinally
+  perturbed in the stance phase with psuedo-random fluctuations in the
+  treadmill belt speed. We provide raw marker and ground reaction loads in
+  addition to processed data that includes gait landmarks, 2D joint angles,
+  angular rates, and joint torques. The protocol is described in detail along
+  with the the additional meta data about each of the data files. This data can
+  be useful for validating or genreating mathematical models that capable of
+  simulating non-periodic and pertubed gaits.
+
+\end{abstract}
 
@@
 
+\section{Raw Data}
+
+The raw data consists of a set of ASCII tab delimited text files output from
+both the Mocap and Record modules in D-Flow in addition a YAML file that
+contains all of the necessary meta data for the given trial. These three files
+are stored in a hierarchy of directories with one trial per directory. The
+directories are named in the following fashion \verb+T001/+ where \verb+T+
+stands for ``trial'' and the following three digits are provide a unique trial
+identification number.
+
+\subsection{mocap-xxx.txt}
+
+The output from the D-Flow mocap module is stored in a tab delimted file named
+\verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id number. The file
+contains several columns:
+
+\begin{description}
+  \item[TimeStamp] The system monotonically time when D-Flow recieves a frame
+    from Cortex. These are approximately at 100 hz and given in seconds.
+  \item[FrameNumber] Monotonically increasing integers that correspond to each
+    frame received from Cortex.
+  \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
+    or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
+    reference frame. These values are in meters.
+  \item[Ground Reaction Loads] The left \verb+FP1+ and right \verb+FP2+ force
+    plates output the forces and moments of the BLANK
+
+Markers
+
+We make use of the full body 47 marker set described in \cite{Ton's HBM paper}.
+
 \end{document}

### 2014-09-05 paper.tex
Beefed up the raw data section, added a make file, and add more to outline.

@@
+% Editing notes
+% - No tabs, just spaces!
+% - No hanging whitespace
+% - Wrap lines to 79 characters
+% - Indentation should be two spaces per indent
+% - Don't worry about formatting details at first, just get the content in
+%   place.
+
 \documentclass{article}
+
+\usepackage{todo}
 
@@
 \section{Introduction}
-
+%
 Emergig powered prothetics are in need of bio-inspired control systems.
@@
 
+\section{Equipment}
+%
+The Human Motion and Control Lab at Cleveland State University has such a
+system. Our system includes:
+
+\begin{itemize}
+  \item A ForceLink R-Mill which has dual 6 DoF force plates, independent belts
+    for each foot, and lateral and pitch motion capabilities.
+  \item A 10 Camera Motion Analysis motion capture system which includes the
+    Cortex software and hardware for collecting analog and camera data
+    simultaneously.  Delsys wireless EMG + 3D Accelerometers.
+  \item Motek Medical’s D-Flow software and visual display system.
+\end{itemize}
+
+Cortex alone is capable of delivering data from the cameras, force plates, and
+analog sensors (EMG/Accelerometer), but D-Flow is required to collect data from
+digital sensors and the treadmill’s motion (lateral, pitch, and belts). D-Flow
+can output multiple types of files which contain the different data.
+
+Our motion capture system’s coordinate system is such that the X coordinate
+points to the right, the Y coordinate points upwards, and the Z coordinate
+follows from the right-hand-rule, i.e. points backwards with respect to a
+person walking forward on the treadmill. The camera’s coordinate system is
+aligned to an origin point on treadmill’s surface during camera calibration.
+
 \section{Raw Data}
-
+%
 The raw data consists of a set of ASCII tab delimited text files output from
-both the Mocap and Record modules in D-Flow in addition a YAML file that
+both the Mocap and Record modules in D-Flow in addition to a YAML file that
 contains all of the necessary meta data for the given trial. These three files
@@
 identification number.
+
+The text file output from the mocap module in DFlow is a tab delimited file.
+The first line is the header and contains a time stamp column, frame number
+column, marker position columns, force plate force/moment columns, force plate
+center of pressure columns, other analog columns, and potentially results from
+the real time Human Body Model \cite{Bogert2013} which is included with the
+D-Flow software. These are detailed below. The numerical values of the
+measurements are provided in decimal floating point notation with 6 decimals of
+precision, e.g. 123456.123456 [\%1.6f].
 
@@
 \begin{description}
-  \item[TimeStamp] The system monotonically time when D-Flow recieves a frame
-    from Cortex. These are approximately at 100 hz and given in seconds.
-  \item[FrameNumber] Monotonically increasing integers that correspond to each
-    frame received from Cortex.
+  \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
+    recieves a frame from Cortex. These are approximately at 100 hz and given
+    in seconds.
+  \item[FrameNumber] Monotonically increasing positive integers that correspond
+    to each frame received from Cortex.
   \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
@@
     reference frame. These values are in meters.
-  \item[Ground Reaction Loads] The left \verb+FP1+ and right \verb+FP2+ force
-    plates output the forces and moments of the BLANK
+  \item[Ground Reaction Loads] There are three ground reaction forces and three
+    ground reaction moments recorded by each of the two force plates in Newtons
+    and Newton-Meters, respectively. The prefix for these columns is either
+    \verb+FP1+ or \verb+FP2+ and represents either force plate 1 (left) or 2
+    (right). The suffixes are either \verb+.For[XYZ]+, \verb+.Mom[XYZ]+ for the
+    forces and moments, respectively. The force plate voltages are sampled at a
+    much higher frequency than the cameras, but delivered at the Cortex camera
+    sample rate, 100 Hz through the D-Flow mocap module. A force/moment
+    calibration matrix stored in Cortex converts the voltages to forces and
+    moments before sending it to D-Flow [1]. Cortex also computes the center of
+    pressure from the forces, moments, and force plate dimensions. These have
+    the same prefixes for the plate number, have the suffix \verb+.Cop[XYZ]+,
+    and are in meters.
+  \item[Analog Channels] Several analog signals are recorded under column
+    headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
+    sampled by Cortex and correspond to the ?92? National Instruments XXXX DAQ
+    box. The first twelve are the voltages from the force plate load cells. We
+    also record the acceleration of 4 points on the treadmill base in chanels
+    X-X.
+\end{description}
 
-Markers
+\subsection{record-xxx.txt}
+%
+The record module also outputs a tab delimited ASCII text file. It includes a
+\verb+Time+ column which records the D-Flow system time in seconds which
+corresponds to the time recorded in the \verb+TimeStamp+ column in mocap module
+tsv file. There are two additional columns \verb+RightBeltSpeed+ and
+\verb+LeftBeltSpeed+ which provide the independent belt speeds measure in
+meters per second by an ?encoder?.
 
-We make use of the full body 47 marker set described in \cite{Ton's HBM paper}.
+Finally, the record module is capable of recording the time at which various
+events occur, as detected or set by D-Flow. It does this by inserting commented
+(\#) lines in between the rows when the event occurred. The record files have
+several events that correspond to the phases of the protocol:
+
+\begin{description}
+  \item[A: Force Plate Zeroing] Marks the time at which there is no load on the
+    force plates and when the voltages were zeroed.
+  \item[B: Calibration Pose] Marks the time at which the person is in the
+    T-Pose.
+  \item[C: First Normal Walking] Marks the time when the treadmill begins Phase
+    1: constant belt speed.
+  \item[D: Longitudinal Perturbation] Marks the time when the treadmill begins
+    Phase 2: longitudinal perturbations in the belt speed.
+  \item[E: Second Normal Walking] Marks the time when phase 3 starts: constant
+    belt speed.
+  \item[F: Unloaded End] Marks the time at which there is no load on the force
+    plates and the belts are stationary.
+\end{description}
+
+\subsection{meta-xxx.yml}
+
+Each trial directory contains a meta data file in the YML format. There are
+three main headings: \verb+study+, \verb+subject+, and \verb+trial+.
+
+The \verb+study+ section contains identififying informatio for the overall
+study, an identification number, name, and description. This is the same for
+all meta data files in the study.
+
+The \verb+subject+ provides key value pairs of information about the subject in
+that trial. Each subject has a unique identification number and basic
+anthropomorphic data is included.
+
+The \verb+trial+ section contains the unique information about the trial. Each
+trial has a unique identification number.
+
+\todo[inline]{Add and example meta data file here with explanations of each
+field.}
+
+\subsection{Markers}
+%
+We make use of the full body 47 marker set described in \cite{Ton's HBM paper}
+and presented in detail in Table \ref{tab:marker-labels}. As with all camera
+based motion capture systems, the markers sometimes go missing in the
+recording.
+
+If the data was recorded in a D-Flow version less than 3.16.2rc4 [3], D-Flow
+records the last non-missing value in all three axes until the marker is
+visible again when a marker goes missing.
+
+In D-Flow versions greater than or equal to 3.16.2rc4 the missing markers are
+indicated in the TSV file as either 0.000000 or -0.000000, which is the same as
+has been in the HBM columns in all versions of D-Flow. The D-Flow version must
+be provided in the meta data yml file, otherwise it will assume D-Flow is at
+the latest version.
+
+\begin{table}
+  \begin{tabular}{rlll}
+    1 & LHEAD & Left head & Just above the ear, in the middle. \\
+    2 & THEAD & Top head & On top of the head, in line with the LHEAD and RHEAD. \\
+    3 & RHEAD & Right head & Just above the ear, in the middle. \\
+    4 & FHEAD & Forehead & Between line LHEAD/RHEAD and THEAD a bit right from center. \\
+    5 & C7 & C7 & On the 7th cervical vertebrae. \\
+    6 & T10 & T10 & On the 10th thoracic vertbrae. \\
+    7 & SACR & Sacrum bone & on the scacral bone. \\
+    8 & NAVE & Navel & On the navel. \\
+    9 & XYPH & Xiphoid process & Xiphoid process of the sternum. \\
+    10 & STRN & Sternum & On the jugular notch of the sternum. \\
+    11 & BBAC & Scapula & On the inferior angle fo the right scapular. \\
+    12 & LSHO & Left shoulder & Left acromion. \\
+    13 & LDELT & Left deltoid muscle & Apex of the deltoid muscle. \\
+    14 & LLEE & Left lateral elbow & Left lateral epicondyle of the elbow. Upper one in the T-Pose. \\
+    15 & LMEE & Left medial elbow & Left medial epicondyle of the elbow. Lower on in the T-Pose. \\
+    16 & LFRM & Left forearm & On 2/3 on the line between the LLEE and LMW. \\
+    17 & LMW & Left medial wrist & On styloid process radius, thumb side. \\
+    18 & LLW & Left lateral wrist & On styloid process ulna, pinky side. \\
+    19 & LFIN & Left fingers & Center of the hand. Caput metatarsal 3. \\
+    20 & RSHO & Right shoulder & Right acromion. \\
+    21 & RDELT & Right deltoid muscle & Apex of deltoid muscle. \\
+    22 & RLEE & Right lateral elbow & Right lateral epicondyle of the elbow. Lower one in the T-pose. \\
+    23 & RMEE & Right medial elbow & Right medial epicondyle of the elbow. Lower one in the T-pose. \\
+    24 & RFRM & Right forearm & On 1/3 on the line between the RLEE and RMW. \\
+    25 & RMW & Right medial wrist & On styloid process radius, thumb side. \\
+    26 & RLW & Right lateral wrist & On styloid process ulna, pinky side. \\
+    27 & RFIN & Right fingers & Center of the hand. Caput metatarsal 3. \\
+    28 & LASIS & Pelvic bone left front & Left anterio superior iliac spine. \\
+    29 & RASIS & Pelvic bone right front & Right anterior superior iliac spine. \\
+    30 & LPSIS & Pelvic bone left back & Left posterior superio iliac spine. \\
+    31 & RPSIS & Pelvic bone right back & Right posterior superior iliac spine. \\
+    32 & LGTRO & Left greater trochanter of the femur & On the cetner of the left greater trochanter. \\
+    33 & FLTHI & Left thigh & On 1/3 on the line between the LFTRO and LLEK. \\
+    34 & LLEK & Left lateral epicondyle of the knee & On the lateral side of the joint axis. \\
+    35 & LATI & Left anterior of the tibia & On 2/3 on the line between the LLEK and LLM. \\
+    36 & LLM & Left lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
+    37 & LHEE & Left heel & Center of the heel at the same height as the toe. \\
+    38 & LTOE & Left toe & Tip of big toe. \\
+    39 & LMT5 & Left 5th metatarsal & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
+    40 & RGTRO & Right greater trochanter of the femur & On the cetner of the right greater trochanter. \\
+    41 & FRTHI & Right thigh & On 2/3 on the line between the RFTRO and RLEK. \\
+    42 & RLEK & Right lateral epicondyle of the knee & On the lateral side of the joint axis. \\
+    43 & RATI & Right anterior of the tibia & On 1/3 on the line between the RLEK and RLM. \\
+    44 & RLM & Right lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
+    45 & RHEE & Right heel & Center of the heel at the same height as the toe. \\
+    46 & RTOE & Right toe & Tip of big toe. \\
+    47 & RMT5 & Right 5th metatarsal & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
+  \end{tabular}
+  \label{tab:marker-labels}
+\end{table}
 

### 2014-09-15 paper.tex
Added more to the meta data section and formatted the marker table in latex.

@@
 
-Each trial directory contains a meta data file in the YML format. There are
+Each trial directory contains a meta data file in the YAML format. There are
 three main headings: \verb+study+, \verb+subject+, and \verb+trial+.
 
-The \verb+study+ section contains identififying informatio for the overall
+The \verb+study+ section contains identififying information for the overall
 study, an identification number, name, and description. This is the same for
@@
 
-\todo[inline]{Add and example meta data file here with explanations of each
-field.}
+An example meta data file is shown in Figure \ref{fig:example-meta-data} and
+the following gives and explanation of the contents.
+
+\begin{figure}
+  \small
+  \begin{verbatim}
+    study:
+            id: 1
+            name: Gait Control Identification
+            description: Perturb the subject during walking and running.
+    subject:
+            id: 0
+    trial:
+            id: 58
+            subject-id: 0
+            datetime: 2014-03-28
+            notes: >
+                    This is an unloaded trial (no subject) 10 minute protocol that
+                    perturbs the subject longitudinally. The treadmill reference
+                    markers were not in place. We have the four custom made metal
+                    feet placed under the four corners of the treadmill to prevent
+                    pitch and sway movements. In addition, we also have the front
+                    left and rear right feet that were installed by Motek. We also
+                    have +/-3g analog accelerometers (ADXL330) affixed to the four
+                    corners of the main frame of the treadmill and wired into the
+                    NI box. The right handrail is removed.
+            nominal-speed: 0.8
+            nominal-speed-units: meters per second
+            stationary-platform: True
+            pitch: False
+            sway: False
+            hardware-settings:
+                    high-performance: True
+            analog-channel-map:
+                    Channel1.Anlg: F1Y1
+                    Channel2.Anlg: F1Y2
+                    Channel3.Anlg: F1Y3
+                    Channel4.Anlg: F1X1
+                    Channel5.Anlg: F1X2
+                    Channel6.Anlg: F1Z1
+                    Channel7.Anlg: F2Y1
+                    Channel8.Anlg: F2Y2
+                    Channel9.Anlg: F2Y3
+                    Channel10.Anlg: F2X1
+                    Channel11.Anlg: F2X2
+                    Channel12.Anlg: F2Z1
+                    Channel61.Anlg: Front_Left_AccX
+                    Channel62.Anlg: Front_Left_AccY
+                    Channel63.Anlg: Front_Left_AccZ
+                    Channel64.Anlg: Back_Left_AccX
+                    Channel65.Anlg: Back_Left_AccY
+                    Channel66.Anlg: Back_Left_AccZ
+                    Channel67.Anlg: Front_Right_AccX
+                    Channel68.Anlg: Front_Right_AccY
+                    Channel69.Anlg: Front_Right_AccZ
+                    Channel70.Anlg: Back_Right_AccX
+                    Channel71.Anlg: Back_Right_AccY
+                    Channel72.Anlg: Back_Right_AccZ
+            events:
+                    A: Force Plate Zeroing
+                    B: Calibration Pose
+                    C: First Normal Walking
+                    D: Longitudinal Perturbation
+                    E: Second Normal Walking
+                    F: Unloaded End
+            files:
+                    mocap: mocap-058.txt
+                    record: record-058.txt
+                    meta: meta-058.yml
+  \end{verbatim}
+  \caption{An example meta data YAML file.}
+  \label{fig:example-meta-data}
+\end{figure}
+
+\begin{description}
+  \item[analog-channel-map] A key value mapping of strings specifying the name
+    of the signal emmitted from the analog channels of the NI USB XXX.
+  \item[cortex-version] (string) Matches the version of Cortex used to record
+    the trial.
+  \item[datetime] (string) A date formatted string giving the date and/or time
+    of the trial in the \verb|YYYY-MM-DD| format.
+  \item[dflow-version] (string) Matches the version of D-Flow used to record
+    the trial.
+  \item[events] A key value map which prescribes names to the events recorded
+    in the record file.
+  \item[files] (list) A key value mapping of files associated with this trial
+    where the key is the file type and the values are the path to the file
+    relative to this meta file.
+  \item[hardware-settings] There are tons of settings for the hardware in both
+    D-Flow, Cortex, and the other software in the system. This contains any
+    non-default settings.
+    \begin{description}
+      \item[high-performance] (boolean) Indicates whether the D-Flow high
+        performance setting was on (True) or off (False).
+    \end{description}
+  \item[id] (integer) A unique three digit identifier for the trial. All of the
+    file names and directories associated with this trial include this.
+  \item[marker-map] A string key value map which contains names for the
+    markers.
+  \item[marker-set] (string) Indicates the HBM \cite{Bogert2013} marker
+    set used during the trial, either full, lower, or NA.
+  \item[nominal-speed] (float) The nominal or base speed during the trial.
+  \item[nominal-speed-units] (string) The units of the nominal speed.
+  \item[notes] {string} Notes about the trial.
+  \item[pitch] (boolean) Indicates if the pitch degree of freedom was acutated
+    during the trial.
+  \item[stationary-platform] (boolean) Indicates whether the treadmill
+      sway or pitch motion was actuated during the trial. If this flag is
+      false, the measured ground reaction forces must be compensated for the
+      inertial affects and be expressed the forces and moments in the motion
+      capture reference frame.
+  \item[subject-id] (integer) A number corresponding to the subject in the
+    trial.
+  \item[sway] (boolean) Indicates if the lateral (sway) degree of freedom was
+    acutated during the trial.
+\end{description}
 
@@
 \begin{table}
+  \centering
+  \caption{Descriptions of the 47 markers used in this study.}
   \begin{tabular}{rlll}
@@
     27 & RFIN  & Right fingers                         & Center of the hand. Caput metatarsal 3. \\
-    28 & LASIS & Pelvic bone left front & Left anterio superior iliac spine. \\
+    28 & LASIS & Pelvic bone left front                & Left anterior superior iliac spine. \\
     29 & RASIS & Pelvic bone right front               & Right anterior superior iliac spine. \\

### 2014-09-22 paper.tex
Start of the processed data section.

@@
 
+\section{Processed Data}
+
+We developed a command line data processing tool for common gait computations
+and provide the processed data from this. The tool is centered around two main
+classes, one to store a complete trial in memory and a second to compute common
+gait variables of interest.
+
+The \verb|motek.DFlowData| class stores all the data presented in the previous
+section and encodes several ``cleaning'' operations to put the data into a
+usable form.
+
+\begin{itemize}
+  \item Loads the meta data file into a Python dictionary.
+  \item Loads the DFlow mocap module TSV file into Pandas \verb|DataFrame|.
+  \item Relabels the column headers to more meaningful names if this is
+    specified in the meta data.
+  \item Identifies the missing values in the mocap marker data and replaces
+    them with \verb|numpy.nan|.
+  \item Interpolates the missing marker values and replaces them with
+    interpolated estimates.
+  \item Loads the DFlow record moudle TSV file into a Pandas \verb|DataFrame|.
+  \item Extracts the events and creates a dictionary mapping the event names in
+    the meta data to the events detected in the record module file.
+  \item Merges the data from the mocap module and record module into one data
+    frame at the maximum common constant sample rate.
+  \item Extracts sections of the data bounded by the events.
+  \item Saves the cleaned data to disk.
+\end{itemize}
+
+The \verb|gait.WalkingData| class is then used to compute things such as gait
+landmarks (toe off and heel strike times), basic 2D inverse dynamics, and to
+store the data into a Pandas \verb|Panel| with each gait cycle on the main axis
+at a specified sample rate. This data is also saved to disk in HDF5 format.
+
 \end{document}

### 2014-10-08 paper.tex

@@
   three speeds on an instrumented treadmill while being longitudinally
-  perturbed in the stance phase with psuedo-random fluctuations in the
+  perturbed in the stance phase with pseudorandom fluctuations in the
   treadmill belt speed. We provide raw marker and ground reaction loads in
@@
 %
-Emergig powered prothetics are in need of bio-inspired control systems.
+Emerging powered prothetics are in need of bio-inspired control systems.
 Ideally, a powered prosthetic would behave in such a way that the user would
@@
 developing bio-inspired control systems some of which aim to mimic the
-reactions and motion of an abled bodied person. A modern gait lab is able to
+reactions and motion of an able-bodied person. A modern gait lab is able to
 collect a variety of kinmetic, kinetic, and physiological data from humans
@@
 
-The output from the D-Flow mocap module is stored in a tab delimted file named
+The output from the D-Flow mocap module is stored in a tab delimited file named
 \verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id number. The file
@@
     headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
-    sampled by Cortex and correspond to the ?92? National Instruments XXXX DAQ
-    box. The first twelve are the voltages from the force plate load cells. We
-    also record the acceleration of 4 points on the treadmill base in chanels
-    X-X.
+    sampled by Cortex and correspond to the 96 analog-channel National 
+    Instruments 6255 DAQ box. The first twelve are the voltages from the 
+    force plate load cells. We also record the acceleration of 4 points on the
+    treadmill base in channels 61-72.
 \end{description}
@@
 
-The \verb+study+ section contains identififying information for the overall
+The \verb+study+ section contains identifying information for the overall
 study, an identification number, name, and description. This is the same for
@@
 
+\section{Perturbation Signals}
+
+  The data set uses a testing protocol containing a segment of normal 
+  walking, followed by a series of pseudorandom longitudinal 
+  perturbations, and ending with a second segment of normal walking.   
+  Three pseudorandom speed control signals, with mean velocities of 
+  0.8, 1.2, and 1.6 m/s, were generated through MATLAB and Simulink. 
+  Initially, a random acceleration signal was created using discrete-time 
+  Gaussian white noise and saturated at the maximum belt acceleration 
+  of 15 m/s2, integrated to obtain belt speed, and high-pass filtered 
+  (second-order Butterworth) to eliminate velocity drift.   Mean velocity was 
+  then added to the signal and limited between 0 and 3.6 m/s.  The cutoff
+  frequencies of the high-pass filter, as well as the variance in the 
+  acceleration signal, were adjusted until the appropriate standard deviations
+  for each mean velocity were obtained.
+
+  These standard deviations were determined through a trial and error method.
+  Initially, perturbations of $\pm$ 10% for each mean velocity were generated. 
+  However, this large deviation in the slowest mean velocity (0.8 m/s) would
+  unintentionally reduce the treadmill speed to values close to zero.  This 
+  would produce an undesired stopping and starting sensation.  Alternatively, 
+  smaller standard perturbations of $\pm$ 5% did not produce noticeable 
+  perturbations in the fastest trial (1.6 m/s).  Therefore, standard deviations
+  of 7.16%, 10.23%, and 13.1% were selected for the 0.8, 1.2, and 
+  1.6 m/s speeds, respectively.   This ensured that the test subject was
+  sufficiently perturbed for each speed, while remaining within the limits of 
+  our testing protocol.
+
+  To ensure that the treadmill could follow the higher frequencies of the 
+  random signals, the maximum treadmill acceleration of 15 m/s2 was used in the
+  D-Flow Software.  As this value remains constant throughout the experiment,
+  the initial starting and stopping of the treadmill would be too rapid for test
+  subjects.  To eliminate this, a ramped acceleration and deceleration was 
+  generated and placed at the beginning and end of the entire signal,
+  respectively.   Other additions include the one minute segments of 
+  unperturbed walking at the beginning and end of the trial and 30 seconds of 
+  no movement for force plate zeroing and subject T-pose calibration.  The 
+  sampling rate of the simulation was set to 100 Hz, corresponding to the 
+  D-Flow software's capabilities.
+
+  The resulting output from the MATLAB script and Simulink produces a comma-
+  delimited text file of six columns: timestamp, slow, normal, and fast walking
+  perturbation signals, and slow and fast running signals.  These running 
+  signals were generated in the event the testing protocol expanded from a
+  walking study to a running study, though they were not used in the data set.
+
 \section{Processed Data}

### 2014-10-16 paper.tex
Cleaned up some compile errors/warnings.

@@
 % - Indentation should be two spaces per indent
-% - Don't worry about formatting details at first, just get the content in
-%   place.
+% - Only one space after periods.
+% - Don't worry about formatting details, just get the content in place.
 
@@
 
+\usepackage{booktabs} % for table /toprule, /midrule, etc
 \usepackage{todo}
+
+\title{Perturbed Gait Data}
 
@@
 
-  Herein we share a data set of typical gait data for 11 subjects walking at a
-  three speeds on an instrumented treadmill while being longitudinally
-  perturbed in the stance phase with pseudorandom fluctuations in the
-  treadmill belt speed. We provide raw marker and ground reaction loads in
-  addition to processed data that includes gait landmarks, 2D joint angles,
-  angular rates, and joint torques. The protocol is described in detail along
-  with the the additional meta data about each of the data files. This data can
-  be useful for validating or genreating mathematical models that capable of
-  simulating non-periodic and pertubed gaits.
+  Herein we share a typical gait data set collected from 11 subjects walking at
+  a three speeds on an instrumented treadmill while being longitudinally
+  perturbed during each stance phase with pseudo-random fluctuations in the
+  treadmill belt. We provide raw marker and ground reaction loads in addition
+  to a presentation of processed data that includes gait landmarks, 2D joint
+  angles, angular rates, and joint torques and the software that does the
+  processing. The protocol is described in detail along with the additional
+  meta data about each of the data files. This data can be useful for
+  validating or generating mathematical models that capable of simulating
+  non-periodic and perturbed gaits.
 
@@
 %
-Emerging powered prothetics are in need of bio-inspired control systems.
+Emerging powered prosthetics are in need of bio-inspired control systems.
 Ideally, a powered prosthetic would behave in such a way that the user would
@@
 reactions and motion of an able-bodied person. A modern gait lab is able to
-collect a variety of kinmetic, kinetic, and physiological data from humans
-during gait. This data can be used to drive the design of the controller. With
-a rich enough data set, one may be able to identify control mechanism used
-during a human's natural gait and recovery from perturbations. We have
-collected one such data set that we believe is richer than previous gait data
-sets and may be rich enough for control identification. The data can also be
-used for verifciation purposes for controllers that have been designed in other
-manners.
+collect a variety of kinematic, kinetic, and physiological data from humans
+during gait. This data can potentially be used to drive the design of the human
+mimicking controller. With a rich enough data set, one may be able to identify
+control mechanism used during a human's natural gait and recovery from
+perturbations. We have collected data that we believe is richer than previous
+gait data sets and may be rich enough for control identification. The data can
+also be used for verification purposes for controllers that have been designed
+in other manners.
 
@@
 %
-The Human Motion and Control Lab at Cleveland State University has such a
-system. Our system includes:
+The Human Motion and Control Lab at Cleveland State University has modern human
+motion measurement equipment. Our system includes:
 
 \begin{itemize}
-  \item A ForceLink R-Mill which has dual 6 DoF force plates, independent belts
-    for each foot, and lateral and pitch motion capabilities.
-  \item A 10 Camera Motion Analysis motion capture system which includes the
-    Cortex software and hardware for collecting analog and camera data
-    simultaneously.  Delsys wireless EMG + 3D Accelerometers.
+  \item A ForceLink R-Mill which has dual 6 degree of freedom force plates,
+    independent belts for each foot, and lateral and pitch motion capabilities.
+  \item A 10 Camera Motion Analysis motion capture system which includes Osprey
+    cameras, the Cortex software, and hardware for collecting analog and camera
+    data simultaneously. Delsys wireless EMG + 3D Accelerometers.
   \item Motek Medical’s D-Flow software and visual display system.
@@
 
-Cortex alone is capable of delivering data from the cameras, force plates, and
-analog sensors (EMG/Accelerometer), but D-Flow is required to collect data from
-digital sensors and the treadmill’s motion (lateral, pitch, and belts). D-Flow
-can output multiple types of files which contain the different data.
+Cortex alone is capable of delivering data from the cameras along with force
+plates, and analog sensors (EMG/Accelerometer) through a National Instruments
+XXXX data acquisition unit. D-Flow is required to collect data from any digital
+sensors and the treadmill's motion (lateral, pitch, and belts). D-Flow can
+output multiple types of files which contain the different data.
 
-Our motion capture system’s coordinate system is such that the X coordinate
+Our motion capture system's coordinate system is such that the X coordinate
 points to the right, the Y coordinate points upwards, and the Z coordinate
 follows from the right-hand-rule, i.e. points backwards with respect to a
-person walking forward on the treadmill. The camera’s coordinate system is
-aligned to an origin point on treadmill’s surface during camera calibration.
+person walking forward on the treadmill. The camera's coordinate system is
+aligned to an origin point on treadmill's surface during camera calibration.
 
@@
 The raw data consists of a set of ASCII tab delimited text files output from
-both the Mocap and Record modules in D-Flow in addition to a YAML file that
-contains all of the necessary meta data for the given trial. These three files
-are stored in a hierarchy of directories with one trial per directory. The
-directories are named in the following fashion \verb+T001/+ where \verb+T+
+both the ``mocap'' and ``record'' modules in D-Flow in addition to a YAML file
+that contains all of the necessary meta data for the given trial. These three
+files are stored in a hierarchy of directories with one trial per directory.
+The directories are named in the following fashion \verb+T001/+ where \verb+T+
 stands for ``trial'' and the following three digits are provide a unique trial
@@
 \begin{table}
+  \caption{Descriptions of the 47 markers used in this study.}
   \centering
-  \caption{Descriptions of the 47 markers used in this study.}
+  \small
   \begin{tabular}{rlll}
+    \toprule
+    Number & Label & Name & Description \\
+    \midrule
     1  & LHEAD & Left head                             & Just above the ear, in the middle. \\
@@
     47 & RMT5  & Right 5th metatarsal                  & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
+    \bottomrule
   \end{tabular}
@@
 \section{Perturbation Signals}
-
-  The data set uses a testing protocol containing a segment of normal 
-  walking, followed by a series of pseudorandom longitudinal 
-  perturbations, and ending with a second segment of normal walking.   
-  Three pseudorandom speed control signals, with mean velocities of 
-  0.8, 1.2, and 1.6 m/s, were generated through MATLAB and Simulink. 
-  Initially, a random acceleration signal was created using discrete-time 
-  Gaussian white noise and saturated at the maximum belt acceleration 
-  of 15 m/s2, integrated to obtain belt speed, and high-pass filtered 
-  (second-order Butterworth) to eliminate velocity drift.   Mean velocity was 
-  then added to the signal and limited between 0 and 3.6 m/s.  The cutoff
-  frequencies of the high-pass filter, as well as the variance in the 
-  acceleration signal, were adjusted until the appropriate standard deviations
-  for each mean velocity were obtained.
+%
+The data set uses a testing protocol containing a segment of normal walking,
+followed by a series of pseudo-random longitudinal perturbations, and ending
+with a second segment of normal walking. Three pseudo-random speed control
+signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
+generated through MATLAB and Simulink. Initially, a random acceleration signal
+was created using discrete-time Gaussian white noise and saturated at the
+maximum belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt
+speed, and high-pass filtered (second-order Butterworth) to eliminate velocity
+drift. Mean velocity was then added to the signal and limited between 0 and 3.6
+m/s. The cutoff frequencies of the high-pass filter, as well as the variance in
+the acceleration signal, were adjusted until the appropriate standard
+deviations for each mean velocity were obtained.
 
 These standard deviations were determined through a trial and error method.
-  Initially, perturbations of $\pm$ 10% for each mean velocity were generated. 
-  However, this large deviation in the slowest mean velocity (0.8 m/s) would
-  unintentionally reduce the treadmill speed to values close to zero.  This 
-  would produce an undesired stopping and starting sensation.  Alternatively, 
-  smaller standard perturbations of $\pm$ 5% did not produce noticeable 
-  perturbations in the fastest trial (1.6 m/s).  Therefore, standard deviations
-  of 7.16%, 10.23%, and 13.1% were selected for the 0.8, 1.2, and 
-  1.6 m/s speeds, respectively.   This ensured that the test subject was
-  sufficiently perturbed for each speed, while remaining within the limits of 
-  our testing protocol.
+Initially, perturbations of $\pm$ 10\% for each mean velocity were generated.
+However, this large deviation in the slowest mean velocity (0.8 $m \cdot
+s^{-1}$) would unintentionally reduce the treadmill speed to values close to
+zero. This would produce an undesired stopping and starting sensation.
+Alternatively, smaller standard perturbations of $\pm$ 5\% did not produce
+noticeable perturbations in the fastest trial (1.6 $m \cdot s^{-1}$).
+Therefore, standard deviations of 7.16\%, 10.23\%, and 13.1\% were selected for
+the 0.8, 1.2, and 1.6 $m \cdot s^{-1}$ speeds, respectively.  This ensured that
+the test subject was sufficiently perturbed for each speed, while remaining
+within the limits of our testing protocol.
 
-  To ensure that the treadmill could follow the higher frequencies of the 
-  random signals, the maximum treadmill acceleration of 15 m/s2 was used in the
-  D-Flow Software.  As this value remains constant throughout the experiment,
+To ensure that the treadmill could follow the higher frequencies of the random
+signals, the maximum treadmill acceleration of 15 $m \cdot s^{-2}$ was used in
+the D-Flow Software. As this value remains constant throughout the experiment,
 the initial starting and stopping of the treadmill would be too rapid for test
@@
 generated and placed at the beginning and end of the entire signal,
-  respectively.   Other additions include the one minute segments of 
-  unperturbed walking at the beginning and end of the trial and 30 seconds of 
-  no movement for force plate zeroing and subject T-pose calibration.  The 
-  sampling rate of the simulation was set to 100 Hz, corresponding to the 
-  D-Flow software's capabilities.
+respectively. Other additions include the one minute segments of unperturbed
+walking at the beginning and end of the trial and 30 seconds of no movement for
+force plate zeroing and subject T-pose calibration.  The sampling rate of the
+simulation was set to 100 Hz, corresponding to the D-Flow software's
+capabilities.
 
@@
 delimited text file of six columns: timestamp, slow, normal, and fast walking
-  perturbation signals, and slow and fast running signals.  These running 
-  signals were generated in the event the testing protocol expanded from a
-  walking study to a running study, though they were not used in the data set.
+perturbation signals, and slow and fast running signals.  These running signals
+were generated in the event the testing protocol expanded from a walking study
+to a running study, though they were not used in the data set.
 

### 2014-10-18 paper.tex
Updated DFlowData pipeline description.

@@
 
-The \verb|motek.DFlowData| class stores all the data presented in the previous
-section and encodes several ``cleaning'' operations to put the data into a
-usable form.
+The \verb|motek.DFlowData| class collects and stores all the raw data presented
+in the previous section and applies several ``cleaning'' operations to put the
+data into a usable form.
 
@@
     specified in the meta data.
-  \item Identifies the missing values in the mocap marker data and replaces
-    them with \verb|numpy.nan|.
-  \item Interpolates the missing marker values and replaces them with
-    interpolated estimates.
-  \item Loads the DFlow record moudle TSV file into a Pandas \verb|DataFrame|.
+  \item Optionally identifies the missing values in the mocap marker data and
+    replaces them with \verb|numpy.nan|.
+  \item Optionally interpolates the missing marker values and replaces them
+    with interpolated estimates using a variety of interpolation methods.
+  \item Loads the D-Flow record moudle TSV file into a Pandas \verb|DataFrame|.
   \item Extracts the events and creates a dictionary mapping the event names in
     the meta data to the events detected in the record module file.
+  \item Compensates the ground reaction loads for treadmill motion.
   \item Merges the data from the mocap module and record module into one data

### 2014-10-18 src/unperturbed_perturbed_comparison.py
Added src dir with README and source for comparison plot.

@@ -1,0 +1,180 @@
+#!/usr/bin/env python
+
+import os
+import random
+
+import yaml
+import matplotlib.pyplot as plt
+import pandas
+from gaitanalysis.motek import DFlowData, markers_for_2D_inverse_dynamics
+from gaitanalysis.gait import WalkingData
+from gaitanalysis.utils import _percent_formatter
+import seaborn as sbn
+
+
+def trial_file_paths(trials_dir, trial_number):
+    """Returns the most common paths to the trials in the gait
+    identification data set.
+
+    Parameters
+    ==========
+    trials_dir : string
+        The path to the main directory for the data. This directory should
+        contain subdirectories: `T001/`, `T002/`, etc.
+    trial_number : string
+        Three digit trial number, e.g. `005`.
+
+    """
+
+    trial_dir = 'T' + trial_number
+    mocap_file = 'mocap-' + trial_number + '.txt'
+    record_file = 'record-' + trial_number + '.txt'
+    meta_file = 'meta-' + trial_number + '.yml'
+
+    mocap_file_path = os.path.join(trials_dir, trial_dir, mocap_file)
+    record_file_path = os.path.join(trials_dir, trial_dir, record_file)
+    meta_file_path = os.path.join(trials_dir, trial_dir, meta_file)
+
+    return mocap_file_path, record_file_path, meta_file_path
+
+
+def load_data(event, tmp):
+
+    file_name = '_'.join([n.lower() for n in event.split(' ')]) + '.h5'
+
+    tmp_data_path = os.path.join(tmp, file_name)
+
+    try:
+        f = open(tmp_data_path, 'r')
+    except IOError:
+        print('Cleaning and processing {} data...'.format(event))
+        # Load raw data, clean it up, and extract the perturbation section.
+        dflow_data = DFlowData(*paths)
+        dflow_data.clean_data(ignore_hbm=True)
+        perturbed_df = \
+            dflow_data.extract_processed_data(event=event,
+                                              index_col='TimeStamp',
+                                              isb_coordinates=True)
+
+        # Compute the lower limb 2D inverse dynamics, identify right hell
+        # strike times, and split the data into gait cycles.
+        gait_data = WalkingData(perturbed_df)
+        marker_set = dflow_data.meta['trial']['marker-set']
+        subject_mass = dflow_data.meta['subject']['mass']
[…]

### 2014-10-18 paper.tex

@@
 
+\usepackage{graphicx}
 \usepackage{booktabs} % for table /toprule, /midrule, etc
@@
 
+Figure~\ref{fig:angle-torque-comparison} compares the mean and standard
+deviation of saggital plane joint angles and torques from the perturbed gait
+cycles and the unperturbed gait cycles computed from trial 20 for 120 gait
+cycles. The variation in the dynamics is larger for the perturbed gait cycles.
+
+\begin{figure}
+  \includegraphics{figures/unperturbed-perturbed-comparison.pdf}
+  \caption{Mean (solid, dashed) and $2\sigma$ (shaded) ($N=120$) joint angles
+    and torques from both unperturbed (purple) and perturbed (blue) gait cycles
+    from trial 20.}
+  \label{fig:angle-torque-comparison}
+\end{figure}
+
+Figure~\ref{fig:gait-cycle-stats-comparison} compares the distribution in
+several statistics of the unperturbed and perturbed gait cycles.
+
+\begin{figure}
+  \includegraphics{figures/unperturbed-perturbed-hist-comparison.pdf}
+  \caption{Histograms of the average belt speed, cadence, step duration, and
+    stride length which compare 120 gait cycles for the unperturbed (purple)
+    and perturbed (blue).}
+  \label{fig:gait-cycle-stats-comparison}
+\end{figure}
+
 \end{document}

### 2014-10-20 src/unperturbed_perturbed_comparison.py
Minor edits to docstrings.

@@ -77,7 +77,7 @@ def load_data(event, paths, tmp):
                                               index_col='TimeStamp',
                                               isb_coordinates=True)
 
-        # Compute the lower limb 2D inverse dynamics, identify right hell
+        # Compute the lower limb 2D inverse dynamics, identify right heel
         # strike times, and split the data into gait cycles.
         gait_data = WalkingData(perturbed_df)
         marker_set = dflow_data.meta['trial']['marker-set']
@@ -102,7 +102,8 @@ def load_data(event, paths, tmp):
 
 
 def remove_bad_steps(gait_data, lower, upper, col):
-    """
+    """Returns the gait cycles with outliers removed based on the step_data
+    DataFrame column.
 
     Parameters
     ==========

### 2014-10-20 paper.tex

@@
 
+\section{Conclusion}
+
+We have presented the details of what we believe to be a rich dataset of motion
+and ground reaction loads humans exhibit when recovering from minor
+longitudinal perturbations. The data provides a large sample of gait cycles
+from both unperturbed and perturbed walking at three different nominal speeds.
+The raw data is provided for reuse with sufficient detail with regards to meta
+data and technical details. In addition to the data, we provide software that
+can process the data for cleaning purposes and to produce typical saggital
+plane gait variables of interest. Among other uses, we believe the dataset is
+ideally suited for control identification purposes. Many researchers are
+working on mathematical models for control in gait and this dataset provides
+both a way to validate these models and data that can be used to generate
+models.
+
 \end{document}

### 2014-10-21 paper.tex
added figures if needed
Simulink diagram, comparison of input and output of perturbation signals

@@
 signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
-generated through MATLAB and Simulink. Initially, a random acceleration signal
-was created using discrete-time Gaussian white noise and saturated at the
-maximum belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt
-speed, and high-pass filtered (second-order Butterworth) to eliminate velocity
-drift. Mean velocity was then added to the signal and limited between 0 and 3.6
-m/s. The cutoff frequencies of the high-pass filter, as well as the variance in
-the acceleration signal, were adjusted until the appropriate standard
-deviations for each mean velocity were obtained.
+generated through MATLAB and Simulink (Figure~\ref{fig:simulink}). Initially, a 
+random acceleration signal was created using discrete-time Gaussian white 
+noise and saturated at the maximum belt acceleration of 15 $m \cdot s^{-2}$,
+integrated to obtain belt speed, and high-pass filtered (second-order
+Butterworth) to eliminate velocity drift. Mean velocity was then added to the 
+signal and limited between 0 and 3.6 m/s. The cutoff frequencies of the high-
+pass filter, as well as the variance in the acceleration signal, were adjusted 
+until the appropriate standard deviations for each mean velocity were obtained.
+
+\begin{figure}
+  \includegraphics{figures/longitudinal_perturbation.PNG}
+  \caption{MATLAB Simulink model for generating longitudinal perturbations,
+                given variance, cutoff frequency of the high-pass filter, and
+                the average speed}
+  \label{fig:simulink}
+\end{figure}
 
@@
 were generated in the event the testing protocol expanded from a walking study
-to a running study, though they were not used in the data set.
+to a running study, though they were not used in the data set.  The actual motion
+of the treadmill belts are compared to the original input signals, shown in 
+Figure~\ref{fig:input_output}.
+
+\begin{figure}
+  \includegraphics{figures/input_vs_output.eps}
+  \caption{Generated input signal (black) and actual treadmill belt movement
+                (red) for average speeds of 0.8, 1.2, and 1.6 m/s, respectively}
+  \label{fig:input_output}
+\end{figure}
 

### 2014-10-21 src/vgait-input-signals/signal_check.m
added source code and pdf images

@@ -1,0 +1,73 @@
+%signal_check.m plots the input vs. output for three average perturbation
+%speeds (0.8, 1.2, 1.6 m/s) for a portion (around 300-310 seconds) of the 
+%entire trial
+
+clear
+clc
+
+%-------------------------------------------------------------------------
+%Loading Files
+%-------------------------------------------------------------------------
+    input_signals=load('Longitudinal_Perturbation_2.4.2014.txt');
+    output_slow=load('record-079_0001.txt');
+    output_normal=load('record-080_0001.txt');
+    output_fast=load('record-081_0001.txt');
+%-------------------------------------------------------------------------
+%Declaring Variables
+%-------------------------------------------------------------------------
+    %Time Variables (necessary because Record Module has varying timestamp)
+        time_input=input_signals(:,1);
+        output_slow_time=output_slow(:,1); 
+        output_slow_time=output_slow_time-output_slow_time(1,1);
+        output_normal_time=output_normal(:,1);
+        output_normal_time=output_normal_time-output_normal_time(1,1);
+        output_fast_time=output_fast(:,1);
+        output_fast_time=output_fast_time-output_fast_time(1,1);
+    %Speed
+        inputs=[input_signals(:,2) input_signals(:,3) input_signals(:,4)];
+        output_slow=output_slow(:,2); 
+        output_normal=output_normal(:,2);
+        output_fast=output_fast(:,2);
+%-------------------------------------------------------------------------
+%Plotting
+%-------------------------------------------------------------------------
+    set(0,'defaulttextinterpreter','latex')
+    set(0,'DefaultAxesFontName', 'latex')
+    ti = get(gca,'TightInset');
+    set(gca,'Position',[ti(1) ti(2) 1-ti(3)-ti(1) 1-ti(4)-ti(2)]);
+    set(gca,'units','centimeters')
+    pos = get(gca,'Position');
+    ti = get(gca,'TightInset');
+    set(gcf, 'PaperUnits','centimeters');
+    set(gcf, 'PaperSize', [pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
+    set(gcf, 'PaperPositionMode', 'manual');
+    set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
+
+    figure(1)
+        subplot(3,1,1)
[…]

### 2014-11-14 src/vgait-input-signals/frequency_analysis.m
added frequency analysis results
graph, source code, and added to draft

@@ -1,0 +1,100 @@
+clear
+clc
+
+%-------------------------------------------------------------------------
+%Loading Files
+%-------------------------------------------------------------------------
+    input_signals=load('Longitudinal_Perturbation_2.4.2014.txt');
+    output_slow=load('record-079_0001.txt');
+    output_normal=load('record-080_0001.txt');
+    output_fast=load('record-081_0001.txt');
+%-------------------------------------------------------------------------
+%Declaring Variables
+%-------------------------------------------------------------------------
+    %Time Variables (necessary because Record Module has varying timestamp)
+        time_input=input_signals(:,1);
+        output_slow_time=output_slow(:,1); 
+        output_slow_time=output_slow_time-output_slow_time(1,1);
+        output_normal_time=output_normal(:,1);
+        output_normal_time=output_normal_time-output_normal_time(1,1);
+        output_fast_time=output_fast(:,1);
+        output_fast_time=output_fast_time-output_fast_time(1,1);
+    %Speed
+        inputs=[input_signals(:,2) input_signals(:,3) input_signals(:,4)];
+        output_slow=output_slow(:,2); 
+        output_normal=output_normal(:,2);
+        output_fast=output_fast(:,2);
+%-------------------------------------------------------------------------
+%Spectral Analysis
+%-------------------------------------------------------------------------
+    Fs=100;
+    T=1/Fs;
+    %Lengths
+        L_inputs = length(inputs);
+        L_output_slow=length(output_slow);
+        L_output_normal = length(output_normal);
+        L_output_fast = length(output_fast);
+    %NFFT
+        NFFT_inputs=2^nextpow2(L_inputs);
+        NFFT_output_slow = 2^nextpow2(L_output_slow);
+        NFFT_output_normal = 2^nextpow2(L_output_normal);
+        NFFT_output_fast = 2^nextpow2(L_output_fast);
+    %FI
+        fi_inputs = Fs/2*linspace(0,1,NFFT_inputs/2+1);
+        fi_output_slow=Fs/2*linspace(0,1,NFFT_output_slow/2+1);
[…]

### 2014-11-15 paper.tex
added source code and pdf images
simulink diagram removed

--- 2014-10-20 paper.tex
@@
 generated through MATLAB and Simulink. Initially, a  random acceleration signal
-was created using discrete-time Gaussian white noise and saturated at the
-maximum belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt
-speed, and high-pass filtered (second-order Butterworth) to eliminate velocity
-drift. Mean velocity was then added to the signal and limited between 0 and 3.6
-m/s. The cutoff frequencies of the high-pass filter, as well as the variance in
-the acceleration signal, were adjusted until the appropriate standard
-deviations for each mean velocity were obtained.
+was created using discrete-time Gaussian white noise and saturated at the maximum
+belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt speed, and 
+high-pass filtered (second-order Butterworth) to eliminate velocity drift. Mean 
+velocity was then added to the signal and limited between 0 and 3.6 m/s. The 
+cutoff frequencies of the high-pass filter, as well as the variance in the acceleration 
+signal, were adjusted until the appropriate standard deviations for each mean 
+velocity were obtained.
 
@@
 were generated in the event the testing protocol expanded from a walking study
-to a running study, though they were not used in the data set.
+to a running study, though they were not used in the data set.  The actual motion
+of the treadmill belts are compared to the original input signals is shown in 
+Figure~\ref{fig:input_output}.   The standard deviations of the outputs were 
+calculated at 6.07\%, 9.68\%, and 12.68\% for the the 0.8, 1.2, and 
+1.6 $m \cdot s^{-1}$ speeds, respectively.
+
+\begin{figure}
+  \includegraphics{figures/input_vs_output.pdf}
+  \caption{Generated input signal (black) and actual treadmill belt movement
+                (red) for average speeds of 0.8, 1.2, and 1.6 m/s, respectively}
+  \label{fig:input_output}
+\end{figure}
+
+For further analysis, the input and output signals were transformed into the 
+frequency domain through the Fast Fourier algorithm, and the results are
+shown in Figure~\ref{fig:freq_analysis}.  It was observed that for the 1.2 $m \cdot s^{-1}$
+average walking speed, the amplitude of the output is significantly lower 
+than the amplitude of the input signal at lower frequencies.  Additionally, the 
+amplitude of output signal in the 0.8 $m \cdot s^{-1}$ walking speed begins to attenuate 
+around 2 Hz, which is a noticeably lower frequency than the other walking 
+speeds.
+
+\begin{figure}
+  \includegraphics{figures/frequency_analysis.pdf}
+  \caption{Frequency analysis of the  input (black) and output (red)
+                for average speeds of 0.8, 1.2, and 1.6 m/s, respectively}
+  \label{fig:freq_analysis}
+\end{figure}
 

### 2014-11-24 paper.tex
Changed the histogram to a box plot.
I removed the redundancy of stride freq and stride duration.

@@
 
-Figure~\ref{fig:gait-cycle-stats-comparison} compares the distribution in
-several statistics of the unperturbed and perturbed gait cycles.
+Figure~\ref{fig:gait-cycle-stats-comparison} compares the distribution in a few
+statistics of the unperturbed and perturbed gait cycles.
 
 \begin{figure}
-  \includegraphics{figures/unperturbed-perturbed-hist-comparison.pdf}
-  \caption{Histograms of the average belt speed, cadence, step duration, and
-    stride length which compare 120 gait cycles for the unperturbed (purple)
-    and perturbed (blue).}
+  \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
+  \caption{Box plots of the average belt speed, stride frequency, and stride
+    length which compare 120 gait cycles for the unperturbed (purple) and
+    perturbed (blue).}
   \label{fig:gait-cycle-stats-comparison}

### 2014-11-24 paper.tex
Update paper.tex
Edits by Ton.

--- 2014-11-15 paper.tex
@@
   perturbed during each stance phase with pseudo-random fluctuations in the
-  treadmill belt. We provide raw marker and ground reaction loads in addition
+  speed of the treadmill belt. We provide raw marker and ground reaction loads in addition
   to a presentation of processed data that includes gait landmarks, 2D joint
@@
 aligned to an origin point on treadmill's surface during camera calibration.
+The same point is used as the origin of the ground reaction force measuring
+system.
 
@@
     force plate load cells. We also record the acceleration of 4 points on the
-    treadmill base in channels 61-72.
+    treadmill base in channels 61-72.  TvdB: are these useful?  If so, we need to 
+    provide information on how to process these channels.  If not, maybe
+    not mention them at all?
 \end{description}
@@
 tsv file. There are two additional columns \verb+RightBeltSpeed+ and
-\verb+LeftBeltSpeed+ which provide the independent belt speeds measure in
-meters per second by an ?encoder?.
+\verb+LeftBeltSpeed+ which provide the independent belt speeds measured in
+meters per second by an encoder in the treadmill.
 
@@
 We developed a command line data processing tool for common gait computations
-and provide the processed data from this. The tool is centered around two main
+and provide the processed data from this. The tool was developed in Pythin and centered around two main
 classes, one to store a complete trial in memory and a second to compute common

### 2014-11-26 src/vgait-input-signals/frequency_analysis.m
updated plot colors and axis labels
Using "dusty purple" and "windows blue" now.  Also the ylabel on the
frequency graphs has been changed to "Amplitude"

@@ -67,34 +67,32 @@ clc
 
     figure(1)
         subplot(3,1,1)
-        plot(fi_inputs,2*abs(Y_input(1:NFFT_inputs/2+1,1)),'k')
+        plot(fi_inputs,2*abs(Y_input(1:NFFT_inputs/2+1,1)),'Color',[0.5098 0.3725 0.5294])
         hold on
-        plot(fi_output_slow,2*abs(Y_output_slow(1:NFFT_output_slow/2+1)),'r')
+        plot(fi_output_slow,2*abs(Y_output_slow(1:NFFT_output_slow/2+1)),'Color',[0.2157 0.4706 0.7490])
         legend('Input','Output')
         title('Slow Walking (0.8 m/s)','Fontweight','bold')
-        ylabel('|Y(f)|','Fontweight','bold')
+        ylabel('Amplitude','Fontweight','bold')
         ylim([0 6*10^-3]); xlim([0 8])
     
         subplot(3,1,2)
-        plot(fi_inputs,2*abs(Y_input(1:NFFT_inputs/2+1,2)),'k')
+        plot(fi_inputs,2*abs(Y_input(1:NFFT_inputs/2+1,2)),'Color',[0.5098 0.3725 0.5294])
         hold on
-        plot(fi_output_normal,2*abs(Y_output_slow(1:NFFT_output_normal/2+1)),'r')
+        plot(fi_output_normal,2*abs(Y_output_slow(1:NFFT_output_normal/2+1)),'Color',[0.2157 0.4706 0.7490])
         legend('Input','Output')
         title('Normal Walking (1.2 m/s)','Fontweight','bold')
-        ylabel('|Y(f)|','Fontweight','bold')
+        ylabel('Amplitude','Fontweight','bold')
         ylim([0 0.015]); xlim([0 8])
         
         subplot(3,1,3)
-         plot(fi_inputs,2*abs(Y_input(1:NFFT_inputs/2+1,3)),'k')
+         plot(fi_inputs,2*abs(Y_input(1:NFFT_inputs/2+1,3)),'Color',[0.5098 0.3725 0.5294])
         hold on
-        plot(fi_output_fast,2*abs(Y_output_fast(1:NFFT_output_fast/2+1)),'r')
+        plot(fi_output_fast,2*abs(Y_output_fast(1:NFFT_output_fast/2+1)),'Color',[0.2157 0.4706 0.7490])
         legend('Input','Output')
         title('Fast Walking (1.6 m/s)','Fontweight','bold')
[…]

### 2014-11-26 src/vgait-input-signals/signal_check.m

@@ -1,7 +1,3 @@
-%signal_check.m plots the input vs. output for three average perturbation
-%speeds (0.8, 1.2, 1.6 m/s) for a portion (around 300-310 seconds) of the 
-%entire trial
-
 clear
 clc
 
@@ -45,26 +41,26 @@ clc
 
     figure(1)
         subplot(3,1,1)
-        plot(time_input, inputs(:,1),'k')
+        plot(time_input, inputs(:,1),'Color',[0.5098 0.3725 0.5294])
         hold on
-        plot(output_slow_time, output_slow,'r')
+        plot(output_slow_time, output_slow,'Color',[0.2157 0.4706 0.7490])
         legend('Input','Output')
         xlim([305 310]); 
         ylabel('Speed (m/s)','Fontweight','bold')
         title('Slow Walking (0.8 m/s)','Fontweight','bold')
     
         subplot(3,1,2)
-        plot(time_input, inputs(:,2),'k')
+        plot(time_input, inputs(:,2),'Color',[0.5098 0.3725 0.5294])
         hold on
-        plot(output_normal_time, output_normal,'r')
+        plot(output_normal_time, output_normal,'Color',[0.2157 0.4706 0.7490])
         ylabel('Speed (m/s)','Fontweight','bold')
         xlim([305 310]); 
         title('Normal Walking (1.2 m/s)','Fontweight','bold')
     
         subplot(3,1,3)
-        plot(time_input,inputs(:,3),'k')
+        plot(time_input,inputs(:,3),'Color',[0.5098 0.3725 0.5294])
         hold on
-        plot(output_fast_time,output_fast,'r')
+        plot(output_fast_time,output_fast,'Color',[0.2157 0.4706 0.7490])
         ylabel('Speed (m/s)','Fontweight','bold')
         xlim([305 310]); xlabel('Time (s)','Fontweight','bold')
         title('Fast Walking (1.6 m/s)','Fontweight','bold')

### 2014-11-26 paper.tex
Update paper.tex
Added acknowledgments.
update text to reflect new plot colors

@@
 to a running study, though they were not used in the data set.  The actual motion
-of the treadmill belts are compared to the original input signals is shown in 
+of the treadmill belts are compared to the original input signals as shown in 
 Figure~\ref{fig:input_output}.   The standard deviations of the outputs were 
@@
   \includegraphics{figures/input_vs_output.pdf}
-  \caption{Generated input signal (black) and actual treadmill belt movement
-                (red) for average speeds of 0.8, 1.2, and 1.6 m/s, respectively}
+  \caption{Treadmill belt velocity input signals (purple) and recorded output 
+                velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
+                respectively}
   \label{fig:input_output}
@@
 
-For further analysis, the input and output signals were transformed into the 
-frequency domain through the Fast Fourier algorithm, and the results are
-shown in Figure~\ref{fig:freq_analysis}.  It was observed that for the 1.2 $m \cdot s^{-1}$
-average walking speed, the amplitude of the output is significantly lower 
-than the amplitude of the input signal at lower frequencies.  Additionally, the 
-amplitude of output signal in the 0.8 $m \cdot s^{-1}$ walking speed begins to attenuate 
-around 2 Hz, which is a noticeably lower frequency than the other walking 
-speeds.
+For further analysis, the input and output signals for each average speed were
+transformed into the frequency domain through the Fast Fourier algorithm, and 
+the results are shown in Figure~\ref{fig:freq_analysis}.  It was observed that for
+the 1.2 $m \cdot s^{-1}$ average walking speed, the amplitude of the output is 
+significantly lower than the amplitude of the input signal at lower frequencies.  
+Additionally, the amplitude of output signal in the 0.8 $m \cdot s^{-1}$ walking
+speed begins to attenuate around 2 Hz, which is a noticeably lower frequency than
+the other walking speeds.
 
@@
   \includegraphics{figures/frequency_analysis.pdf}
-  \caption{Frequency analysis of the  input (black) and output (red)
-                for average speeds of 0.8, 1.2, and 1.6 m/s, respectively}
+  \caption{Frequency analysis of the treadmill belt velocity input signal 
+                 (purple) and the recorded output velocity (blue)
+                for average belt speeds of 0.8, 1.2, and 1.6 m/s, respectively}
   \label{fig:freq_analysis}
@@
 
+\section{Acknowledgments}
+We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.  The work was partially funded by the
+State of Ohio Third Frontier Commission through the Wright Center for Sensor Systems Engineering (WCSSE).
+
 \end{document}

### 2014-12-02 src/unperturbed_perturbed_comparison.py
Now using all perturbed gait cycles for the comparison plots.

@@ -1,7 +1,6 @@
 #!/usr/bin/env python
 
 import os
-import random
 
 import yaml
 from numpy import rad2deg
@@ -147,7 +146,8 @@ if __name__ == '__main__':
               'font.family': 'serif',
               'font.serif': ['Computer Modern'],
               'figure.figsize': (6.0, 6.0 / golden),
-             }
+              }
+
     plt.rcParams.update(params)
 
     trial_number = '020'
@@ -174,13 +174,11 @@ if __name__ == '__main__':
     perturbed_gait_data = load_data('Longitudinal Perturbation', paths, tmp)
 
     # Time series comparison plot.
-    num_gait_cycles = unperturbed_gait_cycles.shape[0]
+    num_unperturbed_cycles = unperturbed_gait_cycles.shape[0]
+    num_perturbed_cycles = perturbed_gait_data.gait_cycles.shape[0]
 
-    idxs = random.sample(range(perturbed_gait_data.gait_cycles.shape[0]),
-                         num_gait_cycles)
-
-    mean_of_perturbed = perturbed_gait_data.gait_cycles.iloc[idxs].mean(axis='items')
-    std_of_perturbed = perturbed_gait_data.gait_cycles.iloc[idxs].std(axis='items')
+    mean_of_perturbed = perturbed_gait_data.gait_cycles.mean(axis='items')
+    std_of_perturbed = perturbed_gait_data.gait_cycles.std(axis='items')
 
     mean_of_unperturbed = unperturbed_gait_cycles.mean(axis='items')
     std_of_unperturbed = unperturbed_gait_cycles.std(axis='items')
@@ -216,11 +214,13 @@ if __name__ == '__main__':
 
                 col_name = side + '.' + col
 
+                sigma_mul = 3.0
+
                 pmean = con[i](mean_of_perturbed[col_name])
-                pstd = 2.0 * con[i](std_of_perturbed[col_name])
+                pstd = sigma_mul * con[i](std_of_perturbed[col_name])
 
                 umean = con[i](mean_of_unperturbed[col_name])
-                ustd = 2.0 * con[i](std_of_unperturbed[col_name])
+                ustd = sigma_mul * con[i](std_of_unperturbed[col_name])
 
                 row[i].fill_between(ppercent,
                                     (pmean - pstd).values,
@@ -237,9 +237,9 @@ if __name__ == '__main__':
                                     label='_nolegend_')
 
[…]

### 2014-12-04 paper.tex
Added a title: closes #32.
Introduced the F1000 sty file.

@@
 
-\documentclass{article}
+\documentclass[10pt,a4paper,twocolumn]{article}
+\usepackage{f1000_styles}
 
@@
 \usepackage{booktabs} % for table /toprule, /midrule, etc
-\usepackage{todo}
 
-\title{Perturbed Gait Data}
+% I get a conflict with the todo stuff in f1000_styles if this is loaded.
+%\usepackage{todo}
 
 \begin{document}
+
+\title{An elaborate data set on human gait and the effect of mechanical
+perturbations}
+
+\author[1]{Jason K. Moore}
+\author[1]{Sandra K. Hnat}
+\author[1]{Antonie van den Bogert}
+
+\affil[1]{Mechanical Engineering, Cleveland State University, Cleveland, Ohio,
+  USA, 44115. j.k.moore19@csuohio.edu, s.hnat@vikes.csuohio.edu,
+  a.vandenbogert@csuohio.edu}
+
+\maketitle
+\thispagestyle{fancy}
 
@@
 \end{abstract}
+\listoftodos[F1000Research review comments] % Ignore until review stage
+\clearpage
 
-\section{Introduction}
+\section*{Introduction}
 %
@@
 
-\section{Equipment}
+\section*{Equipment}
 %
@@
 
-\section{Raw Data}
+\section*{Raw Data}
 %
@@
 
-\subsection{mocap-xxx.txt}
+\subsection*{mocap-xxx.txt}
 
@@
 
-\subsection{record-xxx.txt}
+\subsection*{record-xxx.txt}
 %
@@
 
-\subsection{meta-xxx.yml}
+\subsection*{meta-xxx.yml}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \small
@@
   \label{fig:example-meta-data}
-\end{figure}
+\end{figure*}
 
@@
 
-\subsection{Markers}
+\subsection*{Markers}
 %
@@
 
-\begin{table}
+\begin{table*}
   \caption{Descriptions of the 47 markers used in this study.}
@@
   \label{tab:marker-labels}
-\end{table}
+\end{table*}
 
-\section{Perturbation Signals}
+\section*{Perturbation Signals}
 %
@@
 were generated in the event the testing protocol expanded from a walking study
-to a running study, though they were not used in the data set.  The actual motion
-of the treadmill belts are compared to the original input signals as shown in 
-Figure~\ref{fig:input_output}.   The standard deviations of the outputs were 
-calculated at 6.07\%, 9.68\%, and 12.68\% for the the 0.8, 1.2, and 
-1.6 $m \cdot s^{-1}$ speeds, respectively.
+to a running study, though they were not used in the data set.  The actual
+motion of the treadmill belts are compared to the original input signals as
+shown in Figure~\ref{fig:input_output}.   The standard deviations of the
+outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the the 0.8, 1.2,
+and 1.6 $m \cdot s^{-1}$ speeds, respectively.
 
-\begin{figure}
+\begin{figure*}
   \includegraphics{figures/input_vs_output.pdf}
@@
   \label{fig:input_output}
-\end{figure}
+\end{figure*}
 
@@
 transformed into the frequency domain through the Fast Fourier algorithm, and
-the results are shown in Figure~\ref{fig:freq_analysis}.  It was observed that for
-the 1.2 $m \cdot s^{-1}$ average walking speed, the amplitude of the output is 
-significantly lower than the amplitude of the input signal at lower frequencies.  
-Additionally, the amplitude of output signal in the 0.8 $m \cdot s^{-1}$ walking
-speed begins to attenuate around 2 Hz, which is a noticeably lower frequency than
-the other walking speeds.
+the results are shown in Figure~\ref{fig:freq_analysis}. It was observed that
+for the 1.2 $m \cdot s^{-1}$ average walking speed, the amplitude of the output
+is significantly lower than the amplitude of the input signal at lower
+frequencies. Additionally, the amplitude of output signal in the 0.8 $m \cdot
+s^{-1}$ walking speed begins to attenuate around 2 Hz, which is a noticeably
+lower frequency than the other walking speeds.
 
-\begin{figure}
+\begin{figure*}
   \includegraphics{figures/frequency_analysis.pdf}
@@
   \label{fig:freq_analysis}
-\end{figure}
+\end{figure*}
 
-\section{Processed Data}
+\section*{Processed Data}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \includegraphics{figures/unperturbed-perturbed-comparison.pdf}
@@
   \label{fig:angle-torque-comparison}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
@@
   \label{fig:gait-cycle-stats-comparison}
-\end{figure}
+\end{figure*}
 
-\section{Conclusion}
+\section*{Conclusion}
 
@@
 
-\section{Acknowledgments}
-We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.  The work was partially funded by the
-State of Ohio Third Frontier Commission through the Wright Center for Sensor Systems Engineering (WCSSE).
+\subsection*{Author contributions}
+A.v.d.B. conceived of the experiments and protocol. J.K.M and S.K.H developed
+the protocol, ran the experiments, and collected the data. J.K.M was the
+primary author of the paper with significant contributions from S.K.H and
+A.v.d.B. All authors were involved in the revision of the draft manuscript and
+have agreed to the final content.
+
+\subsection*{Competing interests}
+The authors have no  financial, personal, or professional competing interests
+that could be construed to unduly influence the content of this article.
+
+\subsection*{Grant Information}
+The work was partially funded by the State of Ohio Third Frontier Commission
+through the Wright Center for Sensor Systems Engineering (WCSSE).
+
+\subsection*{Acknowledgments}
+We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.
+
+%{\small\bibliographystyle{unsrt}
+%\bibliography{bibfile}}
 

### 2014-12-05 paper.tex
Added GATK Python snippets, closes issue #24.

@@
 
+% graphicx is also declared in f1000_styles
 \usepackage{graphicx}
-\usepackage{booktabs} % for table /toprule, /midrule, etc
+
+% for nice table formatting, i.e. /toprule, /midrule, etc
+\usepackage{booktabs}
+
+% to allow for \verb++ declarations in captions.
+\usepackage{cprotect}
 
@@
 %\usepackage{todo}
+
+% This allows Figures to be renamed as Listing, as required by F1000Research.
+\usepackage{float}
+\newfloat{listing}{thp}{lop}
+\floatname{listing}{Listing}
 
@@
   \includegraphics{figures/frequency_analysis.pdf}
-  \caption{Frequency analysis of the treadmill belt velocity input signal
-                 (purple) and the recorded output velocity (blue)
-                for average belt speeds of 0.8, 1.2, and 1.6 m/s, respectively}
+  \caption{Frequency spectrum of the treadmill belt velocity input signal
+    (purple) and the recorded output velocity (blue) for average belt speeds of
+    0.8, 1.2, and 1.6 m/s, respectively}
   \label{fig:freq_analysis}
@@
 
-We developed a command line data processing tool for common gait computations
-and provide the processed data from this. The tool was developed in Pythin and centered around two main
-classes, one to store a complete trial in memory and a second to compute common
-gait variables of interest.
+We developed a toolkit for data processing, GaitAnalysisToolKit, for common
+gait computations and provide an example processed trial to show the character
+of the data. The tool was developed in Python, is dependent on the SciPy Stack,
+and provides two main classes: one to do basic gait data cleaning,
+\verb|DFlowData|, and a second to compute common gait variables of interest,
+\verb|GaitData|.
 
-The \verb|motek.DFlowData| class collects and stores all the raw data presented
-in the previous section and applies several ``cleaning'' operations to put the
-data into a usable form.
+The \verb|DFlowData| class collects and stores all the raw data presented in
+the previous section and applies several ``cleaning'' operations to transform
+the data into a usable form. The cleaning process follows these basic steps:
 
-\begin{itemize}
+\begin{enumerate}
   \item Loads the meta data file into a Python dictionary.
@@
     with interpolated estimates using a variety of interpolation methods.
-  \item Loads the D-Flow record moudle TSV file into a Pandas \verb|DataFrame|.
+  \item Loads the D-Flow record module TSV file into a Pandas \verb|DataFrame|.
   \item Extracts the events and creates a dictionary mapping the event names in
     the meta data to the events detected in the record module file.
-  \item Compensates the ground reaction loads for treadmill motion.
+  \item Compensates the ground reaction loads based on whether the meta data
+    indicates there was treadmill motion.
   \item Merges the data from the mocap module and record module into one data
     frame at the maximum common constant sample rate.
-  \item Extracts sections of the data bounded by the events.
-  \item Saves the cleaned data to disk.
+\end{enumerate}
+
+Once the data is cleaned there are two methods that allow you to extract the
+cleaned data:
+
+\begin{itemize}
+  \item Extract sections of the data bounded by the events recorded in the
+    \verb|record-XXX.txt| file.
+  \item Save the cleaned data to disk.
 \end{itemize}
 
-The \verb|gait.WalkingData| class is then used to compute things such as gait
-landmarks (toe off and heel strike times), basic 2D inverse dynamics, and to
-store the data into a Pandas \verb|Panel| with each gait cycle on the main axis
-at a specified sample rate. This data is also saved to disk in HDF5 format.
+An example of the API is given in Listing~\ref{lis:dflowdata}.
 
-Figure~\ref{fig:angle-torque-comparison} compares the mean and standard
-deviation of saggital plane joint angles and torques from the perturbed gait
-cycles and the unperturbed gait cycles computed from trial 20 for 120 gait
-cycles. The variation in the dynamics is larger for the perturbed gait cycles.
+\begin{listing*}
+  \begin{verbatim}
+    >>> from gaitanalysis.motek import DFlowData
+    >>> data = DFlowData('mocap-020.txt', 'record-020.txt', 'meta-020.yml')
+    >>> mass = data.meta['subject']['mass']
+    >>> data.clean()
+    >>> event_df = dflow_data.extract_processed_data(event='Longitudinal Perturbation')
+  \end{verbatim}
+  \cprotect\caption{Python interpreter session showing how one could load a
+    trial into memory, extract the subject's mass from the meta data, run the
+    data cleaning process, and finally extract a Pandas \verb|DataFrame|
+    containing all of the time histories for a specific event in the trial.}
+  \label{lis:dflowdata}
+\end{listing*}
+
+The \verb|GaitData| class is then used to compute things such as gait landmarks
+(toe off and heel strike times), basic 2D inverse dynamics, and to store the
+data into a Pandas \verb|Panel| with each gait cycle on the main axis at a
+specified sample rate. This data is also saved to disk in HDF5 format.
+
+\begin{listing*}
+  \begin{verbatim}
+    >>> from gaitanalysis.gait import GaitData
+    >>> gdata = GaitData(event_df)
+    >>> gdata.inverse_dynamics_2d(left_markers, right_markers, left_loads,
+    ...                           right_loads, mass, 6.0)
+    >>> gdata.grf_landmarks('Right Fy', 'Left Fy', threshhold=20.0)
+    >>> gdata.split_at('right')
+    >>> gdata.plot_gait_cycles('Left Hip Joint Torque', mean=True)
+    >>> gdata.save('gait-data.h5')
+  \end{verbatim}
+  \cprotect\caption{Python interpreter session showing how one could compute
+    the inverse dynamics (joint angles and torques), indentify the gait
+    landmarks (e.g. heelstrikes), split the data with respect to the gait
+    landmarks in a Pandas \verb|Panel|, plot the mean and standard deviation of
+    one time history with respect to the gait cycles, and save the data to disk.}
+  \label{lis:dflowdata}
+\end{listing*}
+
+A similar workflow was used to produce Figure~\ref{fig:angle-torque-comparison}
+which compares the mean and standard deviation of saggital plane joint angles
+and torques from the perturbed gait cycles and the unperturbed gait cycles
+computed from trial 20 for 120 gait cycles. The variation in the dynamics is
+larger for the perturbed gait cycles.
 
 \begin{figure*}
+  \centering
   \includegraphics{figures/unperturbed-perturbed-comparison.pdf}
-  \caption{Mean (solid, dashed) and $2\sigma$ (shaded) ($N=120$) joint angles
-    and torques from both unperturbed (purple) and perturbed (blue) gait cycles
-    from trial 20.}
+  \cprotect\caption{Mean (right: solid, left: dashed) and $3\sigma$ (shaded)
+    joint angles and torques from both unperturbed (purple) and perturbed
+    (blue) gait cycles from trial 20. Produced by \verb|unperturbed_perturbed_comparison.py|.}
   \label{fig:angle-torque-comparison}
@@
 \begin{figure*}
+  \centering
   \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
-  \caption{Box plots of the average belt speed, stride frequency, and stride
-    length which compare 120 gait cycles for the unperturbed (purple) and
-    perturbed (blue).}
+  \cprotect\caption{Box plots of the average belt speed, stride frequency, and
+    stride length which compare 120 gait cycles for the unperturbed (purple) and
+    perturbed (blue). Produced by \verb|unperturbed_perturbed_comparison.py|.}
   \label{fig:gait-cycle-stats-comparison}

### 2014-12-09 paper.tex
Using minted for the listings and source code syntax highlighting.

@@
 
+% for nice source code syntax highlighting, also provides Listing env
+\usepackage{minted}
+
 % I get a conflict with the todo stuff in f1000_styles if this is loaded.
 %\usepackage{todo}
-
-% This allows Figures to be renamed as Listing, as required by F1000Research.
-\usepackage{float}
-\newfloat{listing}{thp}{lop}
-\floatname{listing}{Listing}
 
@@
 \begin{listing*}
-  \begin{verbatim}
+  \begin{minted}{python}
     >>> from gaitanalysis.motek import DFlowData
@@
     >>> event_df = dflow_data.extract_processed_data(event='Longitudinal Perturbation')
-  \end{verbatim}
+  \end{minted}
   \cprotect\caption{Python interpreter session showing how one could load a
@@
 \begin{listing*}
-  \begin{verbatim}
+  \begin{minted}{python}
     >>> from gaitanalysis.gait import GaitData
@@
     >>> gdata.save('gait-data.h5')
-  \end{verbatim}
+  \end{minted}
   \cprotect\caption{Python interpreter session showing how one could compute

### 2014-12-09 paper.tex
Added participants section and table, closes #29.
- Added source to generate participants table.
- Added the generated table to the repo.
- Moved functions from py files into new utils.py.
- Added small text section in paper for participants info and embedded the
  table.

@@
 
-We collected over five and half hours of perturbed gait data from 11 subjects
+We collected over five and half hours of perturbed gait data from 15 subjects
 sampled at 100 hz. The data has been organized and made available for other
 research uses.
+
+\section*{Particpants}
+%
+Fifteen able bodied subjects including four females and eleven males with an
+average age of $24\pm4$ years, height of $1.75\pm0.09$ m, mass of $74\pm13$ kg
+particpated in the study. The data has been anonymized with respect to the
+participants' identities and an unique identification number was assigned to
+each subject. A selection of the meta data collected for each subject is shown
+in Table~\ref{tab:subjects}.
+%
+\begin{table*}
+  \cprotect\caption{Information about the 15 participants. The final three
+    columns give the trial numbers associated with each nominal treadmill
+    speed. Additional trials with subject 0 are trials with no subject, i.e.
+    unloaded trials that can be used for force compensation purposes, and are
+    not shown in the table. Generated by \verb|src/subject_table.py|.}
+  \centering
+  \input{tables/subjects.tex}
+  \label{tab:subjects}
+\end{table*}
 

### 2014-12-09 src/unperturbed_perturbed_comparison.py

@@ -7,130 +7,10 @@ from numpy import rad2deg
 from scipy.constants import golden
 import matplotlib.pyplot as plt
 import pandas
-from gaitanalysis.motek import DFlowData, markers_for_2D_inverse_dynamics
-from gaitanalysis.gait import GaitData
 from gaitanalysis.utils import _percent_formatter
 import seaborn as sbn
 
-
-def trial_file_paths(trials_dir, trial_number):
-    """Returns the most common paths to the trials in the gait
-    identification data set.
-
-    Parameters
-    ==========
-    trials_dir : string
-        The path to the main directory for the data. This directory should
-        contain subdirectories: `T001/`, `T002/`, etc.
-    trial_number : string
-        Three digit trial number, e.g. `005`.
-
-    """
-
-    trial_dir = 'T' + trial_number
-    mocap_file = 'mocap-' + trial_number + '.txt'
-    record_file = 'record-' + trial_number + '.txt'
-    meta_file = 'meta-' + trial_number + '.yml'
-
-    mocap_file_path = os.path.join(trials_dir, trial_dir, mocap_file)
-    record_file_path = os.path.join(trials_dir, trial_dir, record_file)
-    meta_file_path = os.path.join(trials_dir, trial_dir, meta_file)
-
-    return mocap_file_path, record_file_path, meta_file_path
-
-
-def load_data(event, paths, tmp):
-    """Loads an event and processes the data, if necessary, from a trial
-    into a GaitData object.
-
-    Parameters
-    ==========
-    event : string
-        A valid event for the given trial.
-    paths : list of strings
-        The paths to the mocap, record, and meta data files.
-    tmp : string
-        A path to a temporary directory in which the processed data can be
-        stored.
-
-    Returns
-    =======
-    gait_data : gaitanalysis.gait.GaitData
-        The GaitData instance containing the data for the event.
-
-    """
-
-    file_name = '_'.join([n.lower() for n in event.split(' ')]) + '.h5'
-
-    tmp_data_path = os.path.join(tmp, file_name)
-
-    try:
-        f = open(tmp_data_path, 'r')
-    except IOError:
-        print('Cleaning and processing {} data...'.format(event))
-        # Load raw data, clean it up, and extract the perturbation section.
-        dflow_data = DFlowData(*paths)
-        dflow_data.clean_data(ignore_hbm=True)
-        perturbed_df = \
-            dflow_data.extract_processed_data(event=event,
[…]

### 2014-12-09 paper.tex
Metadata description overhaul, closes #21.
- A shorter, more complete example meta file is included.
- Descriptions of every possible meta data item are included.
- meta data section is reorganized.

--- 2014-12-09 paper.tex
@@
 
-Each trial directory contains a meta data file in the YAML format. There are
-three main headings: \verb+study+, \verb+subject+, and \verb+trial+.
+Each trial directory contains a meta data file in the YAML format named in the
+following style \verb|meta-xxx.yml| where \verb|xxx| is the three digit trial
+identification number. There are three main headings in the file: \verb+study+,
+\verb+subject+, and \verb+trial+. An example meta data file is shown in Listing
+\ref{lis:example-meta-data}.
 
@@
 study, an identification number, name, and description. This is the same for
-all meta data files in the study.
+all meta data files in the study. Details are given below:
+
+\begin{description}
+  \item[id] An integer specifying a unique identification number of the study.
+  \item[name] A string giving the name of the study.
+  \item[description] A string with a basic description of the study.
+\end{description}
 
 The \verb+subject+ provides key value pairs of information about the subject in
-that trial. Each subject has a unique identification number and basic
-anthropomorphic data is included.
+that trial. Each subject has a unique identification number along with basic
+anthropomorphic data. The following details the possible meta data for the
+subject:
 
-The \verb+trial+ section contains the unique information about the trial. Each
-trial has a unique identification number.
+\begin{description}
+  \item[age] An integer age in years of the subject at the time of the trial.
+  \item[ankle-width-left] A float specifying the width of the subjects left
+    ankle.
+  \item[ankle-width-right] A float specifying the width of the subjects right
+    ankle.
+  \item[ankle-width-units] A string giving the units of measurement of the
+    ankle widths.
+  \item[id] An unique identification integer for the subject.
+  \item[gender] A string specifying the gender of the subject.
+  \item[height] A float specifying the measured height of the subject (with
+    shoes on) at the time of the trial.
+  \item[height-units] A string giving the units of the height measurement.
+  \item[knee-width-left] A float specifying the width of the subjects left
+    knee.
+  \item[knee-width-right] A float specifying the width of the subjects right
+    knee.
+  \item[knee-width-units] A string giving the units of measurement of the
+    knee widths.
+  \item[mass] A float specifying the self-reported mass of the subject.
+  \item[mass-units] A string specifying the units of the mass measurement.
+\end{description}
 
-An example meta data file is shown in Figure \ref{fig:example-meta-data} and
-the following gives and explanation of the contents.
+The \verb+trial+ section contains the information about the particular trial.
+Each trial has a unique identification number along with a variety of other
+information, detailed below:
 
-\begin{figure*}
+\begin{description}
+  \item[analog-channel-map] A mapping of the strings D-Flow assigns to signals
+    emmitted from the analog channels of the NI USB-6255 to names the user
+    desires.
+  \item[cortex-version] The version of Cortex used to record the trial.
+  \item[datetime]A date formatted string giving the date of the trial in the
+    \verb|YYYY-MM-DD| format.
+  \item[dflow-version] The version of D-Flow used to record the trial.
+  \item[events] A key value map which prescribes names to the alphabetic events
+    recorded in the record file.
+  \item[files] A key value mapping of files associated with this trial
+    where the key is the DFlow file type and the value is the path to the file
+    relative to the meta file.
+  \item[hardware-settings] There are tons of settings for the hardware in both
+    D-Flow, Cortex, and the other software in the system. This contains any
+    non-default settings.
+    \begin{description}
+      \item[high-performance] A boolean value indicatin whether the D-Flow high
+        performance setting was on (True) or off (False).
+    \end{description}
+  \item[id] An unique three digit integer identifier for the trial. All of the
+    file names and directories associated with this trial include this number.
+  \item[marker-map] A key value map which maps marker names in the mocap file
+    to the user's desired names for the markers.
+  \item[marker-set] Indicates the HBM~\cite{Bogert2013} marker set used during
+    the trial, either full, lower, or NA.
+  \item[nominal-speed] A float representing the nominal desired treadmill speed
+    during the trial.
+  \item[nominal-speed-units] (string) The units of the nominal speed.
+  \item[notes] Any notes about the trial.
+  \item[pitch] A boolean that indicates if the treamdill pitch degree of
+    freedom was acutated during the trial.
+  \item[stationary-platform] A boolean that indicates whether the treadmill
+      sway or pitch motion was actuated during the trial. If this flag is
+      false, the measured ground reaction forces must be compensated for the
+      inertial affects and be expressed in the motion capture reference frame.
+  \item[subject-id] An integer corresponding to the subject in the trial.
+  \item[sway] A boolean that indicates if the treadmill lateral (sway) degree
+    of freedom was acutated during the trial.
+\end{description}
+
+\begin{listing*}
   \small
-  \begin{verbatim}
+  \begin{minted}[gobble=4]{yaml}
     study:
@@
     subject:
-            id: 0
+       id: 8
+       age: 20
+       mass: 70.0
+       mass-units: kilograms
+       height: 1.572
+       height-units: meters
+       knee-width-left: 107.43
+       knee-width-right: 107.41
+       knee-width-units: millimeters
+       ankle-width-left: 70.52
+       ankle-width-right: 67.66
+       ankle-width-units: millimeters
+       gender: male
     trial:
        id: 58
-            subject-id: 0
+       subject-id: 8
        datetime: 2014-03-28
        notes: >
-                    This is an unloaded trial (no subject) 10 minute protocol that
-                    perturbs the subject longitudinally. The treadmill reference
-                    markers were not in place. We have the four custom made metal
-                    feet placed under the four corners of the treadmill to prevent
-                    pitch and sway movements. In addition, we also have the front
-                    left and rear right feet that were installed by Motek. We also
-                    have +/-3g analog accelerometers (ADXL330) affixed to the four
-                    corners of the main frame of the treadmill and wired into the
-                    NI box. The right handrail is removed.
+           The subject did a somersault during this trial instead of following
+           instructions to walk. Will have to use for another study.
        nominal-speed: 0.8
@@
            high-performance: True
+       dflow-version: 3.16.1
+       cortex-version: 3.1.1.1290
+       marker-map:
+           M1: LHEAD
+           M2: THEAD
+           M3: RHEAD
+           M4: FHEAD
+           M5: C7
        analog-channel-map:
@@
            Channel4.Anlg: F1X1
-                    Channel5.Anlg: F1X2
-                    Channel6.Anlg: F1Z1
-                    Channel7.Anlg: F2Y1
-                    Channel8.Anlg: F2Y2
-                    Channel9.Anlg: F2Y3
-                    Channel10.Anlg: F2X1
-                    Channel11.Anlg: F2X2
-                    Channel12.Anlg: F2Z1
-                    Channel61.Anlg: Front_Left_AccX
-                    Channel62.Anlg: Front_Left_AccY
-                    Channel63.Anlg: Front_Left_AccZ
-                    Channel64.Anlg: Back_Left_AccX
-                    Channel65.Anlg: Back_Left_AccY
-                    Channel66.Anlg: Back_Left_AccZ
-                    Channel67.Anlg: Front_Right_AccX
-                    Channel68.Anlg: Front_Right_AccY
-                    Channel69.Anlg: Front_Right_AccZ
-                    Channel70.Anlg: Back_Right_AccX
-                    Channel71.Anlg: Back_Right_AccY
-                    Channel72.Anlg: Back_Right_AccZ
        events:
@@
        files:
+           compensation: ../T057/mocap-057.txt
            mocap: mocap-058.txt
@@
            meta: meta-058.yml
-  \end{verbatim}
-  \caption{An example meta data YAML file.}
-  \label{fig:example-meta-data}
-\end{figure*}
-
-\begin{description}
-  \item[analog-channel-map] A key value mapping of strings specifying the name
-    of the signal emmitted from the analog channels of the NI USB XXX.
-  \item[cortex-version] (string) Matches the version of Cortex used to record
-    the trial.
-  \item[datetime] (string) A date formatted string giving the date and/or time
-    of the trial in the \verb|YYYY-MM-DD| format.
-  \item[dflow-version] (string) Matches the version of D-Flow used to record
-    the trial.
-  \item[events] A key value map which prescribes names to the events recorded
-    in the record file.
-  \item[files] (list) A key value mapping of files associated with this trial
-    where the key is the file type and the values are the path to the file
-    relative to this meta file.
-  \item[hardware-settings] There are tons of settings for the hardware in both
-    D-Flow, Cortex, and the other software in the system. This contains any
-    non-default settings.
-    \begin{description}
-      \item[high-performance] (boolean) Indicates whether the D-Flow high
-        performance setting was on (True) or off (False).
-    \end{description}
-  \item[id] (integer) A unique three digit identifier for the trial. All of the
-    file names and directories associated with this trial include this.
-  \item[marker-map] A string key value map which contains names for the
-    markers.
-  \item[marker-set] (string) Indicates the HBM \cite{Bogert2013} marker
-    set used during the trial, either full, lower, or NA.
-  \item[nominal-speed] (float) The nominal or base speed during the trial.
-  \item[nominal-speed-units] (string) The units of the nominal speed.
-  \item[notes] {string} Notes about the trial.
-  \item[pitch] (boolean) Indicates if the pitch degree of freedom was acutated
-    during the trial.
-  \item[stationary-platform] (boolean) Indicates whether the treadmill
-      sway or pitch motion was actuated during the trial. If this flag is
-      false, the measured ground reaction forces must be compensated for the
-      inertial affects and be expressed the forces and moments in the motion
-      capture reference frame.
-  \item[subject-id] (integer) A number corresponding to the subject in the
-    trial.
-  \item[sway] (boolean) Indicates if the lateral (sway) degree of freedom was
-    acutated during the trial.
-\end{description}
+  \end{minted}
+  \caption{A fictious example of a YAML formatted meta data file. All of the
+    possible keys in the data set are shown.}
+  \label{lis:example-meta-data}
+\end{listing*}
 

### 2014-12-09 paper.tex
Closes #34, abstract improved.

@@
 
-  Herein we share a typical gait data set collected from 11 subjects walking at
-  a three speeds on an instrumented treadmill while being longitudinally
-  perturbed during each stance phase with pseudo-random fluctuations in the
-  speed of the treadmill belt. We provide raw marker and ground reaction loads in addition
-  to a presentation of processed data that includes gait landmarks, 2D joint
-  angles, angular rates, and joint torques and the software that does the
-  processing. The protocol is described in detail along with the additional
-  meta data about each of the data files. This data can be useful for
-  validating or generating mathematical models that capable of simulating
-  non-periodic and perturbed gaits.
+  Here we share a rich gait data set collected from fifteen subjects walking at
+  three speeds on an instrumented treadmill. Each trial consists of 120 seconds
+  of normal walking and 480 seconds of walking while being longitudinally
+  and/or laterally perturbed during each stance phase with pseudo-random
+  fluctuations in the speed of the treadmill belt. A total of approximately 1.5
+  hours of normal walking ($>5000$ gait cycles) and 6 hours of perturbed walking
+  ($>20,000$ gait cycles) is included in the data set. We provide raw marker
+  locations and ground reaction loads in addition to a presentation of
+  processed data that includes gait landmarks, 2D joint angles, angular rates,
+  and joint torques and the software used for the computations. The protocol is
+  described in detail along with additional elaborate meta data. This data can
+  likely be useful for validating or generating mathematical models that are
+  capable of simulating normal periodic gait and non-periodic, perturbed gaits.
 

### 2014-12-09 paper.tex
Added 'full body' to abstract.

@@
   hours of normal walking ($>5000$ gait cycles) and 6 hours of perturbed walking
-  ($>20,000$ gait cycles) is included in the data set. We provide raw marker
-  locations and ground reaction loads in addition to a presentation of
+  ($>20,000$ gait cycles) is included in the data set. We provide full body raw
+  marker locations and ground reaction loads in addition to a presentation of
   processed data that includes gait landmarks, 2D joint angles, angular rates,

### 2014-12-09 paper.tex
Minor edits.

@@
 \begin{listing*}
-  \begin{minted}{python}
+  \begin{minted}[gobble=4]{python}
     >>> from gaitanalysis.motek import DFlowData
@@
     >>> mass = data.meta['subject']['mass']
-    >>> data.clean()
+    >>> data.clean_data()
     >>> event_df = dflow_data.extract_processed_data(event='Longitudinal Perturbation')
@@
 \begin{listing*}
-  \begin{minted}{python}
+  \begin{minted}[gobble=4]{python}
     >>> from gaitanalysis.gait import GaitData
@@
 \subsection*{Author contributions}
-A.v.d.B. conceived of the experiments and protocol. J.K.M and S.K.H developed
-the protocol, ran the experiments, and collected the data. J.K.M was the
-primary author of the paper with significant contributions from S.K.H and
-A.v.d.B. All authors were involved in the revision of the draft manuscript and
-have agreed to the final content.
+A.v.d.B. conceived of the experiments and protocol. J.K.M and S.K.H refined the
+protocol, ran the experiments, and collected the data. J.K.M was the primary
+author of the paper with significant contributions from S.K.H and A.v.d.B. All
+authors were involved in the revision of the draft manuscript and have agreed
+to the final content.
 

### 2014-12-10 paper.tex

@@
   hours of normal walking ($>5000$ gait cycles) and 6 hours of perturbed walking
-  ($>20,000$ gait cycles) is included in the data set. We provide full body raw
+  ($>20,000$ gait cycles) is included in the data set. We provide full body
   marker locations and ground reaction loads in addition to a presentation of
@@
 \begin{itemize}
-  \item A ForceLink R-Mill which has dual 6 degree of freedom force plates,
-    independent belts for each foot, and lateral and pitch motion capabilities.
-  \item A 10 Camera Motion Analysis motion capture system which includes Osprey
-    cameras, the Cortex software, and hardware for collecting analog and camera
-    data simultaneously. Delsys wireless EMG + 3D Accelerometers.
-  \item Motek Medical’s D-Flow software and visual display system.
+  \item A ForceLink R-Mill treadmill which has dual 6 degree of freedom force
+    plates, independent belts for each foot, and motion capabilities (lateral,
+    pitch).
+  \item A 10 camera Motion Analysis motion capture system which includes Osprey
+    cameras, the Cortex software (3.1.1.1290), and hardware for collecting
+    analog and camera data simultaneously.
+  \item National Instruments USB-6255 data aquisition unit.
+  \item Four Sparkfun ADXL330 Triple Axis Accelerometer Breakout boards
+    attached to the treadmill.
+  \item Motek Medical’s D-Flow software (3.16.1 to 3.16.2) and visual display
+    system.
 \end{itemize}
@@
 plates, and analog sensors (EMG/Accelerometer) through a National Instruments
-XXXX data acquisition unit. D-Flow is required to collect data from any digital
-sensors and the treadmill's motion (lateral, pitch, and belts). D-Flow can
-output multiple types of files which contain the different data.
+USB-6255 data acquisition unit. D-Flow is required to collect data from any
+digital sensors and the treadmill's motion (lateral, pitch, and belts). D-Flow
+can output multiple types of files which contain the different data.
 
@@
 system.
+
+\section*{Protocol}
+%
+The experimental protocol consisted of static measurements and walking on the
+treadmill for 10 minutes under unperturbed and perturbed conditions. Before a
+set of trials on the same day the following happened:
+%
+\begin{itemize}
+  \item Calibration of the motion capture system.
+  \item Subject changes into atheltic shoes, lyrca shorts, sports bra, baseball
+    cap, and rock climbing harness.
+  \item All 47 markers are applied directly to the skin except for the heel,
+    toe, and head markers.~\footnote{The sacrum and rear pelvic markers were
+    placed on the shorts for a small number of the subjects}.
+  \item Subjects self-reported age, gender, and mass.
+  \item Height is measured by the experimentalist.
+  \item Four reference photographs (front, back, right, left) are taken of
+    subjects marker locations.
+\end{itemize}
+
+After a briefing by the expermentalist on the trial protocol the subject
+follows the verbal instructions of the experimentalist and the on-screen
+instructions from the video display in front of the subject. The protocal goes
+as so:
+%
+\begin{enumerate}
+  \item Subject steps onto the treadmill and markers are identified with
+    Cortex.
+  \item The safety cable is attached loosely to the rock climbing harness such
+    that no undue forces are acting on the subject during walking, but that the
+    harness will prevent a full fall.
+  \item The subject starts by stepping on sides of treadmill so that feet don't
+    touch the force plates and the force plate signals are zero'd.
+  \item Once notified by the screen, the subject stands in the initialization
+    pose: standing straight up, looking forward, arms out by their sides (~45
+    degrees) and the event is manually recorded by the operator.
+  \item A countdown to the first normal walking phase is displayed. At the end
+    of the countodwn the treadmill starts at the specified speed and the
+    subject is instructed to walk normally, to focus on the road in the
+    screen, and not to look at their feet.
+  \item After 1 minute of normal walking, the longitudinal perturbation phase
+    begins.
+  \item After 8 minutes of walking under the influence of the perturbations,
+    the normal walking phase begins.
+  \item After 1 minute, a countdown is shown and the treadmill decelerates to a
+    stop.
+  \item The subject is instructed to step off of the force plates for 10
+    seconds.
+  \item The subject could take a break before each additional trial.
+\end{enumerate}
+
+\section*{Perturbation Signals}
+%
+The data set uses a testing protocol containing a segment of normal walking,
+followed by a series of pseudo-random longitudinal perturbations, and ending
+with a second segment of normal walking. Three pseudo-random speed control
+signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
+generated through MATLAB and Simulink. Initially, a  random acceleration signal
+was created using discrete-time Gaussian white noise and saturated at the
+maximum belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt
+speed, and high-pass filtered (second-order Butterworth) to eliminate velocity
+drift. Mean velocity was then added to the signal and limited between 0 and 3.6
+m/s. The cutoff frequencies of the high-pass filter, as well as the variance in
+the acceleration signal, were adjusted until the appropriate standard
+deviations for each mean velocity were obtained.
+
+These standard deviations were determined through a trial and error method.
+Initially, perturbations of $\pm$ 10\% for each mean velocity were generated.
+However, this large deviation in the slowest mean velocity (0.8 $m \cdot
+s^{-1}$) would unintentionally reduce the treadmill speed to values close to
+zero. This would produce an undesired stopping and starting sensation.
+Alternatively, smaller standard perturbations of $\pm$ 5\% did not produce
+noticeable perturbations in the fastest trial (1.6 $m \cdot s^{-1}$).
+Therefore, standard deviations of 7.16\%, 10.23\%, and 13.1\% were selected for
+the 0.8, 1.2, and 1.6 $m \cdot s^{-1}$ speeds, respectively.  This ensured that
+the test subject was sufficiently perturbed for each speed, while remaining
+within the limits of our testing protocol.
+
+To ensure that the treadmill could follow the higher frequencies of the random
+signals, the maximum treadmill acceleration of 15 $m \cdot s^{-2}$ was used in
+the D-Flow Software. As this value remains constant throughout the experiment,
+the initial starting and stopping of the treadmill would be too rapid for test
+subjects. To eliminate this, a ramped acceleration and deceleration was
+generated and placed at the beginning and end of the entire signal,
+respectively. Other additions include the one minute segments of unperturbed
+walking at the beginning and end of the trial and 30 seconds of no movement for
+force plate zeroing and subject T-pose calibration.  The sampling rate of the
+simulation was set to 100 Hz, corresponding to the D-Flow software's
+capabilities.
+
+The resulting output from the MATLAB script and Simulink produces a comma-
+delimited text file of six columns: timestamp, slow, normal, and fast walking
+perturbation signals, and slow and fast running signals.  These running signals
+were generated in the event the testing protocol expanded from a walking study
+to a running study, though they were not used in the data set.  The actual
+motion of the treadmill belts are compared to the original input signals as
+shown in Figure~\ref{fig:input_output}.   The standard deviations of the
+outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the the 0.8, 1.2,
+and 1.6 $m \cdot s^{-1}$ speeds, respectively.
+
+\begin{figure*}
+  \centering
+  \includegraphics{figures/input_vs_output.pdf}
+  \caption{Treadmill belt velocity input signals (purple) and recorded output
+    velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
+    respectively}
+  \label{fig:input_output}
+\end{figure*}
+
+For further analysis, the input and output signals for each average speed were
+transformed into the frequency domain through the Fast Fourier algorithm, and
+the results are shown in Figure~\ref{fig:freq_analysis}. It was observed that
+for the 1.2 $m \cdot s^{-1}$ average walking speed, the amplitude of the output
+is significantly lower than the amplitude of the input signal at lower
+frequencies. Additionally, the amplitude of output signal in the 0.8 $m \cdot
+s^{-1}$ walking speed begins to attenuate around 2 Hz, which is a noticeably
+lower frequency than the other walking speeds.
+
+\begin{figure*}
+  \centering
+  \includegraphics{figures/frequency_analysis.pdf}
+  \caption{Frequency spectrum of the treadmill belt velocity input signal
+    (purple) and the recorded output velocity (blue) for average belt speeds of
+    0.8, 1.2, and 1.6 m/s, respectively}
+  \label{fig:freq_analysis}
+\end{figure*}
 
@@
 
-\section*{Perturbation Signals}
-%
-The data set uses a testing protocol containing a segment of normal walking,
-followed by a series of pseudo-random longitudinal perturbations, and ending
-with a second segment of normal walking. Three pseudo-random speed control
-signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
-generated through MATLAB and Simulink. Initially, a  random acceleration signal
-was created using discrete-time Gaussian white noise and saturated at the maximum
-belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt speed, and 
-high-pass filtered (second-order Butterworth) to eliminate velocity drift. Mean 
-velocity was then added to the signal and limited between 0 and 3.6 m/s. The 
-cutoff frequencies of the high-pass filter, as well as the variance in the acceleration 
-signal, were adjusted until the appropriate standard deviations for each mean 
-velocity were obtained.
-
-These standard deviations were determined through a trial and error method.
-Initially, perturbations of $\pm$ 10\% for each mean velocity were generated.
-However, this large deviation in the slowest mean velocity (0.8 $m \cdot
-s^{-1}$) would unintentionally reduce the treadmill speed to values close to
-zero. This would produce an undesired stopping and starting sensation.
-Alternatively, smaller standard perturbations of $\pm$ 5\% did not produce
-noticeable perturbations in the fastest trial (1.6 $m \cdot s^{-1}$).
-Therefore, standard deviations of 7.16\%, 10.23\%, and 13.1\% were selected for
-the 0.8, 1.2, and 1.6 $m \cdot s^{-1}$ speeds, respectively.  This ensured that
-the test subject was sufficiently perturbed for each speed, while remaining
-within the limits of our testing protocol.
-
-To ensure that the treadmill could follow the higher frequencies of the random
-signals, the maximum treadmill acceleration of 15 $m \cdot s^{-2}$ was used in
-the D-Flow Software. As this value remains constant throughout the experiment,
-the initial starting and stopping of the treadmill would be too rapid for test
-subjects. To eliminate this, a ramped acceleration and deceleration was
-generated and placed at the beginning and end of the entire signal,
-respectively. Other additions include the one minute segments of unperturbed
-walking at the beginning and end of the trial and 30 seconds of no movement for
-force plate zeroing and subject T-pose calibration.  The sampling rate of the
-simulation was set to 100 Hz, corresponding to the D-Flow software's
-capabilities.
-
-The resulting output from the MATLAB script and Simulink produces a comma-
-delimited text file of six columns: timestamp, slow, normal, and fast walking
-perturbation signals, and slow and fast running signals.  These running signals
-were generated in the event the testing protocol expanded from a walking study
-to a running study, though they were not used in the data set.  The actual
-motion of the treadmill belts are compared to the original input signals as
-shown in Figure~\ref{fig:input_output}.   The standard deviations of the
-outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the the 0.8, 1.2,
-and 1.6 $m \cdot s^{-1}$ speeds, respectively.
-
-\begin{figure*}
-  \includegraphics{figures/input_vs_output.pdf}
-  \caption{Treadmill belt velocity input signals (purple) and recorded output
-    velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
-    respectively}
-  \label{fig:input_output}
-\end{figure*}
-
-For further analysis, the input and output signals for each average speed were
-transformed into the frequency domain through the Fast Fourier algorithm, and
-the results are shown in Figure~\ref{fig:freq_analysis}. It was observed that
-for the 1.2 $m \cdot s^{-1}$ average walking speed, the amplitude of the output
-is significantly lower than the amplitude of the input signal at lower
-frequencies. Additionally, the amplitude of output signal in the 0.8 $m \cdot
-s^{-1}$ walking speed begins to attenuate around 2 Hz, which is a noticeably
-lower frequency than the other walking speeds.
-
-\begin{figure*}
-  \includegraphics{figures/frequency_analysis.pdf}
-  \caption{Frequency spectrum of the treadmill belt velocity input signal
-    (purple) and the recorded output velocity (blue) for average belt speeds of
-    0.8, 1.2, and 1.6 m/s, respectively}
-  \label{fig:freq_analysis}
-\end{figure*}
-
 \section*{Processed Data}

### 2014-12-10 paper.tex
First stab at data limitations, closes #41.

@@
 
+\section*{Data Limitations}
+%
+The data is provided in good faith with great attention to detail. The
+following list gives various notes and warnings about the data that should be
+taken into account when making use of it.
+%
+\begin{itemize}
+  \item Be sure to read the notes in each meta data file for details about
+    possible anomalies in that particular trial.
+  \item The subject identification number 0 stands for "no subject" and was
+    used whenever data was collected from the system with no subject on the
+    treadmill, for example during the trials that were intended to be used for
+    inertial compensation purposes if the treadmill was moving. These trials
+    play through the exact protocol as those with a human subject and the
+    matching trials are indicated in the meta data.
+  \item Trials 1 and 2 were not recorded as part of this study. Those trial
+    identification numbers were reserved for early data exploration from data
+    collected in other studies.
+  \item Trials 37, 38, and 39 do not exist. The numbers were accidentally
+    skipped.
+  \item Trials 9, 10, and 11 used a slightly different event definition where the
+    calibration poses were not explicitly tagged by an event, yet the protocol
+    was the identical to the following trials.
+  \item During trials 9-15 we used wooden blocks to fix the treadmill to the
+    concrete floor to eliminate the treadmill's lowest factory vibration mode
+    (~5 hz). But these blocks seem to have corrupted the force plate
+    measurements by imposing undue stresses on the system. The force plate
+    measurements should not be used from these trials, but the marker data is
+    fine.
+  \item We did not record unloaded compensation trials for trials 9-15.
+    Regardless, they would be useless due to the previous point.
+  \item Trials 6-8 use an early experimental protocol which divided the
+    perturbation sections into three sections: longitudinal perturbations,
+    lateral perturbations, and a combination of each. We then learned the
+    treadmill had a low vibrational mode which significantly affects the force
+    plate measurements. The force measurements during these trials are
+    corrupted by this vibrational mode and should be used with caution or not
+    at all.
+  \item Trials 6-8 use a XX lower body marker set only.
+  \item The ankle joint torques computed from subject 9's data in trials 25-27
+    are abnormal and should be used with caution or not at all. We were not
+    about to locate the source of the error, but it is likely related to the
+    force calibration.
+\end{itemize}
+
 \section*{Conclusion}

### 2014-12-10 src/unperturbed_perturbed_comparison.py
Complete Makefile from the root is actionable, closes #18.
- Added Python and Vim stuff to .gitignore
- Developed a complete Makefile in the root directory.
- Moved the outline into its own file.
- Created build instructions and moved into the root README.
- Added a default config file.
- Regenerated the figures and tables.
- Removed the old config file.
- Added print statements to python scripts and did some cleanup.
- Created a function to load the paths in the config file.
- Added Octave link to README.

@@ -1,8 +1,12 @@
 #!/usr/bin/env python
 
+"""This script generates two plots. The first compares the mean and standard
+deviation of the planar lower extremity joint angles and joint torques from
+perturbed and unperturbed gait cycles. The second is a box plot comparing
+basic gait stats for perturbed and unperturbed gait cycles."""
+
 import os
 
-import yaml
 from numpy import rad2deg
 from scipy.constants import golden
 import matplotlib.pyplot as plt
@@ -10,148 +14,155 @@ import pandas
 from gaitanalysis.utils import _percent_formatter
 import seaborn as sbn
 
-from utils import trial_file_paths, load_data, remove_bad_gait_cycles
+from utils import (trial_file_paths, load_data, remove_bad_gait_cycles,
+                   config_paths)
 
+script_path = os.path.realpath(__file__)
+src_dir = os.path.dirname(script_path)
+root_dir = os.path.realpath(os.path.join(src_dir, '..'))
+raw_dir, tmp = config_paths(root_dir)
 
-if __name__ == '__main__':
+params = {'backend': 'ps',
+          'axes.labelsize': 8,
+          'axes.titlesize': 10,
+          'font.size': 10,
+          'legend.fontsize': 8,
+          'xtick.labelsize': 8,
+          'ytick.labelsize': 8,
+          'text.usetex': True,
+          'font.family': 'serif',
+          'font.serif': ['Computer Modern'],
+          'figure.figsize': (6.0, 6.0 / golden),
+          }
 
-    params = {'backend': 'ps',
-              'axes.labelsize': 8,
-              'axes.titlesize': 10,
-              'text.fontsize': 10,
-              'legend.fontsize': 8,
-              'xtick.labelsize': 8,
-              'ytick.labelsize': 8,
-              'text.usetex': True,
-              'font.family': 'serif',
-              'font.serif': ['Computer Modern'],
-              'figure.figsize': (6.0, 6.0 / golden),
-              }
+plt.rcParams.update(params)
 
-    plt.rcParams.update(params)
+trial_number = '020'
+paths = trial_file_paths(raw_dir, trial_number)
 
-    trial_number = '020'
+unperturbed_gait_data_1 = load_data('First Normal Walking', paths, tmp)
+unperturbed_gait_data_2 = load_data('Second Normal Walking', paths, tmp)
+one = remove_bad_gait_cycles(unperturbed_gait_data_1, 1.18, 1.22,
[…]

### 2014-12-10 src/utils.py

@@ -10,6 +10,33 @@ from gaitanalysis.motek import DFlowData, markers_for_2D_inverse_dynamics
 from gaitanalysis.gait import GaitData
 
 
+def config_paths(root_dir):
+    """Returns the full path to the directories specified in the config.yml
+    file.
+
+    Parameters
+    ----------
+    root_dir : string
+        Absolute path to the root directory of the repository.
+
+    Returns
+    -------
+    raw_dir : string
+        Absolute path to the raw data directory.
+    processed_dir : string
+        Absolute path to the processed data directory.
+
+    """
+
+    with open(os.path.join(root_dir, 'config.yml'), 'r') as f:
+        config = yaml.load(f)
+
+    raw_dir = os.path.join(root_dir, config['raw_data_dir'])
+    processed_dir = os.path.join(root_dir, config['processed_data_dir'])
+
+    return raw_dir, processed_dir
+
+
 def trial_file_paths(trials_dir, trial_number):
     """Returns the most common paths to the trials in the gait
     identification data set.

### 2014-12-11 paper.tex

@@
 \begin{table*}
-  \caption{Descriptions of the 47 markers used in this study.}
+  \cprotect\caption{Descriptions of the 47 markers used in this study. The Set
+    column indicates wheter the marker exists in the lower and/or full body
+    marker set.  The label column matches the column headers in the
+    \verb|mocap-xxx.txt| files and/or the marker map in the \verb|meta-xxx.yml|
+    file.}
   \centering
   \small
-  \begin{tabular}{rlll}
+  \begin{tabular}{lrlll}
     \toprule
-    Number & Label & Name & Description \\
+    Set & \# & Label & Name & Description \\
     \midrule
-    1  & LHEAD & Left head                             & Just above the ear, in the middle. \\
-    2  & THEAD & Top head                              & On top of the head, in line with the LHEAD and RHEAD. \\
-    3  & RHEAD & Right head                            & Just above the ear, in the middle. \\
-    4  & FHEAD & Forehead                              & Between line LHEAD/RHEAD and THEAD a bit right from center. \\
-    5  & C7    & C7                                    & On the 7th cervical vertebrae. \\
-    6  & T10   & T10                                   & On the 10th thoracic vertbrae. \\
-    7  & SACR  & Sacrum bone                           & on the scacral bone. \\
-    8  & NAVE  & Navel                                 & On the navel. \\
-    9  & XYPH  & Xiphoid process                       & Xiphoid process of the sternum. \\
-    10 & STRN  & Sternum                               & On the jugular notch of the sternum. \\
-    11 & BBAC  & Scapula                               & On the inferior angle fo the right scapular. \\
-    12 & LSHO  & Left shoulder                         & Left acromion. \\
-    13 & LDELT & Left deltoid muscle                   & Apex of the deltoid muscle. \\
-    14 & LLEE  & Left lateral elbow                    & Left lateral epicondyle of the elbow. Upper one in the T-Pose. \\
-    15 & LMEE  & Left medial elbow                     & Left medial epicondyle of the elbow. Lower on in the T-Pose. \\
-    16 & LFRM  & Left forearm                          & On 2/3 on the line between the LLEE and LMW. \\
-    17 & LMW   & Left medial wrist                     & On styloid process radius, thumb side. \\
-    18 & LLW   & Left lateral wrist                    & On styloid process ulna, pinky side. \\
-    19 & LFIN  & Left fingers                          & Center of the hand. Caput metatarsal 3. \\
-    20 & RSHO  & Right shoulder                        & Right acromion. \\
-    21 & RDELT & Right deltoid muscle                  & Apex of deltoid muscle. \\
-    22 & RLEE  & Right lateral elbow                   & Right lateral epicondyle of the elbow. Lower one in the T-pose. \\
-    23 & RMEE  & Right medial elbow                    & Right medial epicondyle of the elbow. Lower one in the T-pose. \\
-    24 & RFRM  & Right forearm                         & On 1/3 on the line between the RLEE and RMW. \\
-    25 & RMW   & Right medial wrist                    & On styloid process radius, thumb side. \\
-    26 & RLW   & Right lateral wrist                   & On styloid process ulna, pinky side. \\
-    27 & RFIN  & Right fingers                         & Center of the hand. Caput metatarsal 3. \\
-    28 & LASIS & Pelvic bone left front                & Left anterior superior iliac spine. \\
-    29 & RASIS & Pelvic bone right front               & Right anterior superior iliac spine. \\
-    30 & LPSIS & Pelvic bone left back                 & Left posterior superio iliac spine. \\
-    31 & RPSIS & Pelvic bone right back                & Right posterior superior iliac spine. \\
-    32 & LGTRO & Left greater trochanter of the femur  & On the cetner of the left greater trochanter. \\
-    33 & FLTHI & Left thigh                            & On 1/3 on the line between the LFTRO and LLEK. \\
-    34 & LLEK  & Left lateral epicondyle of the knee   & On the lateral side of the joint axis. \\
-    35 & LATI  & Left anterior of the tibia            & On 2/3 on the line between the LLEK and LLM. \\
-    36 & LLM   & Left lateral malleoulus of the ankle  & The center of the heel at the same height as the toe. \\
-    37 & LHEE  & Left heel                             & Center of the heel at the same height as the toe. \\
-    38 & LTOE  & Left toe                              & Tip of big toe. \\
-    39 & LMT5  & Left 5th metatarsal                   & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
-    40 & RGTRO & Right greater trochanter of the femur & On the cetner of the right greater trochanter. \\
-    41 & FRTHI & Right thigh                           & On 2/3 on the line between the RFTRO and RLEK. \\
-    42 & RLEK  & Right lateral epicondyle of the knee  & On the lateral side of the joint axis. \\
-    43 & RATI  & Right anterior of the tibia           & On 1/3 on the line between the RLEK and RLM. \\
-    44 & RLM   & Right lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
-    45 & RHEE  & Right heel                            & Center of the heel at the same height as the toe. \\
-    46 & RTOE  & Right toe                             & Tip of big toe. \\
-    47 & RMT5  & Right 5th metatarsal                  & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
+    F   & 1  & LHEAD & Left head                             & Just above the ear, in the middle. \\
+    F   & 2  & THEAD & Top head                              & On top of the head, in line with the LHEAD and RHEAD. \\
+    F   & 3  & RHEAD & Right head                            & Just above the ear, in the middle. \\
+    F   & 4  & FHEAD & Forehead                              & Between line LHEAD/RHEAD and THEAD a bit right from center. \\
+    L/F & 5  & C7    & C7                                    & On the 7th cervical vertebrae. \\
+    L/F & 6  & T10   & T10                                   & On the 10th thoracic vertbrae. \\
+    L/F & 7  & SACR  & Sacrum bone                           & On the scacral bone. \\
+    L/F & 8  & NAVE  & Navel                                 & On the navel. \\
+    L/F & 9  & XYPH  & Xiphoid process                       & Xiphoid process of the sternum. \\
+    F   & 10 & STRN  & Sternum                               & On the jugular notch of the sternum. \\
+    F   & 11 & BBAC  & Scapula                               & On the inferior angle fo the right scapular. \\
+    F   & 12 & LSHO  & Left shoulder                         & Left acromion. \\
+    F   & 13 & LDELT & Left deltoid muscle                   & Apex of the deltoid muscle. \\
+    F   & 14 & LLEE  & Left lateral elbow                    & Left lateral epicondyle of the elbow. Upper one in the T-Pose. \\
+    F   & 15 & LMEE  & Left medial elbow                     & Left medial epicondyle of the elbow. Lower on in the T-Pose. \\
+    F   & 16 & LFRM  & Left forearm                          & On 2/3 on the line between the LLEE and LMW. \\
+    F   & 17 & LMW   & Left medial wrist                     & On styloid process radius, thumb side. \\
+    F   & 18 & LLW   & Left lateral wrist                    & On styloid process ulna, pinky side. \\
+    F   & 19 & LFIN  & Left fingers                          & Center of the hand. Caput metatarsal 3. \\
+    F   & 20 & RSHO  & Right shoulder                        & Right acromion. \\
+    F   & 21 & RDELT & Right deltoid muscle                  & Apex of deltoid muscle. \\
+    F   & 22 & RLEE  & Right lateral elbow                   & Right lateral epicondyle of the elbow. Lower one in the T-pose. \\
+    F   & 23 & RMEE  & Right medial elbow                    & Right medial epicondyle of the elbow. Lower one in the T-pose. \\
+    F   & 24 & RFRM  & Right forearm                         & On 1/3 on the line between the RLEE and RMW. \\
+    F   & 25 & RMW   & Right medial wrist                    & On styloid process radius, thumb side. \\
+    F   & 26 & RLW   & Right lateral wrist                   & On styloid process ulna, pinky side. \\
+    F   & 27 & RFIN  & Right fingers                         & Center of the hand. Caput metatarsal 3. \\
+    L/F & 28 & LASIS & Pelvic bone left front                & Left anterior superior iliac spine. \\
+    L/F & 29 & RASIS & Pelvic bone right front               & Right anterior superior iliac spine. \\
+    L/F & 30 & LPSIS & Pelvic bone left back                 & Left posterior superio iliac spine. \\
+    L/F & 31 & RPSIS & Pelvic bone right back                & Right posterior superior iliac spine. \\
+    L/F & 32 & LGTRO & Left greater trochanter of the femur  & On the cetner of the left greater trochanter. \\
+    L/F & 33 & FLTHI & Left thigh                            & On 1/3 on the line between the LFTRO and LLEK. \\
+    L/F & 34 & LLEK  & Left lateral epicondyle of the knee   & On the lateral side of the joint axis. \\
+    L/F & 35 & LATI  & Left anterior of the tibia            & On 2/3 on the line between the LLEK and LLM. \\
+    L/F & 36 & LLM   & Left lateral malleoulus of the ankle  & The center of the heel at the same height as the toe. \\
+    L/F & 37 & LHEE  & Left heel                             & Center of the heel at the same height as the toe. \\
+    L/F & 38 & LTOE  & Left toe                              & Tip of big toe. \\
+    L/F & 39 & LMT5  & Left 5th metatarsal                   & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
+    L/F & 40 & RGTRO & Right greater trochanter of the femur & On the cetner of the right greater trochanter. \\
+    L/F & 41 & FRTHI & Right thigh                           & On 2/3 on the line between the RFTRO and RLEK. \\
+    L/F & 42 & RLEK  & Right lateral epicondyle of the knee  & On the lateral side of the joint axis. \\
+    L/F & 43 & RATI  & Right anterior of the tibia           & On 1/3 on the line between the RLEK and RLM. \\
+    L/F & 44 & RLM   & Right lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
+    L/F & 45 & RHEE  & Right heel                            & Center of the heel at the same height as the toe. \\
+    L/F & 46 & RTOE  & Right toe                             & Tip of big toe. \\
+    L/F & 47 & RMT5  & Right 5th metatarsal                  & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
     \bottomrule

### 2014-12-11 paper.tex
Reworked intro with new citations, closes #36
- new bib file (includes some refs I have yet to cite, but maybe should)
- Different intro flavor in light of lack of gait data.

@@
 \author[1]{Sandra K. Hnat}
-\author[1]{Antonie van den Bogert}
+\author[1]{Antonie J. van den Bogert}
 
@@
   ($>20,000$ gait cycles) is included in the data set. We provide full body
-  marker locations and ground reaction loads in addition to a presentation of
+  marker trajectories and ground reaction loads in addition to a presentation of
   processed data that includes gait landmarks, 2D joint angles, angular rates,
@@
 %
-Emerging powered prosthetics are in need of bio-inspired control systems.
-Ideally, a powered prosthetic would behave in such a way that the user would
-feel like their limb was never disabled. There are a variety of approaches to
-developing bio-inspired control systems some of which aim to mimic the
-reactions and motion of an able-bodied person. A modern gait lab is able to
-collect a variety of kinematic, kinetic, and physiological data from humans
-during gait. This data can potentially be used to drive the design of the human
-mimicking controller. With a rich enough data set, one may be able to identify
-control mechanism used during a human's natural gait and recovery from
-perturbations. We have collected data that we believe is richer than previous
-gait data sets and may be rich enough for control identification. The data can
-also be used for verification purposes for controllers that have been designed
-in other manners.
+The collection of dynamical data during human walking has a long history
+beginning with the first motion pictures and now with modern marker based
+motion capture techniques, image and video analysis, and high fidelity ground
+reaction load measurements. Even though years of data on thousands of subjects
+now exist, this data is not widely disseminated, well organized, nor available
+with few or no restrictions. David Winter's published normative gait
+data~\cite{Winter1990} is widely used in biomechanics studies, yet it comes
+from relatively few subjects and only a small number of gait cycles per
+subject. This small source has successfully inspired powered prosthetic control
+design~\cite{GoldfarbXXXX}, but success in other fields using large sets of
+data for discovery lead one to believe that may be possible in the
+biomechanical field too. To find out if that hypothesis is true, our data needs
+to be shared extensively, organized, and curated to enabled analysts.
 
-We collected over five and half hours of perturbed gait data from 15 subjects
-sampled at 100 hz. The data has been organized and made available for other
-research uses.
+There are some notable gait data sets and databases besides Winter's authoritative
+set that are publicly available. Notably the International Society of
+Biomechanics has maintained a web page (http://isbweb.org/data) since
+approximately 1995 that includes data sets for download and most unencumbered
+use. Vaughn, et. al's data~\cite{Vaughan1992} with kinematics and force plate
+measurement from several subjects. Chris Kirtley shares normative gait from
+several studies on the CGA Normative Gait Database~\cite{Kirtley} and these
+files have influenced and been included in other studies, for
+example the average gait cycles from children used in~\cite{Bogert2003}.
 
-\section*{Particpants}
+Chester et. al~\cite{Chester2007} report on a large gait database comparison
+where one database contained kinematic data of 409 gait cycles of children
+from 1 to 7 years old. Tirosh et. al, recognized the need of a comprehensive
+data base for clinical gait data and created the Gaitabase~\cite{Tirosh2010}.
+This database may contain a substantial amount of data but it is encumbered by
+a very complicated and restrictive license and sharing scheme. The University
+of Wisconsin at LaCrosse has a normative gait data set~\cite{Wilson} from 25
+subjects with lower extremity marker data and force plate measurements from a
+single stance phase.
+
+More recent examples of biomechanist's sharing their data alongside
+publications are: \cite{Bogert2013} which includes processed joint kinematics
+from eleven subjects walking for few gait cycles and \cite{Wang2014} who
+includes data from 10 subjects walking for five minutes each at five different
+speeds but only a small set of lower extremity markers are present. The second
+is notable because it publishes the data in Dryad, a modern citable data
+repository.
+
+The publicly available gait data is small compared to the number of gait
+studies that have been performed over the years. And the data that is available
+in general suffers from things such as few subjects, few gait cycles, few
+markers, no raw data, limited force plate measurements, lack of meta data, and
+restrictive licensing. To help with this situation we are making the data we
+collected for our research purposes publicly available and free of the
+previously mentioned deficiencies. Not only do we provide a larger set of
+normative gait data that has been previously available we include an even large
+set of data in which the subject is being perturbed which does not exist at
+all. We believe both of these sets of data can serve a variety of use cases and
+hope that we can save time for researchers with use cases for this data.
+
+Our use case for the data is centered around emerging powered prosthetics and
+orthotics need of bio-inspired control systems. Ideally, a powered prosthetic
+would behave in such a way that the user would feel like their limb was never
+disabled. There are a variety of approaches to developing bio-inspired control
+systems some of which aim to mimic the reactions and motion of an able-bodied
+person. A modern gait lab is able to collect a variety of kinematic, kinetic,
+and physiological data from humans during gait. This data can potentially be
+used to drive the design of the human-mimicking controller. With a rich enough
+data set, one may be able to identify control mechanisms used during a human's
+natural gait and recovery from perturbations. We have collected data that we
+believe is richer than previous gait data sets and may be rich enough for
+control identification.  The data can also be used for verification purposes
+for controllers that have been designed in other manners.
+
+With all of this in mind, we collected over five and half hours of gait data
+from fifteen able bodied subjects and includes over 25,000 gait cycles. Most of
+the subjects walked at three different speeds on an instrumented treadmill
+while we collected full body marker locations and ground reaction loads from a
+pair of force plates. The protocol for the majority of the trials includes two
+minutes of normal walking and eight minutes of walking under the influence of
+pseudo-random belt speed fluctuations. The data has been organized complete with
+rich meta data and made available in the most unrestrictive form for other
+research uses following modern best practices in data sharing~\cite{White2013}.
+
+Furthermore, we include a small Apache licensed open source software library
+for basic gait analysis and demonstrate it's use in the paper. The combination
+of the open data and open software allow the results presented within to be
+computationally reproducible and instructions are included in the associated
+repository for doing so.
+
+\section*{Participants}
 %
@@
 average age of $24\pm4$ years, height of $1.75\pm0.09$ m, mass of $74\pm13$ kg
-particpated in the study. The data has been anonymized with respect to the
+participated in the study. The data has been anonymized with respect to the
 participants' identities and an unique identification number was assigned to
@@
     analog and camera data simultaneously.
-  \item National Instruments USB-6255 data aquisition unit.
+  \item National Instruments USB-6255 data acquisition unit.
   \item Four Sparkfun ADXL330 Triple Axis Accelerometer Breakout boards
@@
   \item Calibration of the motion capture system.
-  \item Subject changes into atheltic shoes, lyrca shorts, sports bra, baseball
+  \item Subject changes into athletic shoes, shorts, sports bra, baseball
     cap, and rock climbing harness.
@@
 
-After a briefing by the expermentalist on the trial protocol the subject
+After a briefing by the experimentalist on the trial protocol the subject
 follows the verbal instructions of the experimentalist and the on-screen
-instructions from the video display in front of the subject. The protocal goes
+instructions from the video display in front of the subject. The protocol goes
 as so:
@@
   \item The subject starts by stepping on sides of treadmill so that feet don't
-    touch the force plates and the force plate signals are zero'd.
+    touch the force plates and the force plate signals are zeroed.
   \item Once notified by the screen, the subject stands in the initialization
@@
   \item A countdown to the first normal walking phase is displayed. At the end
-    of the countodwn the treadmill starts at the specified speed and the
+    of the countdown the treadmill starts at the specified speed and the
     subject is instructed to walk normally, to focus on the road in the
@@
 The resulting output from the MATLAB script and Simulink produces a comma-
-delimited text file of six columns: timestamp, slow, normal, and fast walking
+delimited text file of six columns: time stamp, slow, normal, and fast walking
 perturbation signals, and slow and fast running signals.  These running signals
@@
 shown in Figure~\ref{fig:input_output}.   The standard deviations of the
-outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the the 0.8, 1.2,
+outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the 0.8, 1.2,
 and 1.6 $m \cdot s^{-1}$ speeds, respectively.
@@
 
-The text file output from the mocap module in DFlow is a tab delimited file.
+The text file output from the mocap module in D-Flow is a tab delimited file.
 The first line is the header and contains a time stamp column, frame number
@@
   \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
-    recieves a frame from Cortex. These are approximately at 100 hz and given
+    receives a frame from Cortex. These are approximately at 100 hz and given
     in seconds.
@@
   \item[analog-channel-map] A mapping of the strings D-Flow assigns to signals
-    emmitted from the analog channels of the NI USB-6255 to names the user
+    emitted from the analog channels of the NI USB-6255 to names the user
     desires.
@@
   \item[files] A key value mapping of files associated with this trial
-    where the key is the DFlow file type and the value is the path to the file
+    where the key is the D-Flow file type and the value is the path to the file
     relative to the meta file.
@@
     \begin{description}
-      \item[high-performance] A boolean value indicatin whether the D-Flow high
+      \item[high-performance] A boolean value indicating whether the D-Flow high
         performance setting was on (True) or off (False).
@@
   \item[notes] Any notes about the trial.
-  \item[pitch] A boolean that indicates if the treamdill pitch degree of
-    freedom was acutated during the trial.
+  \item[pitch] A boolean that indicates if the treadmill pitch degree of
+    freedom was actuated during the trial.
   \item[stationary-platform] A boolean that indicates whether the treadmill
@@
   \item[sway] A boolean that indicates if the treadmill lateral (sway) degree
-    of freedom was acutated during the trial.
+    of freedom was actuated during the trial.
 \end{description}
@@
   \end{minted}
-  \caption{A fictious example of a YAML formatted meta data file. All of the
+  \caption{A fictitious example of a YAML formatted meta data file. All of the
     possible keys in the data set are shown.}
@@
 %
-We make use of the full body 47 marker set described in \cite{Ton's HBM paper}
+We make use of the full body 47 marker set described in \cite{Bogert2013}
 and presented in detail in Table \ref{tab:marker-labels}. As with all camera
@@
   \cprotect\caption{Descriptions of the 47 markers used in this study. The Set
-    column indicates wheter the marker exists in the lower and/or full body
+    column indicates whether the marker exists in the lower and/or full body
     marker set.  The label column matches the column headers in the
@@
     L/F & 6  & T10   & T10                                   & On the 10th thoracic vertbrae. \\
-    L/F & 7  & SACR  & Sacrum bone                           & On the scacral bone. \\
+    L/F & 7  & SACR  & Sacrum bone                           & On the sacral bone. \\
     L/F & 8  & NAVE  & Navel                                 & On the navel. \\
@@
   \item Loads the meta data file into a Python dictionary.
-  \item Loads the DFlow mocap module TSV file into Pandas \verb|DataFrame|.
+  \item Loads the D-Flow mocap module TSV file into Pandas \verb|DataFrame|.
   \item Relabels the column headers to more meaningful names if this is
@@
   \cprotect\caption{Python interpreter session showing how one could compute
-    the inverse dynamics (joint angles and torques), indentify the gait
-    landmarks (e.g. heelstrikes), split the data with respect to the gait
+    the inverse dynamics (joint angles and torques), identify the gait
+    landmarks (e.g. heel strikes), split the data with respect to the gait
     landmarks in a Pandas \verb|Panel|, plot the mean and standard deviation of
@@
 
-A similar workflow was used to produce Figure~\ref{fig:angle-torque-comparison}
+A similar work flow was used to produce Figure~\ref{fig:angle-torque-comparison}
 which compares the mean and standard deviation of saggital plane joint angles
@@
 
-%{\small\bibliographystyle{unsrt}
-%\bibliography{bibfile}}
+{\small\bibliographystyle{unsrt}
+\bibliography{gait-data}}
 

### 2014-12-11 paper.tex
Note that Chester's data isn't available.

--- 2014-12-11 paper.tex
@@
 Chester et. al~\cite{Chester2007} report on a large gait database comparison
-where one database contained kinematic data of 409 gait cycles of children
-from 1 to 7 years old. Tirosh et. al, recognized the need of a comprehensive
-data base for clinical gait data and created the Gaitabase~\cite{Tirosh2010}.
-This database may contain a substantial amount of data but it is encumbered by
-a very complicated and restrictive license and sharing scheme. The University
-of Wisconsin at LaCrosse has a normative gait data set~\cite{Wilson} from 25
-subjects with lower extremity marker data and force plate measurements from a
-single stance phase.
+where one database contained kinematic data of 409 gait cycles of children from
+1 to 7 years old but the data does not seem to be publicly available. Tirosh
+et. al, recognized the need of a comprehensive data base for clinical gait data
+and created the Gaitabase~\cite{Tirosh2010}.  This database may contain a
+substantial amount of data but it is encumbered by a very complicated and
+restrictive license and sharing scheme. The University of Wisconsin at LaCrosse
+has a normative gait data set~\cite{Wilson} from 25 subjects with lower
+extremity marker data and force plate measurements from a single stance phase.
 

### 2014-12-11 paper.tex
Update paper.tex
A few changes in wording.

--- 2014-12-11 paper.tex
@@
 %
-Emerging powered prosthetics are in need of bio-inspired control systems.
+Emerging powered prosthetics and orthotics are in need of bio-inspired control systems.
 Ideally, a powered prosthetic would behave in such a way that the user would
@@
     analog and camera data simultaneously.
-  \item National Instruments USB-6255 data acquisition unit.
+  \item National Instruments USB-6255 data aquisition unit.
   \item Four Sparkfun ADXL330 Triple Axis Accelerometer Breakout boards
@@
   \item Calibration of the motion capture system.
-  \item Subject changes into athletic shoes, shorts, sports bra, baseball
+  \item Subject changes into atheltic shoes, lyrca shorts, sports bra, baseball
     cap, and rock climbing harness.
@@
 
-After a briefing by the experimentalist on the trial protocol the subject
+After a briefing by the expermentalist on the trial protocol the subject
 follows the verbal instructions of the experimentalist and the on-screen
-instructions from the video display in front of the subject. The protocol goes
+instructions from the video display in front of the subject. The protocal goes
 as so:
@@
   \item The subject starts by stepping on sides of treadmill so that feet don't
-    touch the force plates and the force plate signals are zeroed.
+    touch the force plates and the force plate signals are zero'd.
   \item Once notified by the screen, the subject stands in the initialization
@@
   \item A countdown to the first normal walking phase is displayed. At the end
-    of the countdown the treadmill starts at the specified speed and the
+    of the countodwn the treadmill starts at the specified speed and the
     subject is instructed to walk normally, to focus on the road in the
@@
 The resulting output from the MATLAB script and Simulink produces a comma-
-delimited text file of six columns: time stamp, slow, normal, and fast walking
+delimited text file of six columns: timestamp, slow, normal, and fast walking
 perturbation signals, and slow and fast running signals.  These running signals
@@
 shown in Figure~\ref{fig:input_output}.   The standard deviations of the
-outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the 0.8, 1.2,
+outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the the 0.8, 1.2,
 and 1.6 $m \cdot s^{-1}$ speeds, respectively.
@@
 
-The text file output from the mocap module in D-Flow is a tab delimited file.
+The text file output from the mocap module in DFlow is a tab delimited file.
 The first line is the header and contains a time stamp column, frame number
@@
   \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
-    receives a frame from Cortex. These are approximately at 100 hz and given
+    recieves a frame from Cortex. These are approximately at 100 hz and given
     in seconds.
@@
   \item[analog-channel-map] A mapping of the strings D-Flow assigns to signals
-    emitted from the analog channels of the NI USB-6255 to names the user
+    emmitted from the analog channels of the NI USB-6255 to names the user
     desires.
@@
   \item[files] A key value mapping of files associated with this trial
-    where the key is the D-Flow file type and the value is the path to the file
+    where the key is the DFlow file type and the value is the path to the file
     relative to the meta file.
@@
     \begin{description}
-      \item[high-performance] A boolean value indicating whether the D-Flow high
+      \item[high-performance] A boolean value indicatin whether the D-Flow high
         performance setting was on (True) or off (False).
@@
   \item[notes] Any notes about the trial.
-  \item[pitch] A boolean that indicates if the treadmill pitch degree of
-    freedom was actuated during the trial.
+  \item[pitch] A boolean that indicates if the treamdill pitch degree of
+    freedom was acutated during the trial.
   \item[stationary-platform] A boolean that indicates whether the treadmill
@@
   \item[sway] A boolean that indicates if the treadmill lateral (sway) degree
-    of freedom was actuated during the trial.
+    of freedom was acutated during the trial.
 \end{description}
@@
   \end{minted}
-  \caption{A fictitious example of a YAML formatted meta data file. All of the
+  \caption{A fictious example of a YAML formatted meta data file. All of the
     possible keys in the data set are shown.}
@@
   \cprotect\caption{Descriptions of the 47 markers used in this study. The Set
-    column indicates whether the marker exists in the lower and/or full body
+    column indicates wheter the marker exists in the lower and/or full body
     marker set.  The label column matches the column headers in the
@@
     L/F & 6  & T10   & T10                                   & On the 10th thoracic vertbrae. \\
-    L/F & 7  & SACR  & Sacrum bone                           & On the sacral bone. \\
+    L/F & 7  & SACR  & Sacrum bone                           & On the scacral bone. \\
     L/F & 8  & NAVE  & Navel                                 & On the navel. \\
@@
   \item Loads the meta data file into a Python dictionary.
-  \item Loads the D-Flow mocap module TSV file into Pandas \verb|DataFrame|.
+  \item Loads the DFlow mocap module TSV file into Pandas \verb|DataFrame|.
   \item Relabels the column headers to more meaningful names if this is
@@
   \cprotect\caption{Python interpreter session showing how one could compute
-    the inverse dynamics (joint angles and torques), identify the gait
-    landmarks (e.g. heel strikes), split the data with respect to the gait
+    the inverse dynamics (joint angles and torques), indentify the gait
+    landmarks (e.g. heelstrikes), split the data with respect to the gait
     landmarks in a Pandas \verb|Panel|, plot the mean and standard deviation of
@@
 
-A similar work flow was used to produce Figure~\ref{fig:angle-torque-comparison}
+A similar workflow was used to produce Figure~\ref{fig:angle-torque-comparison}
 which compares the mean and standard deviation of saggital plane joint angles

### 2014-12-11 paper.tex

--- 2014-12-11 paper.tex
@@
 subject. This small source has successfully inspired powered prosthetic control
-design~\cite{GoldfarbXXXX}, but success in other fields using large sets of
+design~\cite{Sup2008}, but success in other fields using large sets of
 data for discovery lead one to believe that may be possible in the

### 2014-12-12 paper.tex
Update paper.tex
Edits by Ton.

@@
   of normal walking and 480 seconds of walking while being longitudinally
-  and/or laterally perturbed during each stance phase with pseudo-random
+  perturbed during each stance phase with pseudo-random
   fluctuations in the speed of the treadmill belt. A total of approximately 1.5
@@
   marker trajectories and ground reaction loads in addition to a presentation of
-  processed data that includes gait landmarks, 2D joint angles, angular rates,
+  processed data that includes gait events, 2D joint angles, angular rates,
   and joint torques and the software used for the computations. The protocol is
@@
 design~\cite{Sup2008}, but success in other fields using large sets of
-data for discovery lead one to believe that may be possible in the
-biomechanical field too. To find out if that hypothesis is true, our data needs
+data for discovery lead one to believe that more elaborate data sets may 
+benefit the field of human motion studies.
+To enable such work, data needs
 to be shared extensively, organized, and curated to enabled analysts.
@@
 has a normative gait data set~\cite{Wilson} from 25 subjects with lower
-extremity marker data and force plate measurements from a single stance phase.
+extremity marker data and force plate measurements from a single stance phase (TvdB: gait cycle?).
 
-More recent examples of biomechanist's sharing their data alongside
+More recent examples of biomechanists sharing their data alongside
 publications are: \cite{Bogert2013} which includes processed joint kinematics
@@
 Furthermore, we include a small Apache licensed open source software library
-for basic gait analysis and demonstrate it's use in the paper. The combination
+for basic gait analysis and demonstrate its use in the paper. The combination
 of the open data and open software allow the results presented within to be
@@
 \section*{Participants}
+The study was approved by the Institutional Review Board of Cleveland State University
+and informed consent was obtained from all participants.
 %
@@
 participated in the study. The data has been anonymized with respect to the
-participants' identities and an unique identification number was assigned to
+participants' identities and a unique identification number was assigned to
 each subject. A selection of the meta data collected for each subject is shown
@@
     speed. Additional trials with subject 0 are trials with no subject, i.e.
-    unloaded trials that can be used for force compensation purposes, and are
+    unloaded trials that can be used for inertial compensation purposes, and are
     not shown in the table. Generated by \verb|src/subject_table.py|.}
@@
 %
-The Human Motion and Control Lab at Cleveland State University has modern human
-motion measurement equipment. Our system includes:
+The data were collected in the Laboratory for Human Motion and Control at Cleveland State University,
+using the following equipment:
 
@@
 
-Cortex alone is capable of delivering data from the cameras along with force
-plates, and analog sensors (EMG/Accelerometer) through a National Instruments
+The Cortex software delivers 3D marker trajectories from the cameras along with data
+from force
+plates and analog sensors (EMG/Accelerometer) through a National Instruments
 USB-6255 data acquisition unit. D-Flow is required to collect data from any
-digital sensors and the treadmill's motion (lateral, pitch, and belts). D-Flow
-can output multiple types of files which contain the different data.
+digital sensors and to control the treadmill's motion (lateral, pitch, and belts). D-Flow
+can process the data in real time and/or export data to file.
 
@@
 
-After a briefing by the experimentalist on the trial protocol the subject
+After obtaining informed consent and a briefing by the experimentalist on the trial protocol, the subject
 follows the verbal instructions of the experimentalist and the on-screen
 instructions from the video display in front of the subject. The protocol goes
-as so:
+as follows:
 %
@@
 The \verb|GaitData| class is then used to compute things such as gait landmarks
-(toe off and heel strike times), basic 2D inverse dynamics, and to store the
+(toe off and heel strike times), basic 2D kinematics and inverse dynamics, and to store the
 data into a Pandas \verb|Panel| with each gait cycle on the main axis at a
@@
     play through the exact protocol as those with a human subject and the
-    matching trials are indicated in the meta data.
+    matching trials are indicated in the meta data. (TvdB: can the user tell
+    which compensation file to use with which subject files?  By the date on
+    the file?)
   \item Trials 1 and 2 were not recorded as part of this study. Those trial
@@
     (~5 hz). But these blocks seem to have corrupted the force plate
-    measurements by imposing undue stresses on the system. The force plate
+    measurements by imposing frictional stresses on the system. The force plate
     measurements should not be used from these trials, but the marker data is

### 2014-12-12 src/subject_table.py
Subject table now has measured mass column, closes #42.
- Added a function in utils.py to compute the measured mass for most subjects.
- Updated script to generate the table with the new column.

@@ -6,16 +6,16 @@ from collections import defaultdict
 import numpy as np
 import pandas as pd
 
-from utils import generate_meta_data_tables, config_paths
+import utils
 
 print('Generating subject table.')
 
 script_path = os.path.realpath(__file__)
 src_dir = os.path.dirname(script_path)
 root_dir = os.path.realpath(os.path.join(src_dir, '..'))
-raw_dir, processed_dir = config_paths(root_dir)
+raw_dir, processed_dir = utils.config_paths(root_dir)
 
-tables = generate_meta_data_tables(raw_dir)
+tables = utils.generate_meta_data_tables(raw_dir)
 
 subject_df = tables['TOP|subject']
 
@@ -55,13 +55,30 @@ formatters = {'Height [m]': lambda x: 'NA' if np.isnan(x)
 unique_subjects = unique_subjects.drop_duplicates()
 unique_subjects = unique_subjects.drop(0)  # remove null subject
 
+measured = utils.measured_subject_mass(raw_dir, processed_dir)
+
+def format_sigma(x):
+    if x[1] >= 1.0:
+        return 'dollar{:0.0f}plusminus{:0.0f}dollar'.format(*x)
+    else:
+        return 'dollar{:0.1f}plusminus{:0.1f}dollar'.format(*x)
+
+measured['Measured Mass [kg]'] = zip(measured['Measured Mass'], measured['Standard Deviation'])
+measured['Measured Mass [kg]'] = measured['Measured Mass [kg]'].map(format_sigma)
+
+unique_subjects['Measured Mass [kg]'] = measured['Measured Mass [kg]']
+new_cols.insert(4, 'Measured Mass [kg]')
+
 table_dir = os.path.join(root_dir, 'tables')
 if not os.path.exists(table_dir):
     os.makedirs(table_dir)
 
 table_path = os.path.join(table_dir, 'subjects.tex')
+tex = unique_subjects.sort().to_latex(na_rep='NA', index=False,
+                                      columns=new_cols,
+                                      formatters=formatters)
+tex = tex.replace('dollar', '$')
+tex = tex.replace('plusminus', '\pm')
 with open(table_path, 'w') as f:
-    f.write(unique_subjects.sort().to_latex(na_rep='NA', index=False,
-                                            columns=new_cols,
-                                            formatters=formatters))
+    f.write(tex)
 print('Table at: {}'.format(table_path))

### 2014-12-12 src/utils.py

@@ -1,8 +1,10 @@
 #!/usr/bin/env python
 
+# standard library
 import os
 from collections import defaultdict
 
+# external
 import yaml
 import numpy as np
 import pandas as pd
@@ -224,3 +226,144 @@ def generate_meta_data_tables(trials_dir, top_level_key='TOP', key_sep='|'):
         tables[k] = pd.DataFrame(v, index=ordered_trial_nums)
 
     return tables
+
+
+def measured_subject_mass(raw_data_dir, processed_data_dir):
+    """This script computes the mean mass of each subject based on the force
+    plate data collected just after the calibration pose. It also compares
+    it to the mass provided by the subject. Some subjects may have invalid
+    measurements and will not be included, so you should make use of the
+    self reported mass.
+
+    Parameters
+    ----------
+    raw_data_dir : string
+        The path to the raw data directory.
+    processed_data_dir : string
+        The path to the processed data directory.
+
+    Returns
+    -------
+    mean : pandas.DataFrame
+        A data frame containing columns with mean/std measured mass, the
+        self reported mass, and indexed by subject id.
+
+    """
+
+    # Subject 0 is for the null subject. For subject 1 we use the self
+    # reported value because there is no "Calibration Pose" event. For
+    # subject 11 and subject 4, we use the self reported mass because the
+    # wooden feet were in place and the force measurements are
+    # untrust-worthy.
+    subj_with_invalid_meas = [0, 1, 4, 11]
+
+    # Some of the trials have anomalies in the data after the calibration
+    # pose due to the subjects' movement. The following gives best estimates
+    # of the sections of the event that are suitable to use in the subjects'
+    # mass computation. The entire time series during the "Calibration Pose"
+    # event is acceptable for trials not listed.
+    time_sections = {'020': (None, 14.0),
+                     '021': (None, 14.0),
+                     '031': (-14.0, None),
+                     '047': (None, 12.0),
+                     '048': (None, 7.0),
+                     '055': (-12.0, None),
+                     '056': (-3.0, None),  # also the first 2 seconds are good
+                     '057': (-8.0, None),
+                     '063': (None, 6.0),  # also the last 6 seconds are good
+                     '069': (None, 14.0),
+                     '078': (None, 15.0)}
+
+    trial_dirs = [x[0] for x in os.walk(raw_data_dir) if x[0][-4] == 'T']
+    trial_nums = [x[-3:] for x in trial_dirs if x[-3:] not in ['001', '002']]
+
[…]

### 2014-12-13 paper.tex
Added image of treadmill and motion capture system, closes #53.
Added IRB #, closes #62.

@@
 \section*{Participants}
-The study was approved by the Institutional Review Board of Cleveland State University
-and informed consent was obtained from all participants.
 %
+The study was approved by the Institutional Review Board of Cleveland State
+University (\# 29904-VAN-HS) and informed consent was obtained from all
+participants.
+
 Fifteen able bodied subjects including four females and eleven males with an
@@
 
-The Cortex software delivers 3D marker trajectories from the cameras along with data
-from force
-plates and analog sensors (EMG/Accelerometer) through a National Instruments
-USB-6255 data acquisition unit. D-Flow is required to collect data from any
-digital sensors and to control the treadmill's motion (lateral, pitch, and belts). D-Flow
-can process the data in real time and/or export data to file.
+The Cortex software delivers 3D marker trajectories from the cameras along with
+data from force plates and analog sensors (EMG/Accelerometer) through a
+National Instruments USB-6255 data acquisition unit. D-Flow is required to
+collect data from any digital sensors and to control the treadmill's motion
+(lateral, pitch, and belts). D-Flow can process the data in real time and/or
+export data to file.
 
@@
 The same point is used as the origin of the ground reaction force measuring
-system.
+system. Figure~\ref{fig:treadmill} shows the layout of the equiptment.
+
+\begin{figure*}
+  \centering
+  \includegraphics{figures/treadmill.png}
+  \caption{The treadmill, cameras, projection screen, and saftey rope. The
+    direction of travel is in the $-z$ direction.}
+  \label{fig:treadmill}
+\end{figure*}
 

### 2014-12-13 paper.tex
Switched to PeerJ template, closes #58.

@@
 
-\documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage{f1000_styles}
+\documentclass[fleqn,10pt]{wlpeerj}
 
-% graphicx is also declared in f1000_styles
+% for images: png, pdf, etc
 \usepackage{graphicx}
@@
 \usepackage{minted}
-
-% I get a conflict with the todo stuff in f1000_styles if this is loaded.
-%\usepackage{todo}
-
-\begin{document}
 
@@
 
-\maketitle
-\thispagestyle{fancy}
+\keywords{gait, data, perturbation}
 
@@
 \end{abstract}
-\listoftodos[F1000Research review comments] % Ignore until review stage
-\clearpage
+
+\begin{document}
+
+\flushbottom
+\maketitle
+\thispagestyle{empty}
 
@@
 measurement from several subjects. Chris Kirtley shares normative gait from
-several studies on the CGA Normative Gait Database~\cite{Kirtley} and these
+several studies on the CGA Normative Gait Database~\cite{Kirtley2014} and these
 files have influenced and been included in other studies, for
@@
 restrictive license and sharing scheme. The University of Wisconsin at LaCrosse
-has a normative gait data set~\cite{Wilson} from 25 subjects with lower
+has a normative gait data set~\cite{Wilson2014} from 25 subjects with lower
 extremity marker data and force plate measurements from a single stance phase (TvdB: gait cycle?).
@@
   \centering
-  \small
+  \footnotesize
   \begin{tabular}{lrlll}
@@
 
-{\small\bibliographystyle{unsrt}
-\bibliography{gait-data}}
+\bibliography{gait-data}
 

### 2014-12-13 paper.tex
Proofed the abstract and introduction.

@@
   of normal walking and 480 seconds of walking while being longitudinally
-  perturbed during each stance phase with pseudo-random
-  fluctuations in the speed of the treadmill belt. A total of approximately 1.5
-  hours of normal walking ($>5000$ gait cycles) and 6 hours of perturbed walking
-  ($>20,000$ gait cycles) is included in the data set. We provide full body
-  marker trajectories and ground reaction loads in addition to a presentation of
-  processed data that includes gait events, 2D joint angles, angular rates,
-  and joint torques and the software used for the computations. The protocol is
-  described in detail along with additional elaborate meta data. This data can
-  likely be useful for validating or generating mathematical models that are
-  capable of simulating normal periodic gait and non-periodic, perturbed gaits.
+  perturbed during each stance phase with pseudo-random fluctuations in the
+  speed of the treadmill belt. A total of approximately 1.5 hours of normal
+  walking ($>5000$ gait cycles) and 6 hours of perturbed walking ($>20,000$
+  gait cycles) is included in the data set. We provide full body marker
+  trajectories and ground reaction loads in addition to a presentation of
+  processed data that includes gait events, 2D joint angles, angular rates, and
+  joint torques along with the open source software used for the computations.
+  The protocol is described in detail and supported with additional elaborate
+  meta data for each trial. This data can likely be useful for validating or
+  generating mathematical models that are capable of simulating normal periodic
+  gait and non-periodic, perturbed gaits.
 
@@
 beginning with the first motion pictures and now with modern marker based
-motion capture techniques, image and video analysis, and high fidelity ground
-reaction load measurements. Even though years of data on thousands of subjects
-now exist, this data is not widely disseminated, well organized, nor available
-with few or no restrictions. David Winter's published normative gait
-data~\cite{Winter1990} is widely used in biomechanics studies, yet it comes
-from relatively few subjects and only a small number of gait cycles per
-subject. This small source has successfully inspired powered prosthetic control
-design~\cite{Sup2008}, but success in other fields using large sets of
+motion capture techniques and high fidelity ground reaction load measurements.
+Even though years of data on thousands of subjects now exist, this data is not
+widely disseminated, well organized, nor available with few or no restrictions.
+David Winter's published normative gait data~\cite{Winter1990} is widely used
+in biomechanical studies, yet it comes from relatively few subjects and only a
+small number of gait cycles per subject. This small source has successfully
+inspired many other studies, such as powered prosthetic control
+design~\cite{Sup2008}, but success in other research fields using large sets of
 data for discovery lead one to believe that more elaborate data sets may
-benefit the field of human motion studies.
-To enable such work, data needs
-to be shared extensively, organized, and curated to enabled analysts.
+benefit the field of human motion studies. To enable such work, biomechanical
+data needs to be shared extensively, organized, and curated to enabled
+future analysts.
 
-There are some notable gait data sets and databases besides Winter's authoritative
-set that are publicly available. Notably the International Society of
+There are some notable gait data sets and databases besides Winter's
+authoritative set that are publicly available. The International Society of
 Biomechanics has maintained a web page (http://isbweb.org/data) since
-approximately 1995 that includes data sets for download and most unencumbered
-use. Vaughn, et. al's data~\cite{Vaughan1992} with kinematics and force plate
-measurement from several subjects. Chris Kirtley shares normative gait from
-several studies on the CGA Normative Gait Database~\cite{Kirtley2014} and these
-files have influenced and been included in other studies, for
-example the average gait cycles from children used in~\cite{Bogert2003}.
+approximately 1995 that includes data sets for download and mostly unencumbered
+use. For example, Vaughn, et. al's data~\cite{Vaughan1992} with kinematics and
+force plate measurement from several subjects is available on the site. At
+another website, the CGA Normative Gait Database~\cite{Kirtley2014}, Chris
+Kirtley shares normative gait data from several studies and these files have
+influenced other studies, for example the average gait cycles from children
+used in~\cite{Bogert2003}.
 
@@
 where one database contained kinematic data of 409 gait cycles of children from
-1 to 7 years old but the data does not seem to be publicly available. Tirosh
-et. al, recognized the need of a comprehensive data base for clinical gait data
-and created the Gaitabase~\cite{Tirosh2010}.  This database may contain a
-substantial amount of data but it is encumbered by a very complicated and
-restrictive license and sharing scheme. The University of Wisconsin at LaCrosse
-has a normative gait data set~\cite{Wilson2014} from 25 subjects with lower
-extremity marker data and force plate measurements from a single stance phase (TvdB: gait cycle?).
+1 to 7 years old but the data does not seem to be publicly available. This is
+unfortunately typical. But Tirosh et. al, recognized the need for a
+comprehensive data base for clinical gait data and created the
+Gaitabase~\cite{Tirosh2010}. This database may contain a substantial amount of
+data but it is encumbered by a very complicated and restrictive license and
+sharing scheme. However, there are examples of data with less restrictions. The
+University of Wisconsin at LaCrosse has an easily accessible normative gait
+data set~\cite{Wilson2014} from 25 subjects with lower extremity marker data
+from multiple gait cycles and force plate measurements from a single gait
+cycle.
 
-More recent examples of biomechanists sharing their data alongside
-publications are: \cite{Bogert2013} which includes processed joint kinematics
-from eleven subjects walking for few gait cycles and \cite{Wang2014} who
-includes data from 10 subjects walking for five minutes each at five different
-speeds but only a small set of lower extremity markers are present. The second
-is notable because it publishes the data in Dryad, a modern citable data
-repository.
+More recent examples of biomechanists sharing their data alongside publications
+are: \cite{Bogert2013} which includes full body joint kinematics and kinetics
+from eleven subjects walking for a small number of gait cycles and
+\cite{Wang2014} who includes a larger set of data from ten subjects walking for
+five minutes each at three different speeds but only a small set of lower
+extremity markers are present. The second is notable because it publishes the
+data in Dryad, a modern citable data repository.
 
 The publicly available gait data is small compared to the number of gait
-studies that have been performed over the years. And the data that is available
-in general suffers from things such as few subjects, few gait cycles, few
-markers, no raw data, limited force plate measurements, lack of meta data, and
-restrictive licensing. To help with this situation we are making the data we
-collected for our research purposes publicly available and free of the
-previously mentioned deficiencies. Not only do we provide a larger set of
-normative gait data that has been previously available we include an even large
-set of data in which the subject is being perturbed which does not exist at
-all. We believe both of these sets of data can serve a variety of use cases and
-hope that we can save time for researchers with use cases for this data.
+studies that have been performed over the years. The data that is available
+generally suffers from limitations such as few subjects, few gait cycles, few
+markers, highly clinical, no raw data, limited force plate measurements, lack
+of meta data, non-standard formats, and restrictive licensing. To help with
+this situation we are making the data we collected for our research purposes
+publicly available and free of the previously mentioned deficiencies. Not only
+do we provide a larger set of normative gait data that has been previously
+available, we also include an even larger set of data in which the subject is
+being perturbed, something that does not currently exist. We believe both of
+these sets of data can serve a variety of use cases and hope that we can save
+time and effort for future researchers by sharing it.
 
-Our use case for the data is centered around emerging powered prosthetics and
-orthotics need of bio-inspired control systems. Ideally, a powered prosthetic
-would behave in such a way that the user would feel like their limb was never
-disabled. There are a variety of approaches to developing bio-inspired control
-systems some of which aim to mimic the reactions and motion of an able-bodied
-person. A modern gait lab is able to collect a variety of kinematic, kinetic,
-and physiological data from humans during gait. This data can potentially be
-used to drive the design of the human-mimicking controller. With a rich enough
-data set, one may be able to identify control mechanisms used during a human's
-natural gait and recovery from perturbations. We have collected data that we
-believe is richer than previous gait data sets and may be rich enough for
-control identification.  The data can also be used for verification purposes
-for controllers that have been designed in other manners.
+Our use case for the data is centered around the need of bio-inspired control
+systems for emerging powered prosthetics and orthotics. Ideally, a powered
+prosthetic would behave in such a way that the user would feel like their limb
+was never disabled. There are a variety of approaches to developing
+bio-inspired control systems, some of which aim to mimic the reactions and
+motion of an able-bodied person. A modern gait lab is able to collect a variety
+of kinematic, kinetic, and physiological data from humans during gait. This
+data can potentially be used to drive the design of the human-mimicking
+controller. With a rich enough data set, one may be able to identify control
+mechanisms used during a human's natural gait and recovery from perturbations.
+We have collected data that is richer than previous gait data sets and may be
+rich enough for control identification. The data can also be used for
+verification purposes for controllers that have been designed in other manners.
 
-With all of this in mind, we collected over five and half hours of gait data
-from fifteen able bodied subjects and includes over 25,000 gait cycles. Most of
-the subjects walked at three different speeds on an instrumented treadmill
-while we collected full body marker locations and ground reaction loads from a
-pair of force plates. The protocol for the majority of the trials includes two
-minutes of normal walking and eight minutes of walking under the influence of
-pseudo-random belt speed fluctuations. The data has been organized complete with
-rich meta data and made available in the most unrestrictive form for other
+With all of this in mind, we collected over seven and half hours of gait data
+from fifteen able bodied subjects which amounts to over 25,000 gait cycles. The
+subjects walked at three different speeds on an instrumented treadmill while we
+collected full body marker locations and ground reaction loads from a pair of
+force plates. The protocol for the majority of the trials included two minutes
+of normal walking and eight minutes of walking under the influence of
+pseudo-random belt speed fluctuations. The data has been organized complete
+with rich meta data and made available in the most unrestrictive form for other
 research uses following modern best practices in data sharing~\cite{White2013}.

### 2014-12-14 paper.tex
Proofed the participants section.

@@
 %
-The study was approved by the Institutional Review Board of Cleveland State
-University (\# 29904-VAN-HS) and informed consent was obtained from all
-participants.
-
 Fifteen able bodied subjects including four females and eleven males with an
 average age of $24\pm4$ years, height of $1.75\pm0.09$ m, mass of $74\pm13$ kg
-participated in the study. The data has been anonymized with respect to the
+participated in the study. The study was approved by the Institutional Review
+Board of Cleveland State University (\# 29904-VAN-HS) and informed consent was
+obtained from all participants. The data has been anonymized with respect to the
 participants' identities and a unique identification number was assigned to
@@
     columns give the trial numbers associated with each nominal treadmill
-    speed. Additional trials with subject 0 are trials with no subject, i.e.
-    unloaded trials that can be used for inertial compensation purposes, and are
-    not shown in the table. Generated by \verb|src/subject_table.py|.}
+    speed. The measured mass is computed from the mean total vertical ground
+    reaction force just after the calibration pose event, if possible.
+    Additional trials found in the data set with a subject identification
+    number 0 are trials with no subject, i.e.  unloaded trials that can be used
+    for inertial compensation purposes, and are not shown in the table. Generated
+    by \verb|src/subject_table.py|.}
   \centering
+  \small
   \input{tables/subjects.tex}

### 2014-12-14 src/subject_table.py

@@ -41,8 +41,10 @@ for speed, trials in trials_per.items():
     speed_key = '{:1.1f} m/s'.format(speed)
     unique_subjects[speed_key] = pd.Series(trials, index[speed])
 
-cols = ['id', 'gender', 'age', 'height', 'mass', '0.8 m/s', '1.2 m/s',
-        '1.6 m/s']
+unique_subjects.rename(columns={'mass': 'self-reported mass'}, inplace=True)
+
+cols = ['id', 'gender', 'age', 'height', 'self-reported mass', '0.8 m/s',
+        '1.2 m/s', '1.6 m/s']
 units = ['', '', ' [yr]', ' [m]', ' [kg]', '', '', '']
 new_cols = [s.capitalize() + u for s, u in zip(cols, units)]
 unique_subjects.rename(columns=dict(zip(cols, new_cols)), inplace=True)
@@ -57,14 +59,17 @@ unique_subjects = unique_subjects.drop(0)  # remove null subject
 
 measured = utils.measured_subject_mass(raw_dir, processed_dir)
 
+
 def format_sigma(x):
     if x[1] >= 1.0:
         return 'dollar{:0.0f}plusminus{:0.0f}dollar'.format(*x)
     else:
         return 'dollar{:0.1f}plusminus{:0.1f}dollar'.format(*x)
 
-measured['Measured Mass [kg]'] = zip(measured['Measured Mass'], measured['Standard Deviation'])
-measured['Measured Mass [kg]'] = measured['Measured Mass [kg]'].map(format_sigma)
+measured['Measured Mass [kg]'] = zip(measured['Measured Mass'],
+                                     measured['Standard Deviation'])
+measured['Measured Mass [kg]'] = \
+    measured['Measured Mass [kg]'].map(format_sigma)
 
 unique_subjects['Measured Mass [kg]'] = measured['Measured Mass [kg]']
 new_cols.insert(4, 'Measured Mass [kg]')
@@ -79,6 +84,7 @@ tex = unique_subjects.sort().to_latex(na_rep='NA', index=False,
                                       formatters=formatters)
 tex = tex.replace('dollar', '$')
 tex = tex.replace('plusminus', '\pm')
+tex = tex.replace('rlrrlrlll', 'rlrrrrlll')
 with open(table_path, 'w') as f:
     f.write(tex)
 print('Table at: {}'.format(table_path))

### 2014-12-14 paper.tex
Proofed equipment section.

@@
 \usepackage{minted}
+
+% for nice units
+\usepackage{siunitx}
 
@@
 %
-The data were collected in the Laboratory for Human Motion and Control at Cleveland State University,
-using the following equipment:
+The data were collected in the Laboratory for Human Motion and Control at
+Cleveland State University, using the following equipment:
 
 \begin{itemize}
-  \item A ForceLink R-Mill treadmill which has dual 6 degree of freedom force
-    plates, independent belts for each foot, and motion capabilities (lateral,
-    pitch).
-  \item A 10 camera Motion Analysis motion capture system which includes Osprey
-    cameras, the Cortex software (3.1.1.1290), and hardware for collecting
-    analog and camera data simultaneously.
-  \item National Instruments USB-6255 data acquisition unit.
-  \item Four Sparkfun ADXL330 Triple Axis Accelerometer Breakout boards
-    attached to the treadmill.
-  \item Motek Medical’s D-Flow software (3.16.1 to 3.16.2) and visual display
-    system.
+  \item A R-Mill treadmill which has dual 6 degree of freedom force plates,
+    independent belts for each foot, and lateral/pitch motion capabilities
+    (Forcelink, Culemborg, Netherlands).
+  \item A 10 Osprey camera motion capture system paired with the Cortex
+    3.1.1.1290 software (Motion Analysis, Santa Rosa, CA, USA).
+  \item USB-6255 data acquisition unit (National Instruments, Austin, Texas,
+    USA).
+  \item Four ADXL330 Triple Axis Accelerometer Breakout boards attached to the
+    treadmill (Sparkfun, Niwot, Colorado, USA).
+  \item D-Flow software (versions 3.16.1 to 3.16.2) and visual display system,
+    (Motek Medical, Amsterdam, Netherlands).
 \end{itemize}
 
-The Cortex software delivers 3D marker trajectories from the cameras along with
-data from force plates and analog sensors (EMG/Accelerometer) through a
-National Instruments USB-6255 data acquisition unit. D-Flow is required to
-collect data from any digital sensors and to control the treadmill's motion
-(lateral, pitch, and belts). D-Flow can process the data in real time and/or
-export data to file.
+The Cortex software delivers high accuracy 3D marker trajectories from the
+cameras along with data from force plates and analog sensors
+(EMG/Accelerometer) through a National Instruments USB-6255 data acquisition
+unit. D-Flow is required to collect data from any digital sensors and to
+control the treadmill's motion (lateral, pitch, and belts). D-Flow can process
+the data in real time and/or export data to file.
 
@@
 points to the right, the Y coordinate points upwards, and the Z coordinate
-follows from the right-hand-rule, i.e. points backwards with respect to a
-person walking forward on the treadmill. The camera's coordinate system is
-aligned to an origin point on treadmill's surface during camera calibration.
-The same point is used as the origin of the ground reaction force measuring
-system. Figure~\ref{fig:treadmill} shows the layout of the equiptment.
+follows from the right-hand-rule, i.e. points backwards with respect to the
+walking direction. The camera's coordinate system is aligned to an origin point
+on treadmill's surface during camera calibration. The same point is used as the
+origin of the ground reaction force measuring system.
+Figure~\ref{fig:treadmill} shows the layout of the equipment.
 
+Early on, we discovered that the factory setup of the R-Link treadmill had a
+vibration mode as low as 5 Hz that is detectable in the force measurements,
+likely due to the flexible undercarriage and pitch motion mechanism. Trials 6-8
+are affected by this vibration mode. During trials 9-15 the treadmill was
+stabilized with wooden blocks. During, the remaining trials the treadmill was
+stabilized with metal supports. See the Data Limitations Section for more
+details.
+
+The acceleration of the treadmill was measured during each trial by four
+ADXL330 accelerometers placed at the four corners of the machine. These
+accelerometers were intended to provide information for inertial compensation
+purposes when the treadmill moved laterally, but are extraneous for trials
+greater than 8 due to the treadmill being stabilized.
+%
 \begin{figure*}
@@
   \includegraphics{figures/treadmill.png}
-  \caption{The treadmill, cameras, projection screen, and saftey rope. The
-    direction of travel is in the $-z$ direction.}
+  \caption{The treadmill with coordinate system, cameras (circled in orange),
+    projection screen, and safety rope. The direction of travel is in the $-z$
+    direction.}
   \label{fig:treadmill}
@@
     skipped.
+  \item  Trials 6-15 force measurements are affected by this and should be not
+    be used. We include the trials because the kinematic data is valid and
+    trials 6-8 include lateral perturbations in addition to the longitudinal.
   \item Trials 9, 10, and 11 used a slightly different event definition where the

### 2014-12-14 src/subject_table.py
Fixed subject table formatting.

@@ -51,7 +51,7 @@ unique_subjects.rename(columns=dict(zip(cols, new_cols)), inplace=True)
 
 formatters = {'Height [m]': lambda x: 'NA' if np.isnan(x)
               else '{:0.2f}'.format(x),
-              'Mass [kg]': lambda x: 'NA' if np.isnan(x)
+              'Self-reported mass [kg]': lambda x: 'NA' if np.isnan(x)
               else '{:0.0f}'.format(x)}
 
 unique_subjects = unique_subjects.drop_duplicates()

### 2014-12-14 paper.tex
Proofed protocol section.

@@
 %
-The experimental protocol consisted of static measurements and walking on the
-treadmill for 10 minutes under unperturbed and perturbed conditions. Before a
-set of trials on the same day the following happened:
+The experimental protocol consisted of both static measurements and walking on
+the treadmill for 10 minutes under unperturbed and perturbed conditions. Before
+a set of trials on the same day the following happened:
 %
 \begin{itemize}
-  \item Calibration of the motion capture system.
+  \item Calibration of the motion capture system using the manufacturer's
+    recommended procedure.
   \item Subject changes into athletic shoes, shorts, sports bra, baseball
@@
   \item All 47 markers are applied directly to the skin except for the heel,
-    toe, and head markers.~\footnote{The sacrum and rear pelvic markers were
-    placed on the shorts for a small number of the subjects}.
+    toe, and head markers, which were placed on the respective article of
+    clothing.~\footnote{The sacrum and rear pelvic markers may have been placed
+    on the shorts for a small number of the subjects}.
   \item Subjects self-reported age, gender, and mass.
-  \item Height is measured by the experimentalist.
-  \item Four reference photographs (front, back, right, left) are taken of
-    subjects marker locations.
+  \item Height was measured by the experimentalist.
+  \item Four reference photographs (front, back, right, left) were taken of
+    subject's marker locations.
 \end{itemize}
 
-After obtaining informed consent and a briefing by the experimentalist on the trial protocol, the subject
-follows the verbal instructions of the experimentalist and the on-screen
-instructions from the video display in front of the subject. The protocol goes
-as follows:
+After obtaining informed consent and a briefing by the experimentalist on the
+trial protocol, the subject followed the verbal instructions of the
+experimentalist and the on-screen instructions from the video display. The
+protocol for a single trial was as follows:
 %
 \begin{enumerate}
-  \item Subject steps onto the treadmill and markers are identified with
+  \item Subject stepped onto the treadmill and markers were identified with
     Cortex.
-  \item The safety cable is attached loosely to the rock climbing harness such
-    that no undue forces are acting on the subject during walking, but that the
-    harness will prevent a full fall.
-  \item The subject starts by stepping on sides of treadmill so that feet don't
-    touch the force plates and the force plate signals are zeroed.
-  \item Once notified by the screen, the subject stands in the initialization
-    pose: standing straight up, looking forward, arms out by their sides (~45
-    degrees) and the event is manually recorded by the operator.
-  \item A countdown to the first normal walking phase is displayed. At the end
-    of the countdown the treadmill starts at the specified speed and the
-    subject is instructed to walk normally, to focus on the road in the
-    screen, and not to look at their feet.
+  \item The safety rope was attached loosely to the rock climbing harness such
+    that no undue forces were acting on the subject during walking, but that
+    the harness would prevent a full fall.
+  \item The subject started by stepping on sides of treadmill so that feet did
+    not touch the force plates and the force plate signals are zeroed. This
+    corresponds to the ``Force Plate Zeroing'' event.
+  \item Once notified by the video display, the subject stood in the
+    initialization pose: standing straight up, looking forward, arms out by
+    their sides (~45 degrees) and the event, ``Calibration Pose'', was manually
+    recorded by the operator.
+  \item A countdown to the first normal walking phase was displayed. At the end
+    of the countdown the event ``First Normal Walking'' was recorded and the
+    treadmill ramped up to the specified speed and the subject was instructed
+    to walk normally, to focus on the ``endless'' road on the display, and not
+    to look at their feet.
   \item After 1 minute of normal walking, the longitudinal perturbation phase
-    begins.
+    begun and was recorded as ``Longitudinal Perturbation''.
   \item After 8 minutes of walking under the influence of the perturbations,
-    the normal walking phase begins.
-  \item After 1 minute, a countdown is shown and the treadmill decelerates to a
-    stop.
-  \item The subject is instructed to step off of the force plates for 10
-    seconds.
-  \item The subject could take a break before each additional trial.
+    the second normal walking phase begun and was recorded as ``Second Normal
+    Walking''.
+  \item After 1 minute of normal walking, a countdown was shown on the display
+    and the treadmill decelerated to a stop.
+  \item The subject was instructed to step off of the force plates for 10
+    seconds and the ``Unloaded End'' event was recorded.
+  \item The subject could then take a rest break before each additional trial.
 \end{enumerate}
+
+Trials 6-8 included a calibration pose at the start of the trial but the event
+was not explicitly recorded. In those trials, the ``TreadmillPerturbation''
+event marks the beginning of longitudinal perturbations and the ``Both'' event
+marks the beginning of combined longitudinal and lateral perturbations. The
+force plate zeroing at the end was also not explicitly recorded.
 

### 2014-12-14 paper.tex
Proofed the perturbation signals section.

@@
 %
-The data set uses a testing protocol containing a segment of normal walking,
-followed by a series of pseudo-random longitudinal perturbations, and ending
-with a second segment of normal walking. Three pseudo-random speed control
-signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
-generated through MATLAB and Simulink. Initially, a random acceleration signal
-was created using discrete-time Gaussian white noise and saturated at the
-maximum belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt
-speed, and high-pass filtered (second-order Butterworth) to eliminate velocity
-drift. Mean velocity was then added to the signal and limited between 0 and 3.6
-m/s. The cutoff frequencies of the high-pass filter, as well as the variance in
-the acceleration signal, were adjusted until the appropriate standard
-deviations for each mean velocity were obtained.
+As previously described, the protocol included a phase of normal walking,
+followed by longitudinal belt speed perturbations, and ended with a second
+segment of normal walking. Three pseudo-random belt speed control signals, with
+mean velocities of \SIlist{0.8;1.2;1.6}{\meter\per\second}, were pre-generated
+with MATLAB and Simulink (Mathworks, Natick, Massachusetts, USA).  The same
+control signal was used for all trials at that given speed.
 
-These standard deviations were determined through a trial and error method.
-Initially, perturbations of $\pm$ 10\% for each mean velocity were generated.
-However, this large deviation in the slowest mean velocity (0.8 $m \cdot
-s^{-1}$) would unintentionally reduce the treadmill speed to values close to
-zero. This would produce an undesired stopping and starting sensation.
-Alternatively, smaller standard perturbations of $\pm$ 5\% did not produce
-noticeable perturbations in the fastest trial (1.6 $m \cdot s^{-1}$).
-Therefore, standard deviations of 7.16\%, 10.23\%, and 13.1\% were selected for
-the 0.8, 1.2, and 1.6 $m \cdot s^{-1}$ speeds, respectively.  This ensured that
-the test subject was sufficiently perturbed for each speed, while remaining
-within the limits of our testing protocol.
+To create the signals, we started by generating random 100~\si{\hertz}
+acceleration signals using the Simulink discrete-time Gaussian white noise
+block followed by a saturation block set at the maximum belt acceleration of
+15~\si{\meter\per\second\squared}. The signal was then integrated to obtain
+belt speed and high-pass filtered with a second-order Butterworth filter to
+eliminate drift. One of the three mean speeds were then added to the signal and
+limited between \SIrange{0}{3.6}{\meter\per\second}. The cutoff frequencies of
+the high-pass filter, as well as the variance in the acceleration signal, were
+manually adjusted until acceptable standard deviations for each mean speed were
+obtained: \SIlist{0.06;0.12;0.21}{\meter\per\second} for the three speeds,
+respectively. These ensured that the test subjects were sufficiently perturbed
+at each speed, while remaining within the limits of our equipment and testing
+protocol.
 
-To ensure that the treadmill could follow the higher frequencies of the random
-signals, the maximum treadmill acceleration of 15 $m \cdot s^{-2}$ was used in
-the D-Flow Software. As this value remains constant throughout the experiment,
-the initial starting and stopping of the treadmill would be too rapid for test
-subjects. To eliminate this, a ramped acceleration and deceleration was
-generated and placed at the beginning and end of the entire signal,
-respectively. Other additions include the one minute segments of unperturbed
-walking at the beginning and end of the trial and 30 seconds of no movement for
-force plate zeroing and subject T-pose calibration.  The sampling rate of the
-simulation was set to 100 Hz, corresponding to the D-Flow software's
-capabilities.
+To ensure that the treadmill belts could accelerate to the desired values, the
+high performance mode in the D-Flow software was enabled. This had the side
+effect of enabling too rapid of accelerations when the belt speed changed to or
+from zero speed. To eliminate this, a suitable ramped acceleration and
+deceleration were generated for the speed transitions.
 
-The resulting output from the MATLAB script and Simulink produces a comma-
-delimited text file of six columns: time stamp, slow, normal, and fast walking
-perturbation signals, and slow and fast running signals.  These running signals
-were generated in the event the testing protocol expanded from a walking study
-to a running study, though they were not used in the data set.  The actual
-motion of the treadmill belts are compared to the original input signals as
-shown in Figure~\ref{fig:input_output}.   The standard deviations of the
-outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the 0.8, 1.2,
-and 1.6 $m \cdot s^{-1}$ speeds, respectively.
-
-\begin{figure*}
+The MATLAB script and Simulink model produce a comma-delimited text file of six
+signals: time stamp, slow, normal, and fast walking perturbation signals, and
+slow and fast running signals.~\footnote{The running signals were not used in
+the experiments presented in this paper.} The measured speed of the treadmill
+belts are compared to the control input signals in
+Figure~\ref{fig:input_output} to show the effect of the treadmill and
+controller dynamics. The system introduces a delay and seems to act as a low
+pass filter. The standard deviations of the outputs do not significantly differ
+from the desired values: \SIlist{0.05;0.12;0.2}{\meter\per\second} for the
+three speeds, respectively.
+%
+\begin{figure}
   \centering
   \includegraphics{figures/input_vs_output.pdf}
-  \caption{Treadmill belt velocity input signals (purple) and recorded output
-    velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
-    respectively}
+  \caption{Treadmill belt speed input signals (purple) and recorded output
+    speeds (blue) for average belt speeds of
+    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively.}
   \label{fig:input_output}
-\end{figure*}
+\end{figure}
 
-For further analysis, the input and output signals for each average speed were
-transformed into the frequency domain through the Fast Fourier algorithm, and
-the results are shown in Figure~\ref{fig:freq_analysis}. It was observed that
-for the 1.2 $m \cdot s^{-1}$ average walking speed, the amplitude of the output
-is significantly lower than the amplitude of the input signal at lower
-frequencies. Additionally, the amplitude of output signal in the 0.8 $m \cdot
-s^{-1}$ walking speed begins to attenuate around 2 Hz, which is a noticeably
-lower frequency than the other walking speeds.
-
-\begin{figure*}
+To show the effects of the treadmill dynamics and give an idea of the frequency
+content of the actual perturbations, the input and output for each speed were
+transformed into the frequency domain using the Fast Fourier algorithm, and the
+results are shown in Figure~\ref{fig:freq_analysis}. This shows that for the
+1.2~\si{\meter\per\second} walking speed, the amplitude of the output is
+significantly lower than the amplitude of the input signal at lower
+frequencies. Additionally, the amplitude of output signal in the
+0.8~\si{\meter\per\second}  walking speed begins to attenuate around
+2~\si{\hertz}, which is a noticeably lower frequency than the other walking
+speeds.
+%
+\begin{figure}
   \centering
@@
   \label{fig:freq_analysis}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{listing*}
+\begin{listing}
   \begin{minted}[gobble=4]{python}
@@
   \label{lis:dflowdata}
-\end{listing*}
+\end{listing}
 

### 2014-12-14 paper.tex
Proofed raw data section.

@@
 The raw data consists of a set of ASCII tab delimited text files output from
-both the ``mocap'' and ``record'' modules in D-Flow in addition to a YAML file
-that contains all of the necessary meta data for the given trial. These three
-files are stored in a hierarchy of directories with one trial per directory.
-The directories are named in the following fashion \verb+T001/+ where \verb+T+
-stands for ``trial'' and the following three digits are provide a unique trial
-identification number.
-
-The text file output from the mocap module in D-Flow is a tab delimited file.
-The first line is the header and contains a time stamp column, frame number
-column, marker position columns, force plate force/moment columns, force plate
-center of pressure columns, other analog columns, and potentially results from
-the real time Human Body Model \cite{Bogert2013} which is included with the
-D-Flow software. These are detailed below. The numerical values of the
-measurements are provided in decimal floating point notation with 6 decimals of
-precision, e.g. 123456.123456 [\%1.6f].
+both the ``mocap'' and ``record'' modules in D-Flow in addition to a manually
+generated YAML file that contains all of the necessary meta data for the given
+trial. These three files are stored in a hierarchy of directories with one
+trial per directory. The directories are named in the following fashion
+\verb+T001/+ where \verb+T+ stands for ``trial'' and the following three digits
+are provide a unique trial identification number.
 
@@
 \verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id number. The file
-contains several columns:
-
+is tab delimited and contains a number of time series. The numerical values of
+the time series are provided in decimal fixed point notation with 6 decimals of
+precision, e.g. \verb|123456.123456|, regardless of the units. The first line
+of the file holds the header. The header includes time stamp column, frame
+number column, marker position columns, force plate force/moment columns, force
+plate center of pressure columns, other analog columns, and potentially results
+from the real time Human Body Model \cite{Bogert2013} which is included with
+D-Flow. The columns are further described below:
+%
 \begin{description}
   \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
-    receives a frame from Cortex. These are approximately at 100 hz and given
-    in seconds.
+    receives a frame from Cortex. These are recorded at approximately at
+    100~\si{\hertz} and given in seconds.
   \item[FrameNumber] Monotonically increasing positive integers that correspond
@@
     or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
-    reference frame. These values are in meters.
+    reference frame. The prefixes match the marker labels given in
+    Table~\ref{tab:marker-labels}. These values are in meters.
   \item[Ground Reaction Loads] There are three ground reaction forces and three
@@
     much higher frequency than the cameras, but delivered at the Cortex camera
-    sample rate, 100 Hz through the D-Flow mocap module. A force/moment
-    calibration matrix stored in Cortex converts the voltages to forces and
-    moments before sending it to D-Flow [1]. Cortex also computes the center of
-    pressure from the forces, moments, and force plate dimensions. These have
-    the same prefixes for the plate number, have the suffix \verb+.Cop[XYZ]+,
-    and are in meters.
+    sample rate, \~100~\si{\hertz} through the D-Flow mocap module. A
+    force/moment calibration matrix stored in Cortex converts the voltages to
+    forces and moments before sending it to D-Flow. Cortex also computes
+    the center of pressure from the forces, moments, and force plate
+    dimensions. These have the same prefixes for the plate number, have the
+    suffix \verb+.Cop[XYZ]+, and are given in meters.
   \item[Analog Channels] Several analog signals are recorded under column
     headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
-    sampled by Cortex and correspond to the 96 analog-channel National
-    Instruments 6255 DAQ box. The first twelve are the voltages from the
+    sampled by Cortex and correspond to the 96 analog channels in the National
+    Instruments USB-6255. The first twelve are the voltages from the
     force plate load cells. We also record the acceleration of 4 points on the
-    treadmill base in channels 61-72.  TvdB: are these useful?  If so, we need to
-    provide information on how to process these channels. If not, maybe
-    not mention them at all?
+    treadmill base in analog channels 61-72 that were in place in case inertial
+    compensation for the lateral treadmill movement was required.
 \end{description}
@@
 %
-The record module also outputs a tab delimited ASCII text file. It includes a
-\verb+Time+ column which records the D-Flow system time in seconds which
-corresponds to the time recorded in the \verb+TimeStamp+ column in mocap module
-tsv file. There are two additional columns \verb+RightBeltSpeed+ and
-\verb+LeftBeltSpeed+ which provide the independent belt speeds measured in
-meters per second by an encoder in the treadmill.
+The record module also outputs a tab delimited ASCII text file with numerical
+values at six decimal digits. It includes a \verb+Time+ column which records
+the D-Flow system time in seconds. This time corresponds to the time recorded
+in the \verb+TimeStamp+ column in mocap module tsv file which is necessary for
+time synchronization. There are two additional columns \verb+RightBeltSpeed+
+and \verb+LeftBeltSpeed+ which provide the independent belt speeds measured in
+meters per second by a factory installed encoder in the treadmill.
 
-Finally, the record module is capable of recording the time at which various
-events occur, as detected or set by D-Flow. It does this by inserting commented
-(\#) lines in between the rows when the event occurred. The record files have
-several events that correspond to the phases of the protocol:
-
+Additionally, the record module is capable of recording the time at which
+various preprogrammed events occur, as detected or set by D-Flow. It does this
+by inserting commented (\verb|#|) lines in between the rows when the event
+occurred. The record files have several events that delineate the different
+phases of the protocol:
+%
 \begin{description}
-  \item[A: Force Plate Zeroing] Marks the time at which there is no load on the
-    force plates and when the voltages were zeroed.
+  \item[A: Force Plate Zeroing] Marks the time at the beginning of the trial at
+    which there is no load on the force plates and when the force plate
+    voltages were zeroed.
   \item[B: Calibration Pose] Marks the time at which the person is in the
-    T-Pose.
+    calibration pose.
   \item[C: First Normal Walking] Marks the time when the treadmill begins Phase
@@
 all meta data files in the study. Details are given below:
-
+%
 \begin{description}
@@
 
-The \verb+subject+ provides key value pairs of information about the subject in
-that trial. Each subject has a unique identification number along with basic
-anthropomorphic data. The following details the possible meta data for the
-subject:
-
+The \verb+subject+ section provides key value pairs of information about the
+subject in that trial. Each subject has a unique identification number along
+with basic anthropomorphic data. The following details the possible meta data
+for the subject:
+%
 \begin{description}
@@
   \item[height] A float specifying the measured height of the subject (with
-    shoes on) at the time of the trial.
+    shoes and hat on) at the time of the trial.
   \item[height-units] A string giving the units of the height measurement.
@@
 information, detailed below:
-
+%
 \begin{description}
@@
   \item[cortex-version] The version of Cortex used to record the trial.
-  \item[datetime]A date formatted string giving the date of the trial in the
+  \item[datetime] A date formatted string giving the date of the trial in the
     \verb|YYYY-MM-DD| format.
@@
     where the key is the D-Flow file type and the value is the path to the file
-    relative to the meta file.
+    relative to the meta file. The compensation file corresponds to an unloaded
+    trial collected on the same day that could be used for inertial
+    compensation purposes, if needed.
   \item[hardware-settings] There are tons of settings for the hardware in both
@@
     during the trial.
-  \item[nominal-speed-units] (string) The units of the nominal speed.
+  \item[nominal-speed-units] A string providing the units of the nominal speed.
   \item[notes] Any notes about the trial.
@@
       sway or pitch motion was actuated during the trial. If this flag is
-      false, the measured ground reaction forces must be compensated for the
+      false, the measured ground reaction loads must be compensated for the
       inertial affects and be expressed in the motion capture reference frame.
   \item[subject-id] An integer corresponding to the subject in the trial.
-  \item[sway] A boolean that indicates if the treadmill lateral (sway) degree
-    of freedom was actuated during the trial.
+  \item[sway] A boolean that indicates if the treadmill lateral degree of
+    freedom was actuated during the trial.
 \end{description}
 
-\begin{listing*}
+\begin{listing}
   \small
@@
   \label{lis:example-meta-data}
-\end{listing*}
+\end{listing}
 

### 2014-12-14 paper.tex
Proofed marker section.

@@
 %
-We make use of the full body 47 marker set described in \cite{Bogert2013}
-and presented in detail in Table \ref{tab:marker-labels}. As with all camera
-based motion capture systems, the markers sometimes go missing in the
-recording.
-
-If the data was recorded in a D-Flow version less than 3.16.2rc4 [3], D-Flow
-records the last non-missing value in all three axes until the marker is
-visible again when a marker goes missing.
-
-In D-Flow versions greater than or equal to 3.16.2rc4 the missing markers are
-indicated in the TSV file as either 0.000000 or -0.000000, which is the same as
-has been in the HBM columns in all versions of D-Flow. The D-Flow version must
-be provided in the meta data yml file, otherwise it will assume D-Flow is at
-the latest version.
-
+We make use of the full body 47 marker set described in \cite{Bogert2013} and
+presented in detail in Table \ref{tab:marker-labels}. As with all camera based
+motion capture systems, the markers sometimes go missing in the recording. When
+a marker goes missing, if the data was recorded in a D-Flow version less than
+3.16.2rc4 [3], D-Flow continues to record the last non-missing value in all
+three axes until the marker is visible again. In D-Flow versions greater than
+or equal to 3.16.2rc4, the missing markers are indicated in the TSV file as
+either \verb|0.000000| or \verb|-0.000000|, which is the same as has been in
+the HBM columns in all versions of D-Flow. The D-Flow version must be provided
+in the meta data yml file for each trial to be able to distinguish this detail.
+%
 \begin{table*}
-  \cprotect\caption{Descriptions of the 47 markers used in this study. The Set
-    column indicates whether the marker exists in the lower and/or full body
-    marker set.  The label column matches the column headers in the
+  \cprotect\caption{Descriptions of the 47 markers used in this study. The
+    ``Set'' column indicates whether the marker exists in the lower and/or full
+    body marker set. The label column matches the column headers in the
     \verb|mocap-xxx.txt| files and/or the marker map in the \verb|meta-xxx.yml|

### 2014-12-14 paper.tex
Proofed processed data section.

@@
 %
-\begin{figure*}
+\begin{figure}
   \centering
@@
   \label{fig:treadmill}
-\end{figure*}
+\end{figure}
 
@@
 \section*{Processed Data}
-
-We developed a toolkit for data processing, GaitAnalysisToolKit, for common
-gait computations and provide an example processed trial to show the character
-of the data. The tool was developed in Python, is dependent on the SciPy Stack,
-and provides two main classes: one to do basic gait data cleaning,
-\verb|DFlowData|, and a second to compute common gait variables of interest,
-\verb|GaitData|.
+%
+We developed a toolkit for data processing,
+GaitAnalysisToolKit~\cite{GATK2014}, for common gait computations and provide
+an example processed trial to present the nature of the data. The tool was
+developed in Python, is dependent on the SciPy Stack and Octave, and provides
+two main classes: one to do basic gait data cleaning from D-Flow's output
+files, \verb|DFlowData|, and a second to compute common gait variables of
+interest, \verb|GaitData|.
 
@@
 the previous section and applies several ``cleaning'' operations to transform
-the data into a usable form. The cleaning process follows these basic steps:
-
+the data into a usable form. The cleaning process follows these steps:
+%
 \begin{enumerate}
-  \item Loads the meta data file into a Python dictionary.
-  \item Loads the D-Flow mocap module TSV file into Pandas \verb|DataFrame|.
-  \item Relabels the column headers to more meaningful names if this is
+  \item Load the meta data file into a Python dictionary.
+  \item Load the D-Flow mocap module TSV file into Pandas \verb|DataFrame|.
+  \item Relabel the column headers to more meaningful names if this is
     specified in the meta data.
-  \item Optionally identifies the missing values in the mocap marker data and
-    replaces them with \verb|numpy.nan|.
-  \item Optionally interpolates the missing marker values and replaces them
+  \item Optionally identify the missing values in the mocap marker data and
+    replace them with \verb|numpy.nan|.
+  \item Optionally interpolate the missing marker values and replaces them
     with interpolated estimates using a variety of interpolation methods.
-  \item Loads the D-Flow record module TSV file into a Pandas \verb|DataFrame|.
-  \item Extracts the events and creates a dictionary mapping the event names in
+  \item Load the D-Flow record module TSV file into a Pandas \verb|DataFrame|.
+  \item Extract the events and create a dictionary mapping the event names in
     the meta data to the events detected in the record module file.
-  \item Compensates the ground reaction loads based on whether the meta data
-    indicates there was treadmill motion.
-  \item Merges the data from the mocap module and record module into one data
+  \item Interially compensate the ground reaction loads based on whether the
+    meta data indicates there was treadmill motion.
+  \item Merge the data from the mocap module and record module into one data
     frame at the maximum common constant sample rate.
@@
 Once the data is cleaned there are two methods that allow you to extract the
-cleaned data:
-
-\begin{itemize}
-  \item Extract sections of the data bounded by the events recorded in the
-    \verb|record-XXX.txt| file.
-  \item Save the cleaned data to disk.
-\end{itemize}
-
-An example of the API is given in Listing~\ref{lis:dflowdata}.
-
+cleaned data: either extract sections of the data bounded by the events
+recorded in the \verb|record-xxx.txt| file or save the cleaned data to disk.
+These operations are available as a command line application and as an
+application programming interface (API) in Python. An example of the
+\verb|DFlowData| API in use is provided in Listing~\ref{lis:dflowdata}.
+%
 \begin{listing}
@@
     >>> data.clean_data()
-    >>> event_df = dflow_data.extract_processed_data(event='Longitudinal Perturbation')
+    >>> event_df = dflow_data.extract_processed_data(
+    ...     event='Longitudinal Perturbation')
   \end{minted}
@@
 
-The \verb|GaitData| class is then used to compute things such as gait landmarks
-(toe off and heel strike times), basic 2D kinematics and inverse dynamics, and to store the
-data into a Pandas \verb|Panel| with each gait cycle on the main axis at a
-specified sample rate. This data is also saved to disk in HDF5 format.
-
-\begin{listing*}
+The \verb|GaitData| class is then used to compute things such as gait events
+(toe off and heel strike times), basic 2D kinematics and inverse dynamics, and
+to store the data into a Pandas \verb|Panel| with each gait cycle on the item
+axis at a specified sample rate. This object can also be serialized to disk in
+HDF5 format. An example of using the Python API is shown in
+Listing~\ref{lis:gaitdata}.
+%
+\begin{listing}
   \begin{minted}[gobble=4]{python}
@@
   \end{minted}
-  \cprotect\caption{Python interpreter session showing how one could compute
-    the inverse dynamics (joint angles and torques), identify the gait
-    landmarks (e.g. heel strikes), split the data with respect to the gait
-    landmarks in a Pandas \verb|Panel|, plot the mean and standard deviation of
-    one time history with respect to the gait cycles, and save the data to disk.}
-  \label{lis:dflowdata}
-\end{listing*}
+  \cprotect\caption{Python interpreter session showing how one could use the
+    \verb|GaitData| class to load in the result of \verb|DFlowData| and compute
+    the inverse dynamics (joint angles and torques), identify the gait event
+    (e.g. heel strikes), split the data with respect to the gait events in a
+    Pandas \verb|Panel|, plot the mean and standard deviation of one time
+    history with respect to the gait cycles, and save the data to disk.}
+  \label{lis:gaitdata}
+\end{listing}
 
@@
 and torques from the perturbed gait cycles and the unperturbed gait cycles
-computed from trial 20 for 120 gait cycles. The variation in the dynamics is
-larger for the perturbed gait cycles.
-
-\begin{figure*}
+computed from trial 20. This gives an idea of the more highly variable dynamics
+required to walk while being longitudinally perturbed.
+%
+\begin{figure}
   \centering
@@
     joint angles and torques from both unperturbed (purple) and perturbed
-    (blue) gait cycles from trial 20. Produced by \verb|unperturbed_perturbed_comparison.py|.}
+    (blue) gait cycles from trial 20. Produced by
+    \verb|src/unperturbed_perturbed_comparison.py|.}
   \label{fig:angle-torque-comparison}
-\end{figure*}
+\end{figure}
 
-Figure~\ref{fig:gait-cycle-stats-comparison} compares the distribution in a few
-statistics of the unperturbed and perturbed gait cycles.
-
-\begin{figure*}
+For more insight into the difference in the unperturbed and perturbed data,
+Figure~\ref{fig:gait-cycle-stats-comparison} compares the distribution of a few
+gait cycle statistics. One can see that the perturbed strides have a much
+larger variation in frequency and length.
+%
+\begin{figure}
   \centering
@@
   \cprotect\caption{Box plots of the average belt speed, stride frequency, and
-    stride length which compare 120 gait cycles for the unperturbed (purple) and
-    perturbed (blue). Produced by \verb|unperturbed_perturbed_comparison.py|.}
+    stride length which compare gait cycles for the unperturbed (purple) and
+    perturbed (blue). Produced by \verb|src/unperturbed_perturbed_comparison.py|.}
   \label{fig:gait-cycle-stats-comparison}
-\end{figure*}
+\end{figure}
 

### 2014-12-14 paper.tex
Proofed data limitations section.

@@
 %
-The data is provided in good faith with great attention to detail. The
-following list gives various notes and warnings about the data that should be
-taken into account when making use of it.
+The data is provided in good faith with great attention to detail but as with
+all data there are anomalies that may affect the use and interpretation of
+results emanating from the data. The following list gives various notes and
+warnings about the data that should be taken into account when making use of
+it.
 %
@@
   \item Be sure to read the notes in each meta data file for details about
-    possible anomalies in that particular trial.
+    possible anomalies in that particular trial. Things such as marker dropout,
+    ghost markers, and marker movement are the more prominent notes. Details
+    about variations in the equipment on the day of the trial are also
+    mentioned.
   \item The subject identification number 0 stands for "no subject" and was
@@
     treadmill, for example during the trials that were intended to be used for
-    inertial compensation purposes if the treadmill was moving. These trials
-    play through the exact protocol as those with a human subject and the
-    matching trials are indicated in the meta data. (TvdB: can the user tell
-    which compensation file to use with which subject files?  By the date on
-    the file?)
+    inertial compensation purposes. These trials play through the exact
+    protocol as those with a human subject and the matching trials are
+    indicated in the meta data. Matching unloaded trials were recorded on the
+    same day as the loaded trials and is noted in the
+    \verb|trial:files:compensation| section of the meta data file.
   \item Trials 1 and 2 were not recorded as part of this study. Those trial
@@
     skipped.
-  \item  Trials 6-15 force measurements are affected by this and should be not
-    be used. We include the trials because the kinematic data is valid and
-    trials 6-8 include lateral perturbations in addition to the longitudinal.
   \item Trials 9, 10, and 11 used a slightly different event definition where the
     calibration poses were not explicitly tagged by an event, yet the protocol
-    was the identical to the following trials.
+    was the identical to the following trials. The calibration pose will have
+    to be determined manually.
+  \item  Trials 6-15 have force measurements are affected by the treadmill
+    vibration mode mentioned in the equipment section and the forces should be
+    not be used. We include the trials because both the kinematic data is valid
+    and trials 6-8 include lateral perturbations in addition to the
+    longitudinal.
   \item During trials 9-15 we used wooden blocks to fix the treadmill to the
-    concrete floor to eliminate the treadmill's lowest factory vibration mode
+    concrete floor to eliminate the treadmill's low vibration mode
     (~5 hz). But these blocks seem to have corrupted the force plate
@@
     fine.
-  \item We did not record unloaded compensation trials for trials 9-15.
-    Regardless, they would be useless due to the previous point.
   \item Trials 6-8 use an early experimental protocol which divided the
@@
     treadmill had a low vibrational mode which significantly affects the force
-    plate measurements. The force measurements during these trials are
-    corrupted by this vibrational mode and should be used with caution or not
-    at all.
-  \item Trials 6-8 use a XX lower body marker set only.
+    plate measurements, requiring us to eliminate the lateral perturbation
+    motions. The force measurements during these trials are corrupted by this
+    vibrational mode and should be used with caution or not at all.
+  \item We did not record unloaded compensation trials for trials 9-15.
+    Regardless, they would likely be useless due to the corruption from the
+    wooden blocks.
+  \item Trials 6-8 use a only the lower body marker set. The remaining trials
+    are full body.
   \item The ankle joint torques computed from subject 9's data in trials 25-27
     are abnormal and should be used with caution or not at all. We were not
-    about to locate the source of the error, but it is likely related to the
+    able to locate the source of the error, but it is likely related to the
     force calibration.

### 2014-12-14 paper.tex
Fixed some units.

@@
 Early on, we discovered that the factory setup of the R-Link treadmill had a
-vibration mode as low as 5 Hz that is detectable in the force measurements,
+vibration mode as low as 5\si{\hertz} that is detectable in the force measurements,
 likely due to the flexible undercarriage and pitch motion mechanism. Trials 6-8
@@
     concrete floor to eliminate the treadmill's low vibration mode
-    (~5 hz). But these blocks seem to have corrupted the force plate
+    (\~5\si{\hertz}). But these blocks seem to have corrupted the force plate
     measurements by imposing frictional stresses on the system. The force plate

### 2014-12-14 paper.tex
Proofed conclusion.

@@
 A similar work flow was used to produce Figure~\ref{fig:angle-torque-comparison}
-which compares the mean and standard deviation of saggital plane joint angles
+which compares the mean and standard deviation of sagittal plane joint angles
 and torques from the perturbed gait cycles and the unperturbed gait cycles
@@
 \section*{Conclusion}
-
-We have presented the details of what we believe to be a rich dataset of motion
-and ground reaction loads humans exhibit when recovering from minor
-longitudinal perturbations. The data provides a large sample of gait cycles
-from both unperturbed and perturbed walking at three different nominal speeds.
-The raw data is provided for reuse with sufficient detail with regards to meta
-data and technical details. In addition to the data, we provide software that
-can process the data for cleaning purposes and to produce typical saggital
-plane gait variables of interest. Among other uses, we believe the dataset is
-ideally suited for control identification purposes. Many researchers are
-working on mathematical models for control in gait and this dataset provides
-both a way to validate these models and data that can be used to generate
-models.
+%
+We have presented the details of what we believe to be a rich and elaborate
+data set of motion and ground reaction loads humans exhibit during both normal
+walking and when recovering from longitudinal perturbations. The data provides
+a very large sample of gait cycles from both unperturbed and perturbed walking
+at three different nominal speeds. The raw data is provided for reuse with
+complete meta data. In addition to the data, we provide software that can
+process the data for both cleaning purposes and to produce typical sagittal
+plane gait variables of interest. The variation of these gait variables and
+statistics are provided to give a high level picture of the character of the
+data. Among other uses, we believe the dataset is ideally suited for control
+identification purposes. Many researchers are working on mathematical models
+for control in gait and this dataset provides both a way to validate these
+models and a source for generating them.
 

### 2014-12-15 paper.tex
Added data and software availability sections, closes #70.

@@
 %
-We developed a toolkit for data processing,
-GaitAnalysisToolKit~\cite{GATK2014}, for common gait computations and provide
-an example processed trial to present the nature of the data. The tool was
-developed in Python, is dependent on the SciPy Stack and Octave, and provides
-two main classes: one to do basic gait data cleaning from D-Flow's output
-files, \verb|DFlowData|, and a second to compute common gait variables of
-interest, \verb|GaitData|.
+We developed a toolkit for data processing, GaitAnalysisToolKit v0.1.2,
+\cite{Moore2014a}, for common gait computations and provide an example
+processed trial to present the nature of the data. The tool was developed in
+Python, is dependent on the SciPy Stack and Octave, and provides two main
+classes: one to do basic gait data cleaning from D-Flow's output files,
+\verb|DFlowData|, and a second to compute common gait variables of interest,
+\verb|GaitData|.
 
@@
 
-\subsection*{Author contributions}
+\section*{Data Availability}
+The data set, \cite{Moore2014}, is available via the Zenodo data repository.
+Two approximately 1.2GB gzipped tar balls contain the data and a \verb|README|
+file with a short description of the contents. The data is released under the
+Creative Commons CC0 license (http://creativecommons.org/about/cc0) following
+best practices for sharing scientific data.
+
+\section*{Software Availability}
+The tables and figures in the paper can be reproduced from the source
+repository shared on Github: https://github.com/csu-hmc/perturbed-data-paper.
+Along with the source code in the repository, the computations depend on
+version 0.1.2 of the GaitAnalysisToolKit, \cite{Moore2014a}, which can be
+downloaded from Zenodo or the Python Package Index (http://pypi.python.org).
+
+\section*{Author contributions}
 A.v.d.B. conceived of the experiments and protocol. J.K.M and S.K.H refined the
-protocol, ran the experiments, and collected the data. J.K.M was the primary
-author of the paper with significant contributions from S.K.H and A.v.d.B. All
-authors were involved in the revision of the draft manuscript and have agreed
-to the final content.
+protocol, ran the experiments, collected the data, developed the software, and
+analyzed the data. J.K.M was the primary author of the paper with significant
+contributions from S.K.H and A.v.d.B. All authors were involved in the revision
+of the draft manuscript and have agreed to the final content.
 
-\subsection*{Competing interests}
+\section*{Competing interests}
 The authors have no financial, personal, or professional competing interests
@@
 
-\subsection*{Grant Information}
+\section*{Grant Information}
 The work was partially funded by the State of Ohio Third Frontier Commission
@@
 
-\subsection*{Acknowledgments}
+\section*{Acknowledgments}
 We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.

### 2014-12-15 paper.tex
Preparation for PeerJ preprint submission.
- Added methods section.
- Changed to 2.5cm margins.
- Changed to 12pt font.
- Repaired some citations.
- Added line numbers.

@@
 
-\documentclass[fleqn,10pt]{wlpeerj}
+% 12pt font is recommend for the preprint
+\documentclass[fleqn,12pt]{wlpeerj}
+
+% for the preprint we add line numbers
+\usepackage{lineno}
+\linenumbers
 
@@
 widely disseminated, well organized, nor available with few or no restrictions.
-David Winter's published normative gait data~\cite{Winter1990} is widely used
+David Winter's published normative gait data, \cite{Winter1990}, is widely used
 in biomechanical studies, yet it comes from relatively few subjects and only a
@@
 inspired many other studies, such as powered prosthetic control
-design~\cite{Sup2008}, but success in other research fields using large sets of
-data for discovery lead one to believe that more elaborate data sets may
+design, \cite{Sup2008}, but success in other research fields using large sets
+of data for discovery lead one to believe that more elaborate data sets may
 benefit the field of human motion studies. To enable such work, biomechanical
-data needs to be shared extensively, organized, and curated to enabled
-future analysts.
+data needs to be shared extensively, organized, and curated to enabled future
+analysts.
 
@@
 approximately 1995 that includes data sets for download and mostly unencumbered
-use. For example, Vaughn, et. al's data~\cite{Vaughan1992} with kinematics and
-force plate measurement from several subjects is available on the site. At
-another website, the CGA Normative Gait Database~\cite{Kirtley2014}, Chris
+use. For example, Vaughn, et. al's data, \cite{Vaughan1992}, with kinematics
+and force plate measurement from several subjects is available on the site. At
+another website, the CGA Normative Gait Database, \cite{Kirtley2014}, Chris
 Kirtley shares normative gait data from several studies and these files have
 influenced other studies, for example the average gait cycles from children
-used in~\cite{Bogert2003}.
+used in \cite{Bogert2003}.
 
-Chester et. al~\cite{Chester2007} report on a large gait database comparison
+Chester et. al, \cite{Chester2007}, report on a large gait database comparison
 where one database contained kinematic data of 409 gait cycles of children from
@@
 comprehensive data base for clinical gait data and created the
-Gaitabase~\cite{Tirosh2010}. This database may contain a substantial amount of
+Gaitabase, \cite{Tirosh2010}. This database may contain a substantial amount of
 data but it is encumbered by a very complicated and restrictive license and
@@
 University of Wisconsin at LaCrosse has an easily accessible normative gait
-data set~\cite{Wilson2014} from 25 subjects with lower extremity marker data
+data set, \cite{Willson2014}, from 25 subjects with lower extremity marker data
 from multiple gait cycles and force plate measurements from a single gait
@@
 with rich meta data and made available in the most unrestrictive form for other
-research uses following modern best practices in data sharing~\cite{White2013}.
+research uses following modern best practices in data sharing, \cite{White2013}.
 
@@
 
-\section*{Participants}
+\section*{Methods}
+\subsection*{Participants}
 %
@@
 
-\section*{Equipment}
+\subsection*{Equipment}
 %
@@
 Cleveland State University, using the following equipment:
-
+%
 \begin{itemize}
@@
 
-\section*{Protocol}
+\subsection*{Protocol}
 %
@@
 
-\section*{Perturbation Signals}
+\subsection*{Perturbation Signals}
 %
@@
 \end{figure}
-
-\section*{Raw Data}
+\section*{Results}
+\subsection*{Raw Data}
 %
@@
 
-\subsection*{mocap-xxx.txt}
+\subsubsection*{mocap-xxx.txt}
 
@@
 
-\subsection*{record-xxx.txt}
+\subsubsection*{record-xxx.txt}
 %
@@
 
-\subsection*{meta-xxx.yml}
+\subsubsection*{meta-xxx.yml}
 
@@
 
-\section*{Processed Data}
+\subsection*{Processed Data}
 %
@@
 
-\section*{Data Limitations}
+\subsection*{Data Limitations}
 %

### 2014-12-15 paper.tex

@@
 %
-We have presented the details of what we believe to be a rich and elaborate
-data set of motion and ground reaction loads humans exhibit during both normal
-walking and when recovering from longitudinal perturbations. The data provides
-a very large sample of gait cycles from both unperturbed and perturbed walking
-at three different nominal speeds. The raw data is provided for reuse with
-complete meta data. In addition to the data, we provide software that can
-process the data for both cleaning purposes and to produce typical sagittal
-plane gait variables of interest. The variation of these gait variables and
-statistics are provided to give a high level picture of the character of the
-data. Among other uses, we believe the dataset is ideally suited for control
-identification purposes. Many researchers are working on mathematical models
-for control in gait and this dataset provides both a way to validate these
-models and a source for generating them.
+We have presented a rich and elaborate data set of motion and ground reaction
+loads from human subjects during both normal walking and when recovering from
+longitudinal perturbations. The raw data is provided for reuse with complete
+meta data. In addition to the data, we provide software that can process the
+data for both cleaning purposes and to produce typical sagittal plane gait
+variables of interest. Among other uses, we believe the dataset is ideally
+suited for control identification purposes. Many researchers are working on
+mathematical models for control in gait and this dataset provides both a way to
+validate these models and a source for generating them.
 

### 2014-12-15 paper.tex
Added 'written' to the informed consent sentence.

@@
 participated in the study. The study was approved by the Institutional Review
-Board of Cleveland State University (\# 29904-VAN-HS) and informed consent was
-obtained from all participants. The data has been anonymized with respect to the
-participants' identities and a unique identification number was assigned to
-each subject. A selection of the meta data collected for each subject is shown
-in Table~\ref{tab:subjects}.
+Board of Cleveland State University (\# 29904-VAN-HS) and written informed
+consent was obtained from all participants. The data has been anonymized with
+respect to the participants' identities and a unique identification number was
+assigned to each subject. A selection of the meta data collected for each
+subject is shown in Table~\ref{tab:subjects}.
 %

### 2014-12-15 paper.tex
Made some formatting adjustments to listings for the 12pt font.

@@
 \begin{listing}
-  \small
+  \scriptsize
   \begin{minted}[gobble=4]{yaml}
@@
     >>> from gaitanalysis.motek import DFlowData
-    >>> data = DFlowData('mocap-020.txt', 'record-020.txt', 'meta-020.yml')
+    >>> data = DFlowData('mocap-020.txt', 'record-020.txt',
+    ...                  'meta-020.yml')
     >>> mass = data.meta['subject']['mass']
@@
     >>> gdata = GaitData(event_df)
-    >>> gdata.inverse_dynamics_2d(left_markers, right_markers, left_loads,
-    ...                           right_loads, mass, 6.0)
+    >>> gdata.inverse_dynamics_2d(left_markers, right_markers,
+    ...                           left_loads, right_loads, mass, 6.0)
     >>> gdata.grf_landmarks('Right Fy', 'Left Fy', threshhold=20.0)

### 2014-12-15 src/treadmill_motion.py
Added python version of fft on treadmill speed.

@@ -1,0 +1,41 @@
+import os
+
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.signal import parzen
+from scipy.fftpack import fft
+
+import utils
+
+script_path = os.path.realpath(__file__)
+src_dir = os.path.dirname(script_path)
+root_dir = os.path.realpath(os.path.join(src_dir, '..'))
+raw_dir, tmp = utils.config_paths(root_dir)
+
+trial_number = '020'
+paths = utils.trial_file_paths(raw_dir, trial_number)
+
+gait_data = utils.load_data('Longitudinal Perturbation', paths, tmp)
+
+time = gait_data.data.index.values.astype(float)
+belt_speed = gait_data.data['LeftBeltSpeed'].values
+belt_speed -= belt_speed.mean()
+
+N = len(time)
+T = 1.0 / 100.0
+
+yf = fft(belt_speed)
+window = parzen(N)
+ywf = fft(belt_speed * window)
+
+xf = np.linspace(0.0, 1.0 / (2.0 * T), N / 2)
+
+plt.semilogy(xf[1:N / 2], utils.smooth(2.0 / N * np.abs(yf[1:N / 2]), 20), '-b')
+plt.semilogy(xf[1:N / 2], utils.smooth(2.0 / N * np.abs(ywf[1:N / 2]), 20), '-r')
+
+plt.xlim((0.0, 8.0))
+
+plt.legend(['FFT', 'FFT w. window'])
+
+plt.grid()
+plt.show()

### 2014-12-15 src/utils.py

@@ -367,3 +367,62 @@ def measured_subject_mass(raw_data_dir, processed_data_dir):
     mean['Standard Deviation'] = grouped.agg({'Standard Deviation': uncert})
 
     return mean
+
+
+def smooth(x, window_len=11, window='flat'):
+    """Smooth the data using a window with requested size.
+
+    This method is based on the convolution of a scaled window with the
+    signal.  The signal is prepared by introducing reflected copies of the
+    signal (with the window size) in both ends so that transient parts are
+    minimized in the begining and end part of the output signal.
+
+    Parameters
+    ----------
+    x: the input signal
+    window_len: the dimension of the smoothing window; should be an odd
+    integer
+    window: the type of window from 'flat', 'hanning', 'hamming',
+    'bartlett', 'blackman' flat window will produce a moving average
+    smoothing.
+
+    Returns
+    -------
+    output:
+        the smoothed signal
+
+    Examples
+    --------
+
+    t=linspace(-2,2,0.1)
+    x=sin(t)+randn(len(t))*0.1
+    y=smooth(x)
+
+    See also
+    --------
+
+    numpy.hanning, numpy.hamming, numpy.bartlett, numpy.blackman,
+    numpy.convolve scipy.signal.lfilter
+
+    """
+
+    if x.ndim != 1:
+        raise ValueError("smooth only accepts 1 dimension arrays.")
+
+    if x.size < window_len:
+        raise ValueError("Input vector needs to be bigger than window size.")
+
+    if window_len < 3:
+        return x
+
+    if window not in ['flat', 'hanning', 'hamming', 'bartlett', 'blackman']:
+        raise ValueError("Window is on of 'flat', 'hanning', 'hamming', 'bartlett', 'blackman'")
+
+    s = np.r_[x[window_len - 1:0:-1], x, x[-1:-window_len:-1]]
+    if window == 'flat':  # moving average
+        w = np.ones(window_len, 'd')
+    else:
+        w = eval('np.' + window + '(window_len)')
+
+    y = np.convolve(w/w.sum(), s, mode='valid')
+    return y[(window_len / 2 - 1):-(window_len / 2)]

### 2014-12-16 paper.tex
Fixed error in the DFlowData example.
Spelling.

@@
     >>> data.clean_data()
-    >>> event_df = dflow_data.extract_processed_data(
+    >>> event_df =data.extract_processed_data(
     ...     event='Longitudinal Perturbation')
@@
     \verb|GaitData| class to load in the result of \verb|DFlowData| and compute
-    the inverse dynamics (joint angles and torques), identify the gait event
-    (e.g. heel strikes), split the data with respect to the gait events in a
+    the inverse dynamics (joint angles and torques), identify the gait events
+    (e.g. heel strikes), split the data with respect to the gait events into a
     Pandas \verb|Panel|, plot the mean and standard deviation of one time

### 2014-12-16 src/treadmill_motion.py
Changed to hamming window add second figure.

@@ -2,7 +2,7 @@ import os
 
 import numpy as np
 import matplotlib.pyplot as plt
-from scipy.signal import parzen
+from scipy.signal import hamming
 from scipy.fftpack import fft
 
 import utils
@@ -25,17 +25,25 @@ N = len(time)
 T = 1.0 / 100.0
 
 yf = fft(belt_speed)
-window = parzen(N)
+window = hamming(N)
 ywf = fft(belt_speed * window)
 
 xf = np.linspace(0.0, 1.0 / (2.0 * T), N / 2)
 
-plt.semilogy(xf[1:N / 2], utils.smooth(2.0 / N * np.abs(yf[1:N / 2]), 20), '-b')
-plt.semilogy(xf[1:N / 2], utils.smooth(2.0 / N * np.abs(ywf[1:N / 2]), 20), '-r')
+amp_f = 2.0 / N * np.abs(yf[1:N / 2])
+amp_wf = 2.0 / N * np.abs(ywf[1:N / 2])
 
-plt.xlim((0.0, 8.0))
+fig, ax = plt.subplots()
+ax.semilogy(xf[1:N / 2], amp_f, '-b')
+ax.semilogy(xf[1:N / 2], utils.smooth(amp_wf, 20), '-r')
+ax.set_xlim((0.0, 8.0))
+ax.legend(['FFT', 'FFT Moving Average'])
 
-plt.legend(['FFT', 'FFT w. window'])
+fig, ax = plt.subplots()
+ax.semilogy(xf[1:N / 2], utils.smooth(amp_f, 20), '-b')
+ax.semilogy(xf[1:N / 2], utils.smooth(amp_wf, 20), '-r')
+ax.set_xlim((0.0, 8.0))
+ax.legend(['FFT', 'FFT w. window'])
 
 plt.grid()
 plt.show()

### 2014-12-16 src/treadmill_motion.py
Added trial number to the serialized file.

@@ -15,7 +15,8 @@ raw_dir, tmp = utils.config_paths(root_dir)
 trial_number = '020'
 paths = utils.trial_file_paths(raw_dir, trial_number)
 
-gait_data = utils.load_data('Longitudinal Perturbation', paths, tmp)
+gait_data = utils.load_data(trial_number, 'Longitudinal Perturbation',
+                            paths, tmp)
 
 time = gait_data.data.index.values.astype(float)
 belt_speed = gait_data.data['LeftBeltSpeed'].values

### 2014-12-16 src/utils.py

@@ -65,12 +65,14 @@ def trial_file_paths(trials_dir, trial_number):
     return mocap_file_path, record_file_path, meta_file_path
 
 
-def load_data(event, paths, tmp):
+def load_data(trial_num, event, paths, tmp):
     """Loads an event and processes the data, if necessary, from a trial
     into a GaitData object.
 
     Parameters
     ==========
+    trial_num : string
+        Three digit trial number.
     event : string
         A valid event for the given trial.
     paths : list of strings
@@ -86,7 +88,9 @@ def load_data(event, paths, tmp):
 
     """
 
-    file_name = '_'.join([n.lower() for n in event.split(' ')]) + '.h5'
+    parts = [n.lower() for n in event.split(' ')]
+    parts.append(trial_num)
+    file_name = '_'.join(parts) + '.h5'
 
     tmp_data_path = os.path.join(tmp, file_name)
 

### 2014-12-16 src/treadmill_motion.py
Added spectogram

@@ -2,8 +2,9 @@ import os
 
 import numpy as np
 import matplotlib.pyplot as plt
-from scipy.signal import hamming
+from scipy.signal import hamming, periodogram
 from scipy.fftpack import fft
+from matplotlib.mlab import specgram, window_hanning
 
 import utils
 
@@ -47,4 +48,34 @@ ax.set_xlim((0.0, 8.0))
 ax.legend(['FFT', 'FFT w. window'])
 
 plt.grid()
+
+# Periodogram
+
+belt_speed = gait_data.data['LeftBeltSpeed'].values
+f, Pxx = periodogram(belt_speed, fs=100.0, window='hamming', nfft=len(time),
+                     detrend='constant')
+fig, ax = plt.subplots()
+ax.semilogy(f, Pxx)
+ax.set_xlim((0.0, 8.0))
+
+# Spectogram
+fig, ax = plt.subplots(2)
+ax[0].specgram(belt_speed,
+               NFFT=100,
+               Fs=100,
+               detrend='mean',
+               window=window_hanning,
+               noverlap=50)
+
+spectrum, freqs, t = specgram(belt_speed,
+                              NFFT=100,
+                              Fs=100,
+                              detrend='mean',
+                              window=window_hanning,
+                              noverlap=50)
+
+amp = spectrum.mean(axis=1)
+ax[1].semilogy(freqs, amp)
+ax[1].set_xlim((0.0, 8.0))
+
 plt.show()

### 2014-12-16 src/treadmill_motion.py
Changed spectogram to 48 sections at 1000 samples each.

@@ -61,18 +61,19 @@ ax.set_xlim((0.0, 8.0))
 # Spectogram
 fig, ax = plt.subplots(2)
 ax[0].specgram(belt_speed,
-               NFFT=100,
-               Fs=100,
+               NFFT=1000,
+               Fs=100.0,
                detrend='mean',
                window=window_hanning,
-               noverlap=50)
+               noverlap=500)
+ax[0].set_xlim((0.0, 8.0))
 
 spectrum, freqs, t = specgram(belt_speed,
-                              NFFT=100,
-                              Fs=100,
+                              NFFT=1000,
+                              Fs=100.0,
                               detrend='mean',
                               window=window_hanning,
-                              noverlap=50)
+                              noverlap=500)
 
 amp = spectrum.mean(axis=1)
 ax[1].semilogy(freqs, amp)

### 2014-12-16 src/treadmill_motion.py
Removed log scales from spectogram.

@@ -50,7 +50,6 @@ ax.legend(['FFT', 'FFT w. window'])
 plt.grid()
 
 # Periodogram
-
 belt_speed = gait_data.data['LeftBeltSpeed'].values
 f, Pxx = periodogram(belt_speed, fs=100.0, window='hamming', nfft=len(time),
                      detrend='constant')
@@ -76,7 +75,7 @@ spectrum, freqs, t = specgram(belt_speed,
                               noverlap=500)
 
 amp = spectrum.mean(axis=1)
-ax[1].semilogy(freqs, amp)
+ax[1].plot(freqs, amp)
 ax[1].set_xlim((0.0, 8.0))
 
 plt.show()

### 2014-12-16 src/treadmill_motion.py
Using 20 second sections in the spectogram.

@@ -60,19 +60,19 @@ ax.set_xlim((0.0, 8.0))
 # Spectogram
 fig, ax = plt.subplots(2)
 ax[0].specgram(belt_speed,
-               NFFT=1000,
+               NFFT=2000,
                Fs=100.0,
                detrend='mean',
                window=window_hanning,
-               noverlap=500)
+               noverlap=1000)
 ax[0].set_xlim((0.0, 8.0))
 
 spectrum, freqs, t = specgram(belt_speed,
-                              NFFT=1000,
+                              NFFT=2000,
                               Fs=100.0,
                               detrend='mean',
                               window=window_hanning,
-                              noverlap=500)
+                              noverlap=1000)
 
 amp = spectrum.mean(axis=1)
 ax[1].plot(freqs, amp)

### 2014-12-17 paper.tex

@@
 benefit the field of human motion studies. To enable such work, biomechanical
-data needs to be shared extensively, organized, and curated to enabled future
+data needs to be shared extensively, organized, and curated to enable future
 analysts.
@@
 and force plate measurement from several subjects is available on the site. At
-another website, the CGA Normative Gait Database, \cite{Kirtley2014}, Chris
-Kirtley shares normative gait data from several studies and these files have
-influenced other studies, for example the average gait cycles from children
-used in \cite{Bogert2003}.
+another website, the CGA Normative Gait Database, \cite{Kirtley2014} shares
+normative gait data from several studies and these files have influenced other
+studies, for example the average gait cycles from children used in
+\cite{Bogert2003}.
 
-Chester et. al, \cite{Chester2007}, report on a large gait database comparison
-where one database contained kinematic data of 409 gait cycles of children from
-1 to 7 years old but the data does not seem to be publicly available. This is
+\cite{Chester2007} report on a large gait database comparison where one
+database contained kinematic data of 409 gait cycles of children from 1 to 7
+years old but the data does not seem to be publicly available. This is
 unfortunately typical. But Tirosh et. al, recognized the need for a
-comprehensive data base for clinical gait data and created the
-Gaitabase, \cite{Tirosh2010}. This database may contain a substantial amount of
-data but it is encumbered by a very complicated and restrictive license and
-sharing scheme. However, there are examples of data with less restrictions. The
+comprehensive data base for clinical gait data and created the Gaitabase,
+\cite{Tirosh2010}. This database may contain a substantial amount of data but
+it is encumbered by a very complicated and restrictive license and sharing
+scheme. However, there are examples of data with less restrictions. The
 University of Wisconsin at LaCrosse has an easily accessible normative gait
@@
 are affected by this vibration mode. During trials 9-15 the treadmill was
-stabilized with wooden blocks. During, the remaining trials the treadmill was
+stabilized with wooden blocks. During the remaining trials the treadmill was
 stabilized with metal supports. See the Data Limitations Section for more
@@
 
-The MATLAB script and Simulink model produce a comma-delimited text file of six
-signals: time stamp, slow, normal, and fast walking perturbation signals, and
-slow and fast running signals.~\footnote{The running signals were not used in
-the experiments presented in this paper.} The measured speed of the treadmill
-belts are compared to the control input signals in
-Figure~\ref{fig:input_output} to show the effect of the treadmill and
-controller dynamics. The system introduces a delay and seems to act as a low
-pass filter. The standard deviations of the outputs do not significantly differ
-from the desired values: \SIlist{0.05;0.12;0.2}{\meter\per\second} for the
-three speeds, respectively.
+The MATLAB script and Simulink model produce a comma-delimited text file of
+five desired belt speed signals indexed by the time stamp. There are slow, normal,
+and fast walking speeds and slow and fast running speeds.~\footnote{The running
+signals were not used in the experiments presented in this paper.} The measured
+speed of the treadmill belts are compared to these commanded treadmill control
+input signals in Figure~\ref{fig:input_output} to show the effect of the
+treadmill and controller dynamics. The system introduces a delay and seems to
+act as a low pass filter. The standard deviations of the measured speeds do not
+significantly differ from the desired values:
+\SIlist{0.05;0.12;0.2}{\meter\per\second} for the three speeds, respectively.
 %
@@
   \cprotect\caption{Box plots of the average belt speed, stride frequency, and
-    stride length which compare gait cycles for the unperturbed (purple) and
-    perturbed (blue). Produced by \verb|src/unperturbed_perturbed_comparison.py|.}
+    stride length which compare unperturbed (purple) and perturbed (blue) gait
+    cycles. The median is given with the box bounding the first and third
+    quartiles and the whiskers bound the range of the data. Produced by
+    \verb|src/unperturbed_perturbed_comparison.py|.}
   \label{fig:gait-cycle-stats-comparison}
@@
     concrete floor to eliminate the treadmill's low vibration mode
-    (\~5\si{\hertz}). But these blocks seem to have corrupted the force plate
-    measurements by imposing frictional stresses on the system. The force plate
-    measurements should not be used from these trials, but the marker data is
-    fine.
+    (approximately 5\si{\hertz}). But these blocks seem to have corrupted the
+    force plate measurements by imposing frictional stresses on the system. The
+    force plate measurements should not be used from these trials, but the
+    marker data is fine.
   \item Trials 6-8 use an early experimental protocol which divided the

### 2014-12-17 paper.tex
Added NSF acknowledgement, closes #86.

--- 2014-12-16 paper.tex
@@
 The work was partially funded by the State of Ohio Third Frontier Commission
-through the Wright Center for Sensor Systems Engineering (WCSSE).
+through the Wright Center for Sensor Systems Engineering (WCSSE) and by the
+National Science Foundation under Grant No. 1344954.
 

### 2014-12-17 src/unperturbed_perturbed_comparison.py
Changed unperturbed color to 'red' and removed left perturbed/unperturbed lines and shading

@@ -80,7 +80,7 @@ y_labels = ['Ankle Plantar Flexion',
 fig, axes = plt.subplots(3, 2, sharex=True)
 
 blue = sbn.xkcd_rgb['windows blue']
-purple = sbn.xkcd_rgb['dusty purple']
+red = sbn.xkcd_rgb['pale red']
 
 ppercent = mean_of_perturbed.index.values.astype(float)
 upercent = mean_of_unperturbed['Percent Gait Cycle']
@@ -88,7 +88,7 @@ upercent = mean_of_unperturbed['Percent Gait Cycle']
 con = [lambda x: rad2deg(x), lambda x: x]
 
 for row, angle, torque, y_label in zip(axes, angles, torques, y_labels):
-    for side, linetype in zip(['Right', 'Left'], ['-', '--']):
+    for side, linetype in zip(['Right'], ['-']):
 
         row[0].set_ylabel(y_label)
 
@@ -115,13 +115,13 @@ for row, angle, torque, y_label in zip(axes, angles, torques, y_labels):
                                 (umean - ustd).values,
                                 (umean + ustd).values,
                                 alpha=0.40,
-                                color=purple,
+                                color=red,
                                 label='_nolegend_')
             label = side + ' Perturbed, N = {}'.format(num_perturbed_cycles)
             row[i].plot(ppercent, pmean.values, linetype, color=blue,
                         label=label)
             label = side + ' Unperturbed, N = {}'.format(num_unperturbed_cycles)
-            row[i].plot(upercent, umean.values, linetype, color=purple,
+            row[i].plot(upercent, umean.values, linetype, color=red,
                         label=label)
 
 plt.subplots_adjust(top=0.85)
@@ -156,7 +156,7 @@ for col, ax, unit in zip(columns, axes.flatten(), units):
     ax.set_title('{} {}'.format(col, unit))
     u = unperturbed_gait_cycle_stats[col]
     p = perturbed_gait_data.gait_cycle_stats[col]
-    sbn.boxplot([p.values, u.values], ax=ax, color=[blue, purple],
+    sbn.boxplot([p.values, u.values], ax=ax, color=[blue, red],
                 whis='range',
                 names=['Perturbed\nN = {}'.format(num_perturbed_cycles),
                        'Unperturbed\nN = {}'.format(num_unperturbed_cycles)])

### 2014-12-17 paper.tex
added new frequency plots and code
1) Changed code to average a spectrogram over 20 seconds in the hamming
window.  Also interpolated all signals to the same sampling rate and
clipped out the sections that were not under the influence of
longitudinal perturbations
2) Changed "dusty purple" to "red" in each code and updated legend
3) Edited the text in the report to reflect the results in the new
graphs
4) Fonts are still strange when saving to pdf

@@
 To show the effects of the treadmill dynamics and give an idea of the frequency
-content of the actual perturbations, the input and output for each speed were
-transformed into the frequency domain using the Fast Fourier algorithm, and the
-results are shown in Figure~\ref{fig:freq_analysis}. This shows that for the
-1.2~\si{\meter\per\second} walking speed, the amplitude of the output is
-significantly lower than the amplitude of the input signal at lower
-frequencies. Additionally, the amplitude of output signal in the
-0.8~\si{\meter\per\second} walking speed begins to attenuate around
-2~\si{\hertz}, which is a noticeably lower frequency than the other walking
-speeds.
+content of the actual perturbations, spectral analysis plots were created 
+by averaging a spectrogram over twenty seconds in the hamming window, shown 
+in Figure~\ref{fig:freq_analysis}. For all speeds, the frequency content of the
+commanded and measured belt speeds were noticeably similar.  
 %

### 2014-12-17 paper.tex
Added acknowledgements for Anne, Sabrina, Brad, and Dan.

--- 2014-12-17 paper.tex
@@
 \section*{Acknowledgments}
+%
 We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.
+We also thank Sabrina Abram, Brad Humphreys, and Anne Koelewijn for reviewing
+the preprint and being our guinea pigs on the software/data instructions. Dan
+Simon also gave valuable feedback on the preprint.
 

### 2014-12-17 paper.tex
minor changes to text in report

--- 2014-12-17 paper.tex
@@
 To show the effects of the treadmill dynamics and give an idea of the frequency
-content of the actual perturbations, spectral analysis plots were created 
-by averaging a spectrogram over twenty seconds in the hamming window, shown 
+content of the actual perturbations, spectral density plots were created 
+by averaging a spectrogram over twenty seconds in the Hamming window, shown 
 in Figure~\ref{fig:freq_analysis}. For all speeds, the frequency content of the

### 2014-12-18 paper.tex
Subject table is simpler, closes #83 and #96.
- Got rid of self-reported mass column and replaced the NA's in the Measured
  Mass column with the self-reported mass. Noted this in the caption of the
  table. issue #83
- Added a mkdir to the subject table so that you can run it first in a fresh
  repo with no issues. issue #96

--- 2014-12-17 paper.tex
@@
   \cprotect\caption{Information about the 15 participants. The final three
-    columns give the trial numbers associated with each nominal treadmill
+    columns provide the trial numbers associated with each nominal treadmill
     speed. The measured mass is computed from the mean total vertical ground
-    reaction force just after the calibration pose event, if possible.
-    Additional trials found in the data set with a subject identification
-    number 0 are trials with no subject, i.e.  unloaded trials that can be used
-    for inertial compensation purposes, and are not shown in the table. Generated
+    reaction force just after the calibration pose event, if possible. If the
+    mass is reported without an accompanying standard deviation, it is the
+    subject's self-reported mass. Additional trials found in the data set with
+    a subject identification number 0 are trials with no subject, i.e.
+    unloaded trials that can be used for inertial compensation purposes, and
+    are not shown in the table. Generated
     by \verb|src/subject_table.py|.}
@@
 \section*{Acknowledgments}
-%
 We thank Roman Boychuk and Obinna Nwanna for assistance with the experiments.
-We also thank Sabrina Abram, Brad Humphreys, and Anne Koelewijn for reviewing
-the preprint and being our guinea pigs on the software/data instructions. Dan
-Simon also gave valuable feedback on the preprint.
 

### 2014-12-18 src/utils.py

@@ -249,7 +249,6 @@ def measured_subject_mass(raw_data_dir, processed_data_dir):
         self reported mass, and indexed by subject id.
 
     """
-
     # Subject 0 is for the null subject. For subject 1 we use the self
     # reported value because there is no "Calibration Pose" event. For
     # subject 11 and subject 4, we use the self reported mass because the
@@ -281,6 +280,9 @@ def measured_subject_mass(raw_data_dir, processed_data_dir):
 
     tmp_file_name = '_'.join(event.lower().split(' ')) + '.h5'
     tmp_data_path = os.path.join(processed_data_dir, tmp_file_name)
+
+    if not os.path.exists(processed_data_dir):
+        os.makedirs(processed_data_dir)
 
     subject_data = defaultdict(list)
 
@@ -337,9 +339,9 @@ def measured_subject_mass(raw_data_dir, processed_data_dir):
 
             subject_data['Trial Number'].append(trial_number)
             subject_data['Subject ID'].append(dflow_data.meta['subject']['id'])
-            subject_data['Reported Mass'].append(reported_mass)
-            subject_data['Measured Mass'].append(actual_mass)
-            subject_data['Standard Deviation'].append(std)
+            subject_data['Self Reported Mass'].append(reported_mass)
+            subject_data['Mean Measured Mass'].append(actual_mass)
+            subject_data['Measured Mass Std. Dev.'].append(std)
             subject_data['Gender'].append(dflow_data.meta['subject']['gender'])
 
             print("Measured mass: {} kg".format(actual_mass))
@@ -356,7 +358,7 @@ def measured_subject_mass(raw_data_dir, processed_data_dir):
 
     mean = grouped.mean()
 
-    mean['Diff'] = mean['Measured Mass'] - mean['Reported Mass']
+    mean['Diff'] = mean['Mean Measured Mass'] - mean['Self Reported Mass']
 
     # This sets the grouped standard deviation to the correct value
     # following uncertainty propagation for the mean function.
@@ -364,6 +366,7 @@ def measured_subject_mass(raw_data_dir, processed_data_dir):
     def uncert(x):
         return np.sqrt(np.sum(x**2) / len(x))
 
-    mean['Standard Deviation'] = grouped.agg({'Standard Deviation': uncert})
+    mean['Measured Mass Std. Dev.'] = \
+        grouped.agg({'Measured Mass Std. Dev.': uncert})
 
     return mean

### 2014-12-18 src/subject_table.py
Minor mass formatting in subjects table.

@@ -74,8 +74,7 @@ unique_subjects['mass'] = measured['Mass']
 # for each subject who has an NA for the measured mass, replace that with
 # the self reported mass and an asterik
 na_idx = unique_subjects['mass'].isnull()
-unique_subjects['mass'][na_idx] = \
-    unique_subjects['self-reported mass'][na_idx].astype(str)
+unique_subjects['mass'][na_idx] = unique_subjects['self-reported mass'][na_idx]
 
 # Make nicer column names.
 cols = ['id', 'gender', 'age', 'height', 'mass', '0.8 m/s', '1.2 m/s',
@@ -86,7 +85,8 @@ unique_subjects.rename(columns=dict(zip(cols, new_cols)), inplace=True)
 
 
 formatters = {'Height [m]': lambda x: 'NA' if np.isnan(x)
-              else '{:0.2f}'.format(x)}
+              else '{:0.2f}'.format(x),
+              'Mass [kg]': lambda x: '{:0.0f}'.format(x) if isinstance(x, float) else x}
 
 table_dir = os.path.join(root_dir, 'tables')
 if not os.path.exists(table_dir):

### 2014-12-18 paper.tex
Adds in a new table showing the unloaded trials, closes #102.

--- 2014-12-18 paper.tex
@@
 %
-\begin{table*}
+\begin{table}
   \cprotect\caption{Information about the 15 participants. The final three
-    columns give the trial numbers associated with each nominal treadmill
+    columns provide the trial numbers associated with each nominal treadmill
     speed. The measured mass is computed from the mean total vertical ground
-    reaction force just after the calibration pose event, if possible.
-    Additional trials found in the data set with a subject identification
-    number 0 are trials with no subject, i.e.  unloaded trials that can be used
-    for inertial compensation purposes, and are not shown in the table. Generated
+    reaction force just after the calibration pose event, if possible. If the
+    mass is reported without an accompanying standard deviation, it is the
+    subject's self-reported mass. Additional trials found in the data set with
+    a subject identification number 0 are trials with no subject, i.e.
+    unloaded trials that can be used for inertial compensation purposes, and
+    are not shown in the table. Generated
     by \verb|src/subject_table.py|.}
@@
   \label{tab:subjects}
-\end{table*}
+\end{table}
+%
+\begin{table}
+  \cprotect\caption{A list of unloaded trials collected for each speed. Each
+    loaded trial includes a compensation file listed in its meta data which
+    matches it to these unloaded trials. Generated by
+    \verb|src/subject_table.py|.}
+  \centering
+  \small
+  \input{tables/compensation_trials.tex}
+  \label{tab:comptrials}
+\end{table}
 

### 2014-12-18 src/frequency_analysis.m
addresses #39 and #106

@@ -31,18 +31,6 @@ clc
 %-------------------------------------------------------------------------
 %Spectral Analysis
 %-------------------------------------------------------------------------
-    %Settings for Nice Plot Appearance 
-        set(0,'defaulttextinterpreter','latex')
-        set(0,'DefaultAxesFontName', 'latex')
-        ti = get(gca,'TightInset');
-        set(gca,'Position',[ti(1) ti(2) 1-ti(3)-ti(1) 1-ti(4)-ti(2)]);
-        set(gca,'units','centimeters')
-        pos = get(gca,'Position');
-        ti = get(gca,'TightInset');
-        set(gcf, 'PaperUnits','centimeters');
-        set(gcf, 'PaperSize', [pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
-        set(gcf, 'PaperPositionMode', 'manual');
-        set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
     %Analysis
         fs=100;
         T=1/fs;
@@ -68,10 +56,10 @@ clc
                 hold on
                 semilogy(f_output,p_output_mean,'Color',[0.815 0.3294 0.3020])
                 ylabel('PSD (v^2/Hz)')
-                xlim([0 8])
+                xlim([0 8]); ylim([10^-5 10^0])
                 title(titles{i})
                 if i==1
-                    legend('Commanded','Measured','Location','southeast')
+                    legend('Commanded','Measured','Location','northeast')
                 end
                 if i==3
                     xlabel('Frequency (Hz)');

### 2014-12-18 paper.tex
Added reference to new table.
Fixed caption for the joint angle/torque plot.

@@
   \includegraphics{figures/unperturbed-perturbed-comparison.pdf}
-  \cprotect\caption{Mean (right: solid, left: dashed) and $3\sigma$ (shaded)
-    joint angles and torques from both unperturbed (purple) and perturbed
-    (blue) gait cycles from trial 20. Produced by
-    \verb|src/unperturbed_perturbed_comparison.py|.}
+  \cprotect\caption{Right leg mean and $3\sigma$ (shaded) joint angles and
+    torques from both unperturbed (red) and perturbed (blue) gait cycles
+    from trial 20. Produced by \verb|src/unperturbed_perturbed_comparison.py|.}
   \label{fig:angle-torque-comparison}
@@
     same day as the loaded trials and is noted in the
-    \verb|trial:files:compensation| section of the meta data file.
+    \verb|trial:files:compensation| section of the meta data file. See
+    Table~\ref{tab:comptrials} for a list of all the compensation trials.
   \item Trials 1 and 2 were not recorded as part of this study. Those trial

### 2014-12-18 src/unperturbed_perturbed_comparison.py

@@ -88,41 +88,42 @@ upercent = mean_of_unperturbed['Percent Gait Cycle']
 con = [lambda x: rad2deg(x), lambda x: x]
 
 for row, angle, torque, y_label in zip(axes, angles, torques, y_labels):
-    for side, linetype in zip(['Right'], ['-']):
+    side = 'Right'
+    linetype = '-'
 
-        row[0].set_ylabel(y_label)
+    row[0].set_ylabel(y_label)
 
-        for i, col in enumerate([angle, torque]):
+    for i, col in enumerate([angle, torque]):
 
-            col_name = side + '.' + col
+        col_name = side + '.' + col
 
-            sigma_mul = 3.0
+        sigma_mul = 3.0
 
-            pmean = con[i](mean_of_perturbed[col_name])
-            pstd = sigma_mul * con[i](std_of_perturbed[col_name])
+        pmean = con[i](mean_of_perturbed[col_name])
+        pstd = sigma_mul * con[i](std_of_perturbed[col_name])
 
-            umean = con[i](mean_of_unperturbed[col_name])
-            ustd = sigma_mul * con[i](std_of_unperturbed[col_name])
+        umean = con[i](mean_of_unperturbed[col_name])
+        ustd = sigma_mul * con[i](std_of_unperturbed[col_name])
 
-            row[i].fill_between(ppercent,
-                                (pmean - pstd).values,
-                                (pmean + pstd).values,
-                                alpha=0.25,
-                                color=blue,
-                                label='_nolegend_')
+        row[i].fill_between(ppercent,
+                            (pmean - pstd).values,
+                            (pmean + pstd).values,
+                            alpha=0.25,
+                            color=blue,
+                            label='_nolegend_')
 
-            row[i].fill_between(upercent,
-                                (umean - ustd).values,
-                                (umean + ustd).values,
-                                alpha=0.40,
-                                color=red,
-                                label='_nolegend_')
-            label = side + ' Perturbed, N = {}'.format(num_perturbed_cycles)
-            row[i].plot(ppercent, pmean.values, linetype, color=blue,
-                        label=label)
-            label = side + ' Unperturbed, N = {}'.format(num_unperturbed_cycles)
-            row[i].plot(upercent, umean.values, linetype, color=red,
-                        label=label)
[…]

### 2014-12-18 src/frequency_analysis.m
made adjustments for PDF printing

@@ -31,6 +31,16 @@ clc
 %-------------------------------------------------------------------------
 %Spectral Analysis
 %-------------------------------------------------------------------------
+    %Settings to Improve Graph Appearance
+        ti = get(gca,'TightInset');
+        set(gca,'Position',[ti(1) ti(2) 1-ti(3)-ti(1) 1-ti(4)-ti(2)]);
+        set(gca,'units','centimeters')
+        pos = get(gca,'Position');
+        ti = get(gca,'TightInset');
+        set(gcf, 'PaperUnits','centimeters');
+        set(gcf, 'PaperSize', [pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
+        set(gcf, 'PaperPositionMode', 'manual');
+        set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
     %Analysis
         fs=100;
         T=1/fs;

### 2014-12-18 src/signal_check.m

@@ -27,6 +27,16 @@ clc
 %-------------------------------------------------------------------------
 %Plotting
 %-------------------------------------------------------------------------
+    %Settings to Improve Graph Appearance
+        ti = get(gca,'TightInset');
+        set(gca,'Position',[ti(1) ti(2) 1-ti(3)-ti(1) 1-ti(4)-ti(2)]);
+        set(gca,'units','centimeters')
+        pos = get(gca,'Position');
+        ti = get(gca,'TightInset');
+        set(gcf, 'PaperUnits','centimeters');
+        set(gcf, 'PaperSize', [pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
+        set(gcf, 'PaperPositionMode', 'manual');
+        set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
     figure(1)
         subplot(3,1,1)
         plot(time_input, inputs(:,1),'Color',[0.815 0.3294 0.3020])

### 2014-12-18 src/frequency_analysis.m
changed colors
input and output graphs and frequency graphs are now both red for
commanded and blue for measured

@@ -62,9 +62,9 @@ clc
             titles={'Slow Walking (0.8 m/s)','Normal Walking (1.2 m/s)','Fast Walking (1.6 m/s)'};
             figure(1)
                 subplot(speeds,1,i)
-                semilogy(f_input,p_input_mean,'Color',[0.5098 0.3725 0.5294])
+                semilogy(f_input,p_input_mean,'Color',[0.815 0.3294 0.3020])
                 hold on
-                semilogy(f_output,p_output_mean,'Color',[0.815 0.3294 0.3020])
+                semilogy(f_output,p_output_mean,'Color',[0.2157 0.4706 0.7490])
                 ylabel('PSD (v^2/Hz)')
                 xlim([0 8]); ylim([10^-5 10^0])
                 title(titles{i})

### 2014-12-19 paper.tex
Space.
Edited the paper to fix Ton's final comments.

@@
   speed of the treadmill belt. A total of approximately 1.5 hours of normal
-  walking ($>5000$ gait cycles) and 6 hours of perturbed walking ($>20,000$
+  walking ($>$~5000 gait cycles) and 6 hours of perturbed walking ($>$~20,000
   gait cycles) is included in the data set. We provide full body marker
@@
 David Winter's published normative gait data, \cite{Winter1990}, is widely used
-in biomechanical studies, yet it comes from relatively few subjects and only a
-small number of gait cycles per subject. This small source has successfully
-inspired many other studies, such as powered prosthetic control
-design, \cite{Sup2008}, but success in other research fields using large sets
-of data for discovery lead one to believe that more elaborate data sets may
-benefit the field of human motion studies. To enable such work, biomechanical
-data needs to be shared extensively, organized, and curated to enable future
-analysts.
+in biomechanical studies, yet it comes from few subjects and only a small
+number of gait cycles per subject. This small source has successfully enabled
+many other studies, such as powered prosthetic control design, \cite{Sup2008},
+but success in other research fields using large sets of data for discovery
+lead one to believe that more elaborate data sets may benefit the field of
+human motion studies. To enable such work, biomechanical data needs to be
+shared extensively, organized, and curated to enable future analysts.
 
@@
 approximately 1995 that includes data sets for download and mostly unencumbered
-use. For example, Vaughn, et. al's data, \cite{Vaughan1992}, with kinematics
-and force plate measurement from several subjects is available on the site. At
-another website, the CGA Normative Gait Database, \cite{Kirtley2014} shares
-normative gait data from several studies and these files have influenced other
-studies, for example the average gait cycles from children used in
-\cite{Bogert2003}.
+use. For example, the kinematic and force plate measurements from several
+subjects from \cite{Vaughan1992} is available on the site. At another website,
+the CGA Normative Gait Database, \cite{Kirtley2014} shares normative gait data
+from several studies and these files have influenced other studies, for example
+the average gait cycles from children used in \cite{Bogert2003}.
 
@@
 database contained kinematic data of 409 gait cycles of children from 1 to 7
-years old but the data does not seem to be publicly available. This is
-unfortunately typical. But Tirosh et. al, recognized the need for a
-comprehensive data base for clinical gait data and created the Gaitabase,
-\cite{Tirosh2010}. This database may contain a substantial amount of data but
-it is encumbered by a very complicated and restrictive license and sharing
-scheme. However, there are examples of data with less restrictions. The
-University of Wisconsin at LaCrosse has an easily accessible normative gait
-data set, \cite{Willson2014}, from 25 subjects with lower extremity marker data
-from multiple gait cycles and force plate measurements from a single gait
-cycle.
+years old but the data does not seem to be publicly available. This is,
+unfortunately, typical. Tirosh et. al, recognized the need for a comprehensive
+data base for clinical gait data and created the Gaitabase, \cite{Tirosh2010}.
+This database may contain a substantial amount of data but it is encumbered by
+a complicated and restrictive license and sharing scheme. However, there
+are examples of data with less restrictions. The University of Wisconsin at
+LaCrosse has an easily accessible normative gait data set, \cite{Willson2014},
+from 25 subjects with lower extremity marker data from multiple gait cycles and
+force plate measurements from a single gait cycle.
 
@@
 are: \cite{Bogert2013} which includes full body joint kinematics and kinetics
-from eleven subjects walking for a small number of gait cycles and
-\cite{Wang2014} who includes a larger set of data from ten subjects walking for
-five minutes each at three different speeds but only a small set of lower
-extremity markers are present. The second is notable because it publishes the
-data in Dryad, a modern citable data repository.
+from eleven subjects walking on an instrumented treadmill and \cite{Wang2014}
+who includes a larger set of data from ten subjects walking for five minutes
+each at three different speeds but only a small set of lower extremity markers
+are present. The second is notable because it publishes the data in Dryad, a
+modern citable data repository.
 
-The publicly available gait data is small compared to the number of gait
-studies that have been performed over the years. The data that is available
-generally suffers from limitations such as few subjects, few gait cycles, few
-markers, highly clinical, no raw data, limited force plate measurements, lack
-of meta data, non-standard formats, and restrictive licensing. To help with
-this situation we are making the data we collected for our research purposes
-publicly available and free of the previously mentioned deficiencies. Not only
-do we provide a larger set of normative gait data that has been previously
-available, we also include an even larger set of data in which the subject is
-being perturbed, something that does not currently exist. We believe both of
-these sets of data can serve a variety of use cases and hope that we can save
-time and effort for future researchers by sharing it.
+The amount of publicly available gait data is small compared to the number of
+gait studies that have been performed over the years. The data that is
+available generally suffers from limitations such as few subjects, few gait
+cycles, few markers, highly clinical, no raw data, limited force plate
+measurements, lack of meta data, non-standard formats, and restrictive
+licensing. To help with this situation we are making the data we collected for
+our research purposes publicly available and free of the previously mentioned
+deficiencies. Not only do we provide a larger set of normative gait data that
+has been previously available, we also include an even larger set of data in
+which the subject is being perturbed, something that does not currently exist.
+We believe both of these sets of data can serve a variety of use cases and hope
+that we can save time and effort for future researchers by sharing it.
 
@@
 \begin{table}
-  \cprotect\caption{Information about the 15 participants. The final three
-    columns provide the trial numbers associated with each nominal treadmill
-    speed. The measured mass is computed from the mean total vertical ground
-    reaction force just after the calibration pose event, if possible. If the
-    mass is reported without an accompanying standard deviation, it is the
-    subject's self-reported mass. Additional trials found in the data set with
-    a subject identification number 0 are trials with no subject, i.e.
+  \cprotect\caption{Information about the 15 study participants. The final
+    three columns provide the trial numbers associated with each nominal
+    treadmill speed. The measured mass is computed from the mean total vertical
+    ground reaction force just after the calibration pose event, if possible.
+    If the mass is reported without an accompanying standard deviation, it is
+    the subject's self-reported mass. Additional trials found in the data set
+    with a subject identification number 0 are trials with no subject, i.e.
     unloaded trials that can be used for inertial compensation purposes, and
-    are not shown in the table. Generated
-    by \verb|src/subject_table.py|.}
+    are not shown in the table. Generated by \verb|src/subject_table.py|.}
   \centering
@@
   \label{tab:subjects}
-\end{table}
-%
-\begin{table}
-  \cprotect\caption{A list of unloaded trials collected for each speed. Each
-    loaded trial includes a compensation file listed in its meta data which
-    matches it to these unloaded trials. Generated by
-    \verb|src/subject_table.py|.}
-  \centering
-  \small
-  \input{tables/compensation_trials.tex}
-  \label{tab:comptrials}
 \end{table}
@@
 accelerometers were intended to provide information for inertial compensation
-purposes when the treadmill moved laterally, but are extraneous for trials
-greater than 8 due to the treadmill being stabilized.
+purposes when the treadmill moved laterally or in pitch, but are extraneous for
+trials greater than number 8 due to the treadmill being stabilized in those
+degrees of freedom.
 %
@@
 the treadmill for 10 minutes under unperturbed and perturbed conditions. Before
-a set of trials on the same day the following happened:
-%
-\begin{itemize}
-  \item Calibration of the motion capture system using the manufacturer's
-    recommended procedure.
-  \item Subject changes into athletic shoes, shorts, sports bra, baseball
-    cap, and rock climbing harness.
-  \item All 47 markers are applied directly to the skin except for the heel,
-    toe, and head markers, which were placed on the respective article of
-    clothing.~\footnote{The sacrum and rear pelvic markers may have been placed
-    on the shorts for a small number of the subjects}.
-  \item Subjects self-reported age, gender, and mass.
-  \item Height was measured by the experimentalist.
-  \item Four reference photographs (front, back, right, left) were taken of
-    subject's marker locations.
-\end{itemize}
+a set of trials on the same day, the motion capture system was calibrated using
+the manufacturer's recommended procedure. Before each subject's gait data were
+collected, the subject changed into athletic shoes, shorts, a sports bra,
+a baseball cap, and a rock climbing harness. All 47 markers were applied
+directly to the skin at the landmarks noted in Table~\ref{tab:marker-labels}
+except for the heel, toe, and head markers, which were placed on the respective
+article of clothing.~\footnote{The sacrum and rear pelvic markers may have been
+placed on the shorts for a small number of the subjects.} Then the subject
+self-reported their age, gender, and mass. Finally, their height was measured
+by the experimentalist and four reference photographs (front, back, right,
+left) were taken of subject's marker locations.
 
@@
 signals were not used in the experiments presented in this paper.} The measured
-speed of the treadmill belts are compared to these commanded treadmill control
-input signals in Figure~\ref{fig:input_output} to show the effect of the
-treadmill and controller dynamics. The system introduces a delay and seems to
-act as a low pass filter. The standard deviations of the measured speeds do not
+speed of the treadmill belts from unloaded trials (79, 80, 81) are compared to
+these commanded treadmill control input signals in
+Figure~\ref{fig:input_output} to show the effect of the treadmill and
+controller dynamics. The system introduces a delay and seems to act as a low
+pass filter. The standard deviations of the measured speeds do not
 significantly differ from the desired values:
@@
 content of the actual perturbations, spectral density plots were created
-by averaging a spectrogram over twenty seconds in the Hamming window, shown 
+by averaging a spectrogram of twenty second Hamming window, shown
 in Figure~\ref{fig:freq_analysis}. For all speeds, the frequency content of the
-commanded and measured belt speeds were noticeably similar.  
+commanded and measured time series show similarity below 4~\si{\hertz} and
+attenuation in the measured spectral density above 4~\si{\hertz}.
 %
@@
 \end{figure}
+
+When belt speed is not constant, the inertia of the rollers and motor will
+induce error in the force plate $x$ axis moment, and hence, the
+anterior-posterior coordinate ($z$ axis) of the center of pressure that is
+measured by the instrumentation in the treadmill. If needed, this error can be
+compensated by a linear model as shown by \cite{Hnat2014}. The model
+coefficients can be identified from the unloaded trials given in
+Table~\ref{tab:comptrials}. The error due to inertia is random and does not
+affect the averaged joint moments presented in
+Figure~\ref{fig:angle-torque-comparison}. Compensation should, however, be done
+if joint moments from individual gait cycles are of interest rather than the
+ensemble average.
+%
+\begin{table}
+  \cprotect\caption{A list of unloaded trials collected for each speed. Each
+    loaded trial includes a compensation file listed in its meta data which
+    matches it to these unloaded trials. Generated by
+    \verb|src/subject_table.py|.}
+  \centering
+  \small
+  \input{tables/compensation_trials.tex}
+  \label{tab:comptrials}
+\end{table}
+
 \section*{Results}
@@
 \subsubsection*{mocap-xxx.txt}
-
-The output from the D-Flow mocap module is stored in a tab delimited file named
-\verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id number. The file
-is tab delimited and contains a number of time series. The numerical values of
-the time series are provided in decimal fixed point notation with 6 decimals of
+%
+The output from the D-Flow mocap module is stored in a tab separated value
+(TSV) file named \verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id
+number. The file contains a number of time series. The numerical values of the
+time series are provided in decimal fixed point notation with 6 decimals of
 precision, e.g. \verb|123456.123456|, regardless of the units. The first line
@@
 number column, marker position columns, force plate force/moment columns, force
-plate center of pressure columns, other analog columns, and potentially results
-from the real time Human Body Model \cite{Bogert2013} which is included with
-D-Flow. The columns are further described below:
+plate center of pressure columns, and other analog columns. The columns are
+further described below:
 %
@@
   \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
-    receives a frame from Cortex. These are recorded at approximately at
-    100~\si{\hertz} and given in seconds.
+    receives a frame from Cortex. These are recorded approximately at
+    100~\si{\hertz} sampling rate and given in seconds.
   \item[FrameNumber] Monotonically increasing positive integers that correspond
@@
     force/moment calibration matrix stored in Cortex converts the voltages to
-    forces and moments before sending it to D-Flow. Cortex also computes
+    forces and moments before sending it to D-Flow. The software also computes
     the center of pressure from the forces, moments, and force plate
@@
 the D-Flow system time in seconds. This time corresponds to the time recorded
-in the \verb+TimeStamp+ column in mocap module tsv file which is necessary for
+in the \verb+TimeStamp+ column in mocap module TSV file which is necessary for
 time synchronization. There are two additional columns \verb+RightBeltSpeed+
@@
 a marker goes missing, if the data was recorded in a D-Flow version less than
-3.16.2rc4 [3], D-Flow continues to record the last non-missing value in all
+3.16.2rc4, D-Flow continues to record the last non-missing value in all
 three axes until the marker is visible again. In D-Flow versions greater than
 or equal to 3.16.2rc4, the missing markers are indicated in the TSV file as
-either \verb|0.000000| or \verb|-0.000000|, which is the same as has been in
-the HBM columns in all versions of D-Flow. The D-Flow version must be provided
-in the meta data yml file for each trial to be able to distinguish this detail.
+either \verb|0.000000| or \verb|-0.000000|. The D-Flow version must be provided
+in the meta data YAML file to be able to distinguish this detail.
 %
@@
   \centering
-  \footnotesize
+  \scriptsize
   \begin{tabular}{lrlll}
@@
 processed trial to present the nature of the data. The tool was developed in
-Python, is dependent on the SciPy Stack and Octave, and provides two main
-classes: one to do basic gait data cleaning from D-Flow's output files,
-\verb|DFlowData|, and a second to compute common gait variables of interest,
-\verb|GaitData|.
+Python, is dependent on the SciPy Stack (NumPy, SciPy, matplotlib, Pandas, etc)
+and Octave, and provides two main classes: one to do basic gait data cleaning
+from D-Flow's output files, \verb|DFlowData|, and a second to compute common
+gait variables of interest, \verb|GaitData|.
 
@@
 
-Once the data is cleaned there are two methods that allow you to extract the
-cleaned data: either extract sections of the data bounded by the events
+Once the data is cleaned there are two methods that allow the user to extract
+the cleaned data: either extract sections of the data bounded by the events
 recorded in the \verb|record-xxx.txt| file or save the cleaned data to disk.
@@
     >>> data.clean_data()
-    >>> event_df =data.extract_processed_data(
+    >>> event_df = data.extract_processed_data(
     ...     event='Longitudinal Perturbation')
@@
 
-The \verb|GaitData| class is then used to compute things such as gait events
-(toe off and heel strike times), basic 2D kinematics and inverse dynamics, and
-to store the data into a Pandas \verb|Panel| with each gait cycle on the item
-axis at a specified sample rate. This object can also be serialized to disk in
-HDF5 format. An example of using the Python API is shown in
-Listing~\ref{lis:gaitdata}.
+The \verb|GaitData| class is then used to compute gait events (toe off and heel
+strike times), basic 2D kinematics and inverse dynamics, and to store the data
+into a Pandas \verb|Panel| with each gait cycle on the item axis at a specified
+sampling rate. This object can also be serialized to disk in HDF5 format. An
+example of using the Python API is shown in Listing~\ref{lis:gaitdata}.
 %
@@
 \section*{Data Availability}
+%
 The data set, \cite{Moore2014}, is available via the Zenodo data repository.
@@
 \section*{Software Availability}
-The tables and figures in the paper can be reproduced from the source
+%
+The tables, figures, and the paper can be reproduced from the source
 repository shared on Github: https://github.com/csu-hmc/perturbed-data-paper.

### 2014-12-19 paper.tex
Grammar.

@@
 content of the actual perturbations, spectral density plots were created
-by averaging a spectrogram of twenty second Hamming window, shown
+by averaging a spectrogram of a twenty second Hamming window, shown
 in Figure~\ref{fig:freq_analysis}. For all speeds, the frequency content of the

### 2014-12-19 paper.tex
Spelling.

@@
     calibration poses were not explicitly tagged by an event, yet the protocol
-    was the identical to the following trials. The calibration pose will have
-    to be determined manually.
+    was identical to the following trials. The calibration pose will have to be
+    determined manually.
   \item  Trials 6-15 have force measurements are affected by the treadmill

### 2014-12-19 paper.tex
Cleaned up data limitations list.

@@
     collected in other studies.
-  \item Trials 37, 38, and 39 do not exist. The numbers were accidentally
-    skipped.
+  \item Trials 6-8 use a only the lower body marker set. The remaining trials
+    are full body.
+  \item Trials 6-8 use an early experimental protocol which divided the
+    perturbation sections into three sections: longitudinal perturbations,
+    lateral perturbations, and a combination of each. The calibration pose and
+    zeroing are present in the data but not identified explicitly by an event.
+  \item  Trials 6-8's force measurements are affected by the treadmill
+    vibration mode mentioned in the equipment section and the force plate data
+    should not be used. We include the trials because both the kinematic data
+    is valid and it includes lateral perturbations in addition to the
+    longitudinal.
   \item Trials 9, 10, and 11 used a slightly different event definition where the
@@
     determined manually.
-  \item  Trials 6-15 have force measurements are affected by the treadmill
-    vibration mode mentioned in the equipment section and the forces should be
-    not be used. We include the trials because both the kinematic data is valid
-    and trials 6-8 include lateral perturbations in addition to the
-    longitudinal.
-  \item During trials 9-15 we used wooden blocks to fix the treadmill to the
+  \item During trials 9-15, we used wooden blocks to fix the treadmill to the
     concrete floor to eliminate the treadmill's low vibration mode
@@
     marker data is fine.
-  \item Trials 6-8 use an early experimental protocol which divided the
-    perturbation sections into three sections: longitudinal perturbations,
-    lateral perturbations, and a combination of each. We then learned the
-    treadmill had a low vibrational mode which significantly affects the force
-    plate measurements, requiring us to eliminate the lateral perturbation
-    motions. The force measurements during these trials are corrupted by this
-    vibrational mode and should be used with caution or not at all.
   \item We did not record unloaded compensation trials for trials 9-15.
@@
     wooden blocks.
-  \item Trials 6-8 use a only the lower body marker set. The remaining trials
-    are full body.
   \item The ankle joint torques computed from subject 9's data in trials 25-27
@@
     force calibration.
+  \item Trials 37, 38, and 39 do not exist. The numbers were accidentally
+    skipped.
 \end{itemize}

### 2015-01-13 paper.tex
Added major section introductions, fixes #91.
- Added intros to the Methods and Results sections.
- Moved the markers subsection in as a paragraph in the mocap.txt section.
- Fixed figure captions to say red instead of purple.
- Got rid of last "velocity" mentions and changed to "speed".
- Moved wording to "commanded belt speed" instead of "input speed".

@@
 \section*{Methods}
+%
+In this section, we describe our experimental design beginning with
+descriptions of the participants and equipment. This is then followed by the
+protocol details and specifics on the perturbation design.
+
 \subsection*{Participants}
@@
   \includegraphics{figures/input_vs_output.pdf}
-  \caption{Treadmill belt speed input signals (purple) and recorded output
-    speeds (blue) for average belt speeds of
+  \caption{Commanded treadmill belt speed (red) and the recorded
+    speed (blue) for average belt speeds of
     \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively.}
@@
   \includegraphics{figures/frequency_analysis.pdf}
-  \caption{Frequency spectrum of the treadmill belt velocity input signal
-    (purple) and the recorded output velocity (blue) for average belt speeds of
-    0.8, 1.2, and 1.6 m/s, respectively}
+  \caption{Power spectral density of the commanded treadmill belt speed
+    (red) and the recorded speed (blue) for average belt speeds of
+    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively}
   \label{fig:freq_analysis}
@@
 \section*{Results}
+%
+Here we present some basic results. We first provide a detailed description of
+the raw data followed by an overview of several computed variables that give an
+idea of the characteristics of both the unperturbed and perturbed gait.
+
 \subsection*{Raw Data}
@@
 
+We make use of the full body 47 marker set described in \cite{Bogert2013} and
+presented in detail in Table \ref{tab:marker-labels}. As with all camera based
+motion capture systems, the markers sometimes go missing in the recording. When
+a marker goes missing, if the data was recorded in a D-Flow version less than
+3.16.2rc4, D-Flow continues to record the last non-missing value in all
+three axes until the marker is visible again. In D-Flow versions greater than
+or equal to 3.16.2rc4, the missing markers are indicated in the TSV file as
+either \verb|0.000000| or \verb|-0.000000|. The D-Flow version must be provided
+in the meta data YAML file to be able to distinguish this detail.
+%
+\begin{table*}
+  \cprotect\caption{Descriptions of the 47 markers used in this study. The
+    ``Set'' column indicates whether the marker exists in the lower and/or full
+    body marker set. The label column matches the column headers in the
+    \verb|mocap-xxx.txt| files and/or the marker map in the \verb|meta-xxx.yml|
+    file.}
+  \centering
+  \scriptsize
+  \begin{tabular}{lrlll}
+    \toprule
+    Set & \# & Label & Name & Description \\
+    \midrule
+    F   & 1  & LHEAD & Left head                             & Just above the ear, in the middle. \\
+    F   & 2  & THEAD & Top head                              & On top of the head, in line with the LHEAD and RHEAD. \\
+    F   & 3  & RHEAD & Right head                            & Just above the ear, in the middle. \\
+    F   & 4  & FHEAD & Forehead                              & Between line LHEAD/RHEAD and THEAD a bit right from center. \\
+    L/F & 5  & C7    & C7                                    & On the 7th cervical vertebrae. \\
+    L/F & 6  & T10   & T10                                   & On the 10th thoracic vertbrae. \\
+    L/F & 7  & SACR  & Sacrum bone                           & On the sacral bone. \\
+    L/F & 8  & NAVE  & Navel                                 & On the navel. \\
+    L/F & 9  & XYPH  & Xiphoid process                       & Xiphoid process of the sternum. \\
+    F   & 10 & STRN  & Sternum                               & On the jugular notch of the sternum. \\
+    F   & 11 & BBAC  & Scapula                               & On the inferior angle fo the right scapular. \\
+    F   & 12 & LSHO  & Left shoulder                         & Left acromion. \\
+    F   & 13 & LDELT & Left deltoid muscle                   & Apex of the deltoid muscle. \\
+    F   & 14 & LLEE  & Left lateral elbow                    & Left lateral epicondyle of the elbow. Upper one in the T-Pose. \\
+    F   & 15 & LMEE  & Left medial elbow                     & Left medial epicondyle of the elbow. Lower on in the T-Pose. \\
+    F   & 16 & LFRM  & Left forearm                          & On 2/3 on the line between the LLEE and LMW. \\
+    F   & 17 & LMW   & Left medial wrist                     & On styloid process radius, thumb side. \\
+    F   & 18 & LLW   & Left lateral wrist                    & On styloid process ulna, pinky side. \\
+    F   & 19 & LFIN  & Left fingers                          & Center of the hand. Caput metatarsal 3. \\
+    F   & 20 & RSHO  & Right shoulder                        & Right acromion. \\
+    F   & 21 & RDELT & Right deltoid muscle                  & Apex of deltoid muscle. \\
+    F   & 22 & RLEE  & Right lateral elbow                   & Right lateral epicondyle of the elbow. Lower one in the T-pose. \\
+    F   & 23 & RMEE  & Right medial elbow                    & Right medial epicondyle of the elbow. Lower one in the T-pose. \\
+    F   & 24 & RFRM  & Right forearm                         & On 1/3 on the line between the RLEE and RMW. \\
+    F   & 25 & RMW   & Right medial wrist                    & On styloid process radius, thumb side. \\
+    F   & 26 & RLW   & Right lateral wrist                   & On styloid process ulna, pinky side. \\
+    F   & 27 & RFIN  & Right fingers                         & Center of the hand. Caput metatarsal 3. \\
+    L/F & 28 & LASIS & Pelvic bone left front                & Left anterior superior iliac spine. \\
+    L/F & 29 & RASIS & Pelvic bone right front               & Right anterior superior iliac spine. \\
+    L/F & 30 & LPSIS & Pelvic bone left back                 & Left posterior superio iliac spine. \\
+    L/F & 31 & RPSIS & Pelvic bone right back                & Right posterior superior iliac spine. \\
+    L/F & 32 & LGTRO & Left greater trochanter of the femur  & On the cetner of the left greater trochanter. \\
+    L/F & 33 & FLTHI & Left thigh                            & On 1/3 on the line between the LFTRO and LLEK. \\
+    L/F & 34 & LLEK  & Left lateral epicondyle of the knee   & On the lateral side of the joint axis. \\
+    L/F & 35 & LATI  & Left anterior of the tibia            & On 2/3 on the line between the LLEK and LLM. \\
+    L/F & 36 & LLM   & Left lateral malleoulus of the ankle  & The center of the heel at the same height as the toe. \\
+    L/F & 37 & LHEE  & Left heel                             & Center of the heel at the same height as the toe. \\
+    L/F & 38 & LTOE  & Left toe                              & Tip of big toe. \\
+    L/F & 39 & LMT5  & Left 5th metatarsal                   & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
+    L/F & 40 & RGTRO & Right greater trochanter of the femur & On the cetner of the right greater trochanter. \\
+    L/F & 41 & FRTHI & Right thigh                           & On 2/3 on the line between the RFTRO and RLEK. \\
+    L/F & 42 & RLEK  & Right lateral epicondyle of the knee  & On the lateral side of the joint axis. \\
+    L/F & 43 & RATI  & Right anterior of the tibia           & On 1/3 on the line between the RLEK and RLM. \\
+    L/F & 44 & RLM   & Right lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
+    L/F & 45 & RHEE  & Right heel                            & Center of the heel at the same height as the toe. \\
+    L/F & 46 & RTOE  & Right toe                             & Tip of big toe. \\
+    L/F & 47 & RMT5  & Right 5th metatarsal                  & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
+    \bottomrule
+  \end{tabular}
+  \label{tab:marker-labels}
+\end{table*}
+
 \subsubsection*{record-xxx.txt}
@@
 identification number. There are three main headings in the file: \verb+study+,
-\verb+subject+, and \verb+trial+. An example meta data file is shown in Listing
-\ref{lis:example-meta-data}.
+\verb+subject+, and \verb+trial+. An example meta data file is shown in
+Listing~\ref{lis:example-meta-data}.
 
@@
 
-\subsection*{Markers}
-%
-We make use of the full body 47 marker set described in \cite{Bogert2013} and
-presented in detail in Table \ref{tab:marker-labels}. As with all camera based
-motion capture systems, the markers sometimes go missing in the recording. When
-a marker goes missing, if the data was recorded in a D-Flow version less than
-3.16.2rc4, D-Flow continues to record the last non-missing value in all
-three axes until the marker is visible again. In D-Flow versions greater than
-or equal to 3.16.2rc4, the missing markers are indicated in the TSV file as
-either \verb|0.000000| or \verb|-0.000000|. The D-Flow version must be provided
-in the meta data YAML file to be able to distinguish this detail.
-%
-\begin{table*}
-  \cprotect\caption{Descriptions of the 47 markers used in this study. The
-    ``Set'' column indicates whether the marker exists in the lower and/or full
-    body marker set. The label column matches the column headers in the
-    \verb|mocap-xxx.txt| files and/or the marker map in the \verb|meta-xxx.yml|
-    file.}
-  \centering
-  \scriptsize
-  \begin{tabular}{lrlll}
-    \toprule
-    Set & \# & Label & Name & Description \\
-    \midrule
-    F   & 1  & LHEAD & Left head                             & Just above the ear, in the middle. \\
-    F   & 2  & THEAD & Top head                              & On top of the head, in line with the LHEAD and RHEAD. \\
-    F   & 3  & RHEAD & Right head                            & Just above the ear, in the middle. \\
-    F   & 4  & FHEAD & Forehead                              & Between line LHEAD/RHEAD and THEAD a bit right from center. \\
-    L/F & 5  & C7    & C7                                    & On the 7th cervical vertebrae. \\
-    L/F & 6  & T10   & T10                                   & On the 10th thoracic vertbrae. \\
-    L/F & 7  & SACR  & Sacrum bone                           & On the sacral bone. \\
-    L/F & 8  & NAVE  & Navel                                 & On the navel. \\
-    L/F & 9  & XYPH  & Xiphoid process                       & Xiphoid process of the sternum. \\
-    F   & 10 & STRN  & Sternum                               & On the jugular notch of the sternum. \\
-    F   & 11 & BBAC  & Scapula                               & On the inferior angle fo the right scapular. \\
-    F   & 12 & LSHO  & Left shoulder                         & Left acromion. \\
-    F   & 13 & LDELT & Left deltoid muscle                   & Apex of the deltoid muscle. \\
-    F   & 14 & LLEE  & Left lateral elbow                    & Left lateral epicondyle of the elbow. Upper one in the T-Pose. \\
-    F   & 15 & LMEE  & Left medial elbow                     & Left medial epicondyle of the elbow. Lower on in the T-Pose. \\
-    F   & 16 & LFRM  & Left forearm                          & On 2/3 on the line between the LLEE and LMW. \\
-    F   & 17 & LMW   & Left medial wrist                     & On styloid process radius, thumb side. \\
-    F   & 18 & LLW   & Left lateral wrist                    & On styloid process ulna, pinky side. \\
-    F   & 19 & LFIN  & Left fingers                          & Center of the hand. Caput metatarsal 3. \\
-    F   & 20 & RSHO  & Right shoulder                        & Right acromion. \\
-    F   & 21 & RDELT & Right deltoid muscle                  & Apex of deltoid muscle. \\
-    F   & 22 & RLEE  & Right lateral elbow                   & Right lateral epicondyle of the elbow. Lower one in the T-pose. \\
-    F   & 23 & RMEE  & Right medial elbow                    & Right medial epicondyle of the elbow. Lower one in the T-pose. \\
-    F   & 24 & RFRM  & Right forearm                         & On 1/3 on the line between the RLEE and RMW. \\
-    F   & 25 & RMW   & Right medial wrist                    & On styloid process radius, thumb side. \\
-    F   & 26 & RLW   & Right lateral wrist                   & On styloid process ulna, pinky side. \\
-    F   & 27 & RFIN  & Right fingers                         & Center of the hand. Caput metatarsal 3. \\
-    L/F & 28 & LASIS & Pelvic bone left front                & Left anterior superior iliac spine. \\
-    L/F & 29 & RASIS & Pelvic bone right front               & Right anterior superior iliac spine. \\
-    L/F & 30 & LPSIS & Pelvic bone left back                 & Left posterior superio iliac spine. \\
-    L/F & 31 & RPSIS & Pelvic bone right back                & Right posterior superior iliac spine. \\
-    L/F & 32 & LGTRO & Left greater trochanter of the femur  & On the cetner of the left greater trochanter. \\
-    L/F & 33 & FLTHI & Left thigh                            & On 1/3 on the line between the LFTRO and LLEK. \\
-    L/F & 34 & LLEK  & Left lateral epicondyle of the knee   & On the lateral side of the joint axis. \\
-    L/F & 35 & LATI  & Left anterior of the tibia            & On 2/3 on the line between the LLEK and LLM. \\
-    L/F & 36 & LLM   & Left lateral malleoulus of the ankle  & The center of the heel at the same height as the toe. \\
-    L/F & 37 & LHEE  & Left heel                             & Center of the heel at the same height as the toe. \\
-    L/F & 38 & LTOE  & Left toe                              & Tip of big toe. \\
-    L/F & 39 & LMT5  & Left 5th metatarsal                   & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
-    L/F & 40 & RGTRO & Right greater trochanter of the femur & On the cetner of the right greater trochanter. \\
-    L/F & 41 & FRTHI & Right thigh                           & On 2/3 on the line between the RFTRO and RLEK. \\
-    L/F & 42 & RLEK  & Right lateral epicondyle of the knee  & On the lateral side of the joint axis. \\
-    L/F & 43 & RATI  & Right anterior of the tibia           & On 1/3 on the line between the RLEK and RLM. \\
-    L/F & 44 & RLM   & Right lateral malleoulus of the ankle & The center of the heel at the same height as the toe. \\
-    L/F & 45 & RHEE  & Right heel                            & Center of the heel at the same height as the toe. \\
-    L/F & 46 & RTOE  & Right toe                             & Tip of big toe. \\
-    L/F & 47 & RMT5  & Right 5th metatarsal                  & Caput of the 5th metatarsal bone, on joint line midfoot/toes. \\
-    \bottomrule
-  \end{tabular}
-  \label{tab:marker-labels}
-\end{table*}
-
 \subsection*{Processed Data}
@@
   \cprotect\caption{Box plots of the average belt speed, stride frequency, and
-    stride length which compare unperturbed (purple) and perturbed (blue) gait
+    stride length which compare unperturbed (red) and perturbed (blue) gait
     cycles. The median is given with the box bounding the first and third

### 2015-01-13 paper.tex
Spelling.

--- 2014-12-19 paper.tex
@@
 speed of the treadmill belts from unloaded trials (79, 80, 81) are compared to
-these commanded treadmill control input signals in
+the commanded treadmill control input signals in
 Figure~\ref{fig:input_output} to show the effect of the treadmill and

### 2015-01-13 paper.tex

@@
 processed trial to present the nature of the data. The tool was developed in
-Python, is dependent on the SciPy Stack (NumPy, SciPy, matplotlib, Pandas, etc)
-and Octave, and provides two main classes: one to do basic gait data cleaning
-from D-Flow's output files, \verb|DFlowData|, and a second to compute common
-gait variables of interest, \verb|GaitData|.
+Python, is dependent on the SciPy Stack [NumPy~\citep{Walt2011},
+SciPy~\citep{Jones2001}, matplotlib~\citep{Hunter2007},
+Pandas~\citep{McKinney2010}, etc] and Octave~\citep{Octave2014}, and provides
+two main classes: one to do basic gait data cleaning from D-Flow's output
+files, \verb|DFlowData|, and a second to compute common gait variables of
+interest, \verb|GaitData|.
 

### 2015-01-14 paper.tex
Accentuate Figures 2 and 3 in the text, fixes #92
Moved the subject of the sentence, the figures, to the beginning of the
paragraphs.

@@
 and fast walking speeds and slow and fast running speeds.~\footnote{The running
-signals were not used in the experiments presented in this paper.} The measured
-speed of the treadmill belts from unloaded trials (79, 80, 81) are compared to
-the commanded treadmill control input signals in
-Figure~\ref{fig:input_output} to show the effect of the treadmill and
-controller dynamics. The system introduces a delay and seems to act as a low
-pass filter. The standard deviations of the measured speeds do not
-significantly differ from the desired values:
-\SIlist{0.05;0.12;0.2}{\meter\per\second} for the three speeds, respectively.
+signals were not used in the experiments presented in this paper.}
+
+Figure~\ref{fig:input_output} gives an idea of the effect of the treadmill and
+controller dynamics by plotting the measured speed of the treadmill belts from
+unloaded trials (79, 80, 81) against the commanded treadmill control input
+signal. The system introduces a delay and seems to act as a low pass filter.
+The standard deviations of the measured speeds do not significantly differ from
+the desired values: \SIlist{0.05;0.12;0.2}{\meter\per\second} for the three
+speeds, respectively.
 %
@@
 
-To show the effects of the treadmill dynamics and give an idea of the frequency
-content of the actual perturbations, spectral density plots were created
-by averaging a spectrogram of a twenty second Hamming window, shown
-in Figure~\ref{fig:freq_analysis}. For all speeds, the frequency content of the
-commanded and measured time series show similarity below 4~\si{\hertz} and
-attenuation in the measured spectral density above 4~\si{\hertz}.
+Figure~\ref{fig:freq_analysis} gives a frequency domain view of the effects of
+the treadmill dynamics. These spectral density plots were created by averaging
+a spectrogram of a twenty second Hamming window. For all speeds, the frequency
+content of the commanded and measured time series show similarity below
+4~\si{\hertz} and attenuation in the measured spectral density above
+4~\si{\hertz}.
 %

### 2015-02-18 src/unperturbed_perturbed_comparison.py
Added a fourth box plot: stride width.

@@ -7,7 +7,7 @@ basic gait stats for perturbed and unperturbed gait cycles."""
 
 import os
 
-from numpy import rad2deg
+from numpy import rad2deg, hstack
 from scipy.constants import golden
 import matplotlib.pyplot as plt
 import pandas
@@ -148,12 +148,12 @@ print('Time series plot at: {}'.format(file_path))
 # length.
 print('Generating the box plot comparison.')
 
-fig, axes = plt.subplots(1, 3)
+fig, axes = plt.subplots(1, 4)
 
 columns = ['Average Belt Speed', 'Stride Frequency', 'Stride Length']
 units = ['[m/s]', '[Hz]', '[m]']
 
-for col, ax, unit in zip(columns, axes.flatten(), units):
+for col, ax, unit in zip(columns, axes[:-1], units):
     ax.set_title('{} {}'.format(col, unit))
     u = unperturbed_gait_cycle_stats[col]
     p = perturbed_gait_data.gait_cycle_stats[col]
@@ -161,6 +161,24 @@ for col, ax, unit in zip(columns, axes.flatten(), units):
                 whis='range',
                 names=['Perturbed\nN = {}'.format(num_perturbed_cycles),
                        'Unperturbed\nN = {}'.format(num_unperturbed_cycles)])
+
+perturbed_stride_width = (perturbed_gait_data.data['RHEE.PosZ'] -
+                          perturbed_gait_data.data['LHEE.PosZ'])
+
+unperturbed_stride_width_1 = (unperturbed_gait_data_1.data['RHEE.PosZ'].iloc[500:] -
+                              unperturbed_gait_data_1.data['LHEE.PosZ'].iloc[500:])
+
+unperturbed_stride_width_2 = (unperturbed_gait_data_2.data['RHEE.PosZ'].iloc[:5500] -
+                              unperturbed_gait_data_2.data['LHEE.PosZ'].iloc[:5500])
+
+sbn.boxplot([perturbed_stride_width.values,
+             hstack((unperturbed_stride_width_1.values,
+                     unperturbed_stride_width_2.values))],
+             ax=axes[-1],
+             color=[blue, red],
+             whis='range',
+             names=['Perturbed', 'Unperturbed'])
+axes[-1].set_title('Stride Width [m]'.format(col, unit))
 
 plt.tight_layout()
 

### 2015-02-18 paper.tex
Simplified stride width calc and shorted x tick labels on boxplot.

@@
 gait cycle statistics. One can see that the perturbed strides have a much
-larger variation in frequency and length.
+larger variation in frequency and length and even larger variation in stride
+width.
 %
@@
   \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
-  \cprotect\caption{Box plots of the average belt speed, stride frequency, and
-    stride length which compare unperturbed (red) and perturbed (blue) gait
-    cycles. The median is given with the box bounding the first and third
-    quartiles and the whiskers bound the range of the data. Produced by
-    \verb|src/unperturbed_perturbed_comparison.py|.}
+  % NOTE : Make sure to update these values (120, 519) if anything in the
+  % script changes.
+  \cprotect\caption{Box plots of the average belt speed, stride frequency,
+    stride length, and stride width which compare 120 unperturbed (U: red) and
+    519 perturbed (P: blue) gait cycles. The median is given with the box
+    bounding the first and third quartiles and the whiskers bound the range of
+    the data. Produced by \verb|src/unperturbed_perturbed_comparison.py|.}
   \label{fig:gait-cycle-stats-comparison}

### 2015-02-18 src/unperturbed_perturbed_comparison.py

@@ -7,7 +7,7 @@ basic gait stats for perturbed and unperturbed gait cycles."""
 
 import os
 
-from numpy import rad2deg, hstack
+from numpy import rad2deg
 from scipy.constants import golden
 import matplotlib.pyplot as plt
 import pandas
@@ -159,25 +159,20 @@ for col, ax, unit in zip(columns, axes[:-1], units):
     p = perturbed_gait_data.gait_cycle_stats[col]
     sbn.boxplot([p.values, u.values], ax=ax, color=[blue, red],
                 whis='range',
-                names=['Perturbed\nN = {}'.format(num_perturbed_cycles),
-                       'Unperturbed\nN = {}'.format(num_unperturbed_cycles)])
+                names=['P', 'U'])
 
-perturbed_stride_width = (perturbed_gait_data.data['RHEE.PosZ'] -
-                          perturbed_gait_data.data['LHEE.PosZ'])
+unperturbed_stride_widths = \
+    (unperturbed_gait_cycles[:, :, 'RHEE.PosZ'] -
+     unperturbed_gait_cycles[:, :, 'LHEE.PosZ']).values.flatten()
+perturbed_stride_widths = \
+    (perturbed_gait_data.gait_cycles[:, :, 'RHEE.PosZ'] -
+     perturbed_gait_data.gait_cycles[:, :, 'LHEE.PosZ']).values.flatten()
 
-unperturbed_stride_width_1 = (unperturbed_gait_data_1.data['RHEE.PosZ'].iloc[500:] -
-                              unperturbed_gait_data_1.data['LHEE.PosZ'].iloc[500:])
-
-unperturbed_stride_width_2 = (unperturbed_gait_data_2.data['RHEE.PosZ'].iloc[:5500] -
-                              unperturbed_gait_data_2.data['LHEE.PosZ'].iloc[:5500])
-
-sbn.boxplot([perturbed_stride_width.values,
-             hstack((unperturbed_stride_width_1.values,
-                     unperturbed_stride_width_2.values))],
-             ax=axes[-1],
-             color=[blue, red],
-             whis='range',
-             names=['Perturbed', 'Unperturbed'])
+sbn.boxplot([perturbed_stride_widths, unperturbed_stride_widths],
+            ax=axes[-1],
+            color=[blue, red],
+            whis='range',
+            names=['P', 'U'])
 axes[-1].set_title('Stride Width [m]'.format(col, unit))
 
 plt.tight_layout()

### 2015-02-18 paper.tex
Added definition of YAML in a footnote and point to the YAML listing for an example, fixes #131.

--- 2015-01-14 paper.tex
@@
 both the ``mocap'' and ``record'' modules in D-Flow in addition to a manually
-generated YAML file that contains all of the necessary meta data for the given
-trial. These three files are stored in a hierarchy of directories with one
-trial per directory. The directories are named in the following fashion
-\verb+T001/+ where \verb+T+ stands for ``trial'' and the following three digits
-are provide a unique trial identification number.
+generated YAML~\footnote{YAML is a human readable data serialization format.
+See Listing~\ref{lis:example-meta-data} for an example.} file that contains all
+of the necessary meta data for the given trial. These three files are stored
+in a hierarchy of directories with one trial per directory. The directories are
+named in the following fashion \verb+T001/+ where \verb+T+ stands for ``trial''
+and the following three digits are provide a unique trial identification
+number.
 
@@
   \end{minted}
-  \caption{A fictitious example of a YAML formatted meta data file. All of the
-    possible keys in the data set are shown.}
+  \caption{A fictitious example of a YAML formatted meta data file. Examples of
+    all of the possible keys in the data set are shown.}
   \label{lis:example-meta-data}

### 2015-02-19 paper.tex

@@
 widely disseminated, well organized, nor available with few or no restrictions.
-David Winter's published normative gait data, \cite{Winter1990}, is widely used
+David Winter's published normative gait data \citep{Winter1990} is widely used
 in biomechanical studies, yet it comes from few subjects and only a small
 number of gait cycles per subject. This small source has successfully enabled
-many other studies, such as powered prosthetic control design, \cite{Sup2008},
+many other studies, such as powered prosthetic control design \citep{Sup2008}
 but success in other research fields using large sets of data for discovery
@@
 years old but the data does not seem to be publicly available. This is,
-unfortunately, typical. Tirosh et. al, recognized the need for a comprehensive
-data base for clinical gait data and created the Gaitabase, \cite{Tirosh2010}.
-This database may contain a substantial amount of data but it is encumbered by
-a complicated and restrictive license and sharing scheme. However, there
-are examples of data with less restrictions. The University of Wisconsin at
-LaCrosse has an easily accessible normative gait data set, \cite{Willson2014},
+unfortunately, typical. \cite{Tirosh2010} recognized the need for a
+comprehensive data base for clinical gait data and created the Gaitabase. This
+database may contain a substantial amount of data but it is encumbered by a
+complicated and restrictive license and sharing scheme. However, there are
+examples of data with less restrictions. The University of Wisconsin at
+LaCrosse has an easily accessible normative gait data set \citep{Willson2014}
 from 25 subjects with lower extremity marker data from multiple gait cycles and
@@
 with rich meta data and made available in the most unrestrictive form for other
-research uses following modern best practices in data sharing, \cite{White2013}.
+research uses following modern best practices in data sharing \citep{White2013}.
 
@@
 measured by the instrumentation in the treadmill. If needed, this error can be
-compensated by a linear model as shown by \cite{Hnat2014}. The model
+compensated by a linear model as shown in \cite{Hnat2014}. The model
 coefficients can be identified from the unloaded trials given in
@@
     to the user's desired names for the markers.
-  \item[marker-set] Indicates the HBM~\cite{Bogert2013} marker set used during
+  \item[marker-set] Indicates the HBM~\citep{Bogert2013} marker set used during
     the trial, either full, lower, or NA.
@@
 %
-We developed a toolkit for data processing, GaitAnalysisToolKit v0.1.2,
-\cite{Moore2014a}, for common gait computations and provide an example
+We developed a toolkit for data processing, GaitAnalysisToolKit v0.1.2
+\citep{Moore2014a} for common gait computations and provide an example
 processed trial to present the nature of the data. The tool was developed in
@@
 %
-The data set, \cite{Moore2014}, is available via the Zenodo data repository.
+The data set \citep{Moore2014} is available via the Zenodo data repository.
 Two approximately 1.2GB gzipped tar balls contain the data and a \verb|README|
@@
 Along with the source code in the repository, the computations depend on
-version 0.1.2 of the GaitAnalysisToolKit, \cite{Moore2014a}, which can be
+version 0.1.2 of the GaitAnalysisToolKit \citep{Moore2014a} which can be
 downloaded from Zenodo or the Python Package Index (http://pypi.python.org).

### 2015-02-19 paper.tex

@@
 subjects from \cite{Vaughan1992} is available on the site. At another website,
-the CGA Normative Gait Database, \cite{Kirtley2014} shares normative gait data
-from several studies and these files have influenced other studies, for example
-the average gait cycles from children used in \cite{Bogert2003}.
+the CGA Normative Gait Database, \cite{Kirtley2014} curates and shares
+normative clinical gait data collected from multiple labs and these datasets
+have influenced other studies, for example the average gait cycles from
+children used in \cite{Bogert2003}.
 

### 2015-02-19 src/unperturbed_perturbed_comparison.py
Switch orderd of U and P in box plots plus raised the titles a bit.

@@ -154,12 +154,12 @@ columns = ['Average Belt Speed', 'Stride Frequency', 'Stride Length']
 units = ['[m/s]', '[Hz]', '[m]']
 
 for col, ax, unit in zip(columns, axes[:-1], units):
-    ax.set_title('{} {}'.format(col, unit))
+    ax.set_title('{} {}'.format(col, unit), y=1.08)
     u = unperturbed_gait_cycle_stats[col]
     p = perturbed_gait_data.gait_cycle_stats[col]
-    sbn.boxplot([p.values, u.values], ax=ax, color=[blue, red],
+    sbn.boxplot([u.values, p.values], ax=ax, color=[red, blue],
                 whis='range',
-                names=['P', 'U'])
+                names=['U', 'P'])
 
 unperturbed_stride_widths = \
     (unperturbed_gait_cycles[:, :, 'RHEE.PosZ'] -
@@ -168,12 +168,12 @@ perturbed_stride_widths = \
     (perturbed_gait_data.gait_cycles[:, :, 'RHEE.PosZ'] -
      perturbed_gait_data.gait_cycles[:, :, 'LHEE.PosZ']).values.flatten()
 
-sbn.boxplot([perturbed_stride_widths, unperturbed_stride_widths],
+sbn.boxplot([unperturbed_stride_widths, perturbed_stride_widths],
             ax=axes[-1],
-            color=[blue, red],
+            color=[red, blue],
             whis='range',
-            names=['P', 'U'])
-axes[-1].set_title('Stride Width [m]'.format(col, unit))
+            names=['U', 'P'])
+axes[-1].set_title('Stride Width [m]'.format(col, unit), y=1.08)
 
 plt.tight_layout()
 

### 2015-03-04 paper.tex
treadmill > treadmill base, fixes #127.

@@
 
-The acceleration of the treadmill was measured during each trial by four
+The acceleration of the treadmill base was measured during each trial by four
 ADXL330 accelerometers placed at the four corners of the machine. These
@@
   \item Once notified by the video display, the subject stood in the
-    initialization pose: standing straight up, looking forward, arms out by
-    their sides (~45 degrees) and the event, ``Calibration Pose'', was manually
-    recorded by the operator.
+    calibration pose: standing straight up, looking forward, arms out by their
+    sides (approximately 45 degree abduction) and the event, ``Calibration
+    Pose'', was manually recorded by the operator.
   \item A countdown to the first normal walking phase was displayed. At the end
@@
     F   & 13 & LDELT & Left deltoid muscle                   & Apex of the deltoid muscle. \\
-    F   & 14 & LLEE  & Left lateral elbow                    & Left lateral epicondyle of the elbow. Upper one in the T-Pose. \\
-    F   & 15 & LMEE  & Left medial elbow                     & Left medial epicondyle of the elbow. Lower on in the T-Pose. \\
+    F   & 14 & LLEE  & Left lateral elbow                    & Left lateral epicondyle of the elbow. \\
+    F   & 15 & LMEE  & Left medial elbow                     & Left medial epicondyle of the elbow. \\
     F   & 16 & LFRM  & Left forearm                          & On 2/3 on the line between the LLEE and LMW. \\
@@
     F   & 21 & RDELT & Right deltoid muscle                  & Apex of deltoid muscle. \\
-    F   & 22 & RLEE  & Right lateral elbow                   & Right lateral epicondyle of the elbow. Lower one in the T-pose. \\
-    F   & 23 & RMEE  & Right medial elbow                    & Right medial epicondyle of the elbow. Lower one in the T-pose. \\
+    F   & 22 & RLEE  & Right lateral elbow                   & Right lateral epicondyle of the elbow. \\
+    F   & 23 & RMEE  & Right medial elbow                    & Right medial epicondyle of the elbow. \\
     F   & 24 & RFRM  & Right forearm                         & On 1/3 on the line between the RLEE and RMW. \\
@@
     torques from both unperturbed (red) and perturbed (blue) gait cycles
-    from trial 20. Produced by \verb|src/unperturbed_perturbed_comparison.py|.}
+    from trial 20. We define the nominal configuration, i.e. all joint angles
+    equal to zero, such that the vectors from the shoulder to the hip, the hip
+    to the knee, the knee to the ankle, and the heel to the toe are all aligned.
+    Produced by \verb|src/unperturbed_perturbed_comparison.py|.}
   \label{fig:angle-torque-comparison}

### 2015-03-10 paper.tex
Footnote about baseball cap. Fixes 122.

@@
 collected, the subject changed into athletic shoes, shorts, a sports bra,
-a baseball cap, and a rock climbing harness. All 47 markers were applied
-directly to the skin at the landmarks noted in Table~\ref{tab:marker-labels}
-except for the heel, toe, and head markers, which were placed on the respective
-article of clothing.~\footnote{The sacrum and rear pelvic markers may have been
-placed on the shorts for a small number of the subjects.} Then the subject
-self-reported their age, gender, and mass. Finally, their height was measured
-by the experimentalist and four reference photographs (front, back, right,
-left) were taken of subject's marker locations.
+a baseball cap~\footnote{A cap was used to avoid having to shave participants'
+hair but at the expense of accuracy.}, and a rock climbing harness. All 47 markers
+were applied directly to the skin at the landmarks noted in
+Table~\ref{tab:marker-labels} except for the heel, toe, and head markers, which
+were placed on the respective article of clothing.~\footnote{The sacrum and
+rear pelvic markers may have been placed on the shorts for a small number of
+the subjects.} Then the subject self-reported their age, gender, and mass.
+Finally, their height was measured by the experimentalist and four reference
+photographs (front, back, right, left) were taken of subject's marker
+locations.
 

### 2015-03-10 paper.tex

--- 2015-03-04 paper.tex
@@
     collected in other studies.
+  \item Subject 1 walked only at a single speed with three trial repetitions.
+    This is because these trials were originally for protocol testing purposed
+    but the kinematic data is valid and the trials uniquely have lateral
+    perturbations, so we have included them. See the next items for more
+    details about these trials.
   \item Trials 6-8 use a only the lower body marker set. The remaining trials

### 2015-03-11 paper.tex
Added note about stride width increasing with long perturbations. Fixes #129.
Added note on the x moment error with variable belt speed.

@@
 When belt speed is not constant, the inertia of the rollers and motor will
-induce error in the force plate $x$ axis moment, and hence, the
+likely induce error in the force plate $x$ axis moment, and hence, the
 anterior-posterior coordinate ($z$ axis) of the center of pressure that is
-measured by the instrumentation in the treadmill. If needed, this error can be
+measured by the instrumentation in the treadmill. This error may or may not be
+pertinent to different analyses. If needed, this error can be partially
 compensated by a linear model as shown in \cite{Hnat2014}. The model

### 2015-03-11 paper.tex

--- 2015-03-10 paper.tex
@@
 widely disseminated, well organized, nor available with few or no restrictions.
-David Winter's published normative gait data \citep{Winter1990} is widely used
+David Winter's published normative gait data~\citep{Winter1990} is widely used
 in biomechanical studies, yet it comes from few subjects and only a small
 number of gait cycles per subject. This small source has successfully enabled
-many other studies, such as powered prosthetic control design \citep{Sup2008}
+many other studies, such as powered prosthetic control design~\citep{Sup2008}
 but success in other research fields using large sets of data for discovery
-lead one to believe that more elaborate data sets may benefit the field of
+lead us to believe that more elaborate data sets may benefit the field of
 human motion studies. To enable such work, biomechanical data needs to be
@@
 
-Our use case for the data is centered around the need of bio-inspired control
-systems for emerging powered prosthetics and orthotics. Ideally, a powered
+But our reasons are not entirely altruistic, as governments and granting
+agencies are also encouraging researchers to share data with recent policy
+changes. For example, the \cite{EuropeanCommission2012} has outlined publicly
+funded data's role in innovation and the \cite{WhiteHouse2013} laid out a plan
+for public access to publications and data in 2013. The National Science
+Foundation, which partially funds this work, was ahead of the White House and
+required all grants to include a data management plan in 2011. This work is a
+partial fulfillment of the grant requirements laid out in our grant's data
+management plan and we hope that this work can be a good model for
+dissemination of biomechanical data.
+
+Our use case for the data is centered around the need for bio-inspired control
+systems in emerging powered prosthetics and orthotics. Ideally, a powered
 prosthetic would behave in such a way that the user would feel like their limb
@@
 mechanisms used during a human's natural gait and recovery from perturbations.
-We have collected data that is richer than previous gait data sets and may be
-rich enough for control identification. The data can also be used for
-verification purposes for controllers that have been designed in other manners.
+We hypothesize that by forcing the human to recover from external
+perturbations, the resulting reactive actions can be used along with system
+identification techniques to expose the feedback related relationships among
+the human's sensors and actuators. We have collected data that is richer than
+previous gait data sets and may be rich enough for control identification. The
+data can also be used for verification purposes for controllers that have been
+designed in other manners, such as those constructed from first principles
+\citep{Geyer2010}.
 

### 2015-03-11 paper.tex

@@
 larger variation in frequency and length and even larger variation in stride
-width.
+width. It is also interesting to note that the coupled nature of system's
+degrees of freedom can be exploited to increase the stride width with only
+longitudinal perturbations, although not relatively as much as is in the other
+statistics.
 %

### 2015-03-11 paper.tex
Added more gait database citations.
Fixes #126.

@@
 database may contain a substantial amount of data but it is encumbered by a
-complicated and restrictive license and sharing scheme. However, there are
+complicated and restrictive license and sharing scheme. \cite{Yun2014} provides
+lower body kinematic data of single gait cycles from over one hundred subjects
+extracted from the large KIST Human Gait Pattern Data database which may also
+include a substantial amount of raw data but it is private. However, there are
 examples of data with less restrictions. The University of Wisconsin at
@@
 from 25 subjects with lower extremity marker data from multiple gait cycles and
-force plate measurements from a single gait cycle.
+force plate measurements from a single gait cycle. The CMU Graphics Lab
+Motion Capture Database~\citep{CMU2015} is also a good example and contains
+full body marker kinematics for a fair number of trials with small number of
+gait cycles during both walking and running.
 
@@
 are present. The second is notable because it publishes the data in Dryad, a
-modern citable data repository.
+modern citable data repository. It is also worth noting purely visual data
+collections of gait, like the one presented in \cite{Makihara2012}, which
+contain videos of subjects walking on a treadmill in full clothing. This
+database is also unfortunately tightly secured with an extensive release
+agreement for reuse.
 

### 2015-03-11 src/frequency_analysis.m
Initial commit addressing some MATLAB issues
#69 - MATLAB dependencies
#117 - Frequency Spectrum Plot Units
#110 - Parse Events when Loading Files
Also renamed MATLAB scripts

@@ -4,10 +4,34 @@ clc
 %-------------------------------------------------------------------------
 %Loading Files
 %-------------------------------------------------------------------------
-    input_signals=load('Longitudinal_Perturbation_2.4.2014.txt');
-    output_slow=load('record-079_0001.txt');
-    output_normal=load('record-080_0001.txt');
-    output_fast=load('record-081_0001.txt');
+    input_signals = importdata('Longitudinal_Perturbation_2.4.2014.txt');
+    output_slow_file = fopen('record-076_0001.txt');
+    output_normal_file = fopen('record-077_0001.txt');
+    output_fast_file = fopen('record-078_0001.txt');
+%-------------------------------------------------------------------------
+%Parse Events from Files
+%-------------------------------------------------------------------------
+    %Remove Lines Starting with #
+        output_slow_parse = textscan(output_slow_file,'%s', 'Delimiter', '\n', 'CommentStyle', '#');
+        output_slow_parse = output_slow_parse{:};
+        output_normal_parse = textscan(output_normal_file,'%s', 'Delimiter', '\n', 'CommentStyle', '#');
+        output_normal_parse = output_normal_parse{:};
+        output_fast_parse = textscan(output_fast_file,'%s', 'Delimiter', '\n', 'CommentStyle', '#');
+        output_fast_parse = output_fast_parse{:};
+    %Preallocating
+        output_slow = zeros(length(output_slow_parse)-1,4);
+        output_normal = zeros(length(output_normal_parse)-1,4);
+        output_fast = zeros(length(output_fast_parse)-1,4);
+    %Recreate Matrices 
+        for i = 2:length(output_slow_parse)
+            output_slow(i-1,:) = str2num(output_slow_parse{i,:});
+        end
+        for i = 2:length(output_normal_parse)
+            output_normal(i-1,:) = str2num(output_normal_parse{i,:});
+        end
+        for i = 2:length(output_fast_parse)
+            output_fast(i-1,:) = str2num(output_fast_parse{i,:});
+        end
 %-------------------------------------------------------------------------
 %Declaring Variables
 %-------------------------------------------------------------------------
@@ -65,7 +89,7 @@ clc
                 semilogy(f_input,p_input_mean,'Color',[0.815 0.3294 0.3020])
                 hold on
                 semilogy(f_output,p_output_mean,'Color',[0.2157 0.4706 0.7490])
-                ylabel('PSD (v^2/Hz)')
[…]

### 2015-03-11 src/input_output_plot.m

@@ -4,26 +4,26 @@ clc
 %-------------------------------------------------------------------------
 %Loading Files
 %-------------------------------------------------------------------------
-    input_signals=load('Longitudinal_Perturbation_2.4.2014.txt');
-    output_slow=load('record-079_0001.txt');
-    output_normal=load('record-080_0001.txt');
-    output_fast=load('record-081_0001.txt');
+    input_signals = importdata('Longitudinal_Perturbation_2.4.2014.txt');
+    output_slow = importdata('record-076.txt');
+    output_normal = importdata('record-077.txt');
+    output_fast = importdata('record-078.txt');
 %-------------------------------------------------------------------------
 %Declaring Variables
 %-------------------------------------------------------------------------
     %Time Variables (necessary because Record Module has varying timestamp)
-        time_input=input_signals(:,1);
-        output_slow_time=output_slow(:,1); 
-        output_slow_time=output_slow_time-output_slow_time(1,1);
-        output_normal_time=output_normal(:,1);
-        output_normal_time=output_normal_time-output_normal_time(1,1);
-        output_fast_time=output_fast(:,1);
-        output_fast_time=output_fast_time-output_fast_time(1,1);
+        time_input = input_signals(:,1);
+        output_slow_time = output_slow.data(:,1); 
+        output_slow_time = output_slow_time-output_slow_time(1,1);
+        output_normal_time = output_normal.data(:,1);
+        output_normal_time = output_normal_time-output_normal_time(1,1);
+        output_fast_time = output_fast.data(:,1);
+        output_fast_time  =output_fast_time-output_fast_time(1,1);
     %Speed
         inputs=[input_signals(:,2) input_signals(:,3) input_signals(:,4)];
-        output_slow=output_slow(:,2); 
-        output_normal=output_normal(:,2);
-        output_fast=output_fast(:,2);
+        output_slow = output_slow.data(:,2); 
+        output_normal = output_normal.data(:,2);
+        output_fast = output_fast.data(:,2);
 %-------------------------------------------------------------------------
 %Plotting
 %-------------------------------------------------------------------------

### 2015-03-11 src/frequency_analysis.m
Update input vs output plot using loaded trials

@@ -5,9 +5,9 @@ clc
 %Loading Files
 %-------------------------------------------------------------------------
     input_signals = importdata('Longitudinal_Perturbation_2.4.2014.txt');
-    output_slow_file = fopen('record-076_0001.txt');
-    output_normal_file = fopen('record-077_0001.txt');
-    output_fast_file = fopen('record-078_0001.txt');
+    output_slow_file = fopen('record-076.txt');
+    output_normal_file = fopen('record-077.txt');
+    output_fast_file = fopen('record-078.txt');
 %-------------------------------------------------------------------------
 %Parse Events from Files
 %-------------------------------------------------------------------------

### 2015-03-11 paper.tex
#128 Lateral perturbations
added plot, source code, and brief description in text

--- 2015-03-11 paper.tex
@@
 widely disseminated, well organized, nor available with few or no restrictions.
-David Winter's published normative gait data~\citep{Winter1990} is widely used
+David Winter's published normative gait data \citep{Winter1990} is widely used
 in biomechanical studies, yet it comes from few subjects and only a small
 number of gait cycles per subject. This small source has successfully enabled
-many other studies, such as powered prosthetic control design~\citep{Sup2008}
+many other studies, such as powered prosthetic control design \citep{Sup2008}
 but success in other research fields using large sets of data for discovery
-lead us to believe that more elaborate data sets may benefit the field of
+lead one to believe that more elaborate data sets may benefit the field of
 human motion studies. To enable such work, biomechanical data needs to be
@@
 
-But our reasons are not entirely altruistic, as governments and granting
-agencies are also encouraging researchers to share data with recent policy
-changes. For example, the \cite{EuropeanCommission2012} has outlined publicly
-funded data's role in innovation and the \cite{WhiteHouse2013} laid out a plan
-for public access to publications and data in 2013. The National Science
-Foundation, which partially funds this work, was ahead of the White House and
-required all grants to include a data management plan in 2011. This work is a
-partial fulfillment of the grant requirements laid out in our grant's data
-management plan and we hope that this work can be a good model for
-dissemination of biomechanical data.
-
-Our use case for the data is centered around the need for bio-inspired control
-systems in emerging powered prosthetics and orthotics. Ideally, a powered
+Our use case for the data is centered around the need of bio-inspired control
+systems for emerging powered prosthetics and orthotics. Ideally, a powered
 prosthetic would behave in such a way that the user would feel like their limb
@@
 mechanisms used during a human's natural gait and recovery from perturbations.
-We hypothesize that by forcing the human to recover from external
-perturbations, the resulting reactive actions can be used along with system
-identification techniques to expose the feedback related relationships among
-the human's sensors and actuators. We have collected data that is richer than
-previous gait data sets and may be rich enough for control identification. The
-data can also be used for verification purposes for controllers that have been
-designed in other manners, such as those constructed from first principles
-\citep{Geyer2010}.
+We have collected data that is richer than previous gait data sets and may be
+rich enough for control identification. The data can also be used for
+verification purposes for controllers that have been designed in other manners.
 
@@
 controller dynamics by plotting the measured speed of the treadmill belts from
-unloaded trials (79, 80, 81) against the commanded treadmill control input
+loaded trials (76, 77, 78) against the commanded treadmill control input
 signal. The system introduces a delay and seems to act as a low pass filter.
@@
 \end{table}
+
+In addition to the longitudinal perturbations, lateral perturbations were 
+also prescribed for a duration of four minutes for one subject in earlier 
+trials (6,7,8).  Figure~\ref{fig:lateral_perturbation} shows the effect of the
+lateral control of the treadmill by plotting the commanded and measured treadmill
+positions. These signals were generated in a similar manner using MATLAB and 
+Simulink in which a Gaussian white noise block that twice integrated to obtain the 
+lateral position.  The signal was then high-pass filtered with a second-order
+Butterworth filter to eliminate drift and then saturated so that the signal 
+remained within the 5cm ~\si{\centimeter}lateral  range of the physical hardware.  
+The same perturbation signal was used for each of the three trials.
 

### 2015-03-11 src/lateral_perturbation_plot.m

@@ -1,0 +1,40 @@
+clear
+clc
+
+%-------------------------------------------------------------------------
+%Loading, Parsing, and Truncating
+%-------------------------------------------------------------------------
+    perturbation_input = importdata('Lateral_Perturbation.txt');
+    perturbation_output = importdata('mocap-006.txt');
+    frame=1:1:length(perturbation_input);            
+    time_input=(frame-1)*0.0033;                    
+    signal_input = perturbation_input(:,2);
+    time_output = perturbation_output.data(33090:57336,1);
+    time_output = time_output - time_output(1,1);
+    signal_output = perturbation_output.data(33090:57336,3);
+    signal_output = signal_output - signal_output(1,1);
+    signal_output = interp1(time_output,signal_output,time_input);
+%-------------------------------------------------------------------------
+%Plotting 
+%-------------------------------------------------------------------------
+    %Settings to Improve Graph Appearance
+        ti = get(gca,'TightInset');
+        set(gca,'Position',[ti(1) ti(2) 1-ti(3)-ti(1) 1-ti(4)-ti(2)]);
+        set(gca,'units','centimeters')
+        pos = get(gca,'Position');
+        ti = get(gca,'TightInset');
+        set(gcf, 'PaperUnits','centimeters');
+        set(gcf, 'PaperSize', [pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
+        set(gcf, 'PaperPositionMode', 'manual');
+        set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]); 
+    %Plot
+        plot(signal_input,'Color',[0.815 0.3294 0.3020])
+        hold on
+        plot(signal_output,'Color',[0.2157 0.4706 0.7490])
+        xlabel('Time (s)')
+        ylabel('Lateral Position (m)')
+        title('Lateral Perturbation Signal')
+        legend('Commanded','Measured')
+    %Saving
+        h=figure(1);
+        saveas(h,'lateral_perturbation.pdf');

### 2015-03-11 paper.tex
Added the new figure in the tex file and removed some whitespace.

@@
 
-In addition to the longitudinal perturbations, lateral perturbations were 
-also prescribed for a duration of four minutes for one subject in earlier 
-trials (6,7,8).  Figure~\ref{fig:lateral_perturbation} shows the effect of the
-lateral control of the treadmill by plotting the commanded and measured treadmill
+In addition to the longitudinal perturbations, lateral perturbations were also
+prescribed for a duration of four minutes for one subject in earlier trials
+(6,7,8). Figure~\ref{fig:lateral_perturbation} shows the effect of the lateral
+control of the treadmill by plotting the commanded and measured treadmill
 positions. These signals were generated in a similar manner using MATLAB and
-Simulink in which a Gaussian white noise block that twice integrated to obtain the 
-lateral position.  The signal was then high-pass filtered with a second-order
-Butterworth filter to eliminate drift and then saturated so that the signal 
-remained within the 5cm ~\si{\centimeter}lateral  range of the physical hardware.  
-The same perturbation signal was used for each of the three trials.
+Simulink in which a Gaussian white noise block that twice integrated to obtain
+the lateral position. The signal was then high-pass filtered with a
+second-order Butterworth filter to eliminate drift and then saturated so that
+the signal remained within the 5~\si{\centi\meter} lateral range of the
+physical hardware.  The same perturbation signal was used for each of the three
+trials.
+%
+\begin{figure}
+  \centering
+  \includegraphics{figures/lateral_perturbation.pdf}
+  \caption{Commanded treadmill lateral deviation (red) and the recorded
+    deviation (blue) for the trials 6, 7, and 8.}
+  \label{fig:lateral_perturbation}
+\end{figure}
 

### 2015-03-11 src/lateral_perturbation_plot.m
fixed X-axis label and scale

@@ -28,13 +28,14 @@ clc
         set(gcf, 'PaperPositionMode', 'manual');
         set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]); 
     %Plot
-        plot(signal_input,'Color',[0.815 0.3294 0.3020])
+        plot(time_input, signal_input,'Color',[0.815 0.3294 0.3020])
         hold on
-        plot(signal_output,'Color',[0.2157 0.4706 0.7490])
+        plot(time_input,signal_output,'Color',[0.2157 0.4706 0.7490])
         xlabel('Time (s)')
         ylabel('Lateral Position (m)')
         title('Lateral Perturbation Signal')
         legend('Commanded','Measured')
+        xlim([100 200])
     %Saving
         h=figure(1);
         saveas(h,'lateral_perturbation.pdf');

### 2015-03-11 paper.tex

--- 2015-03-11 paper.tex
@@
 \end{table}
+
+In addition to the longitudinal perturbations, lateral perturbations were also
+prescribed for a duration of four minutes for one subject in earlier trials
+(6,7,8). Figure~\ref{fig:lateral_perturbation} shows the effect of the lateral
+control of the treadmill by plotting the commanded and measured treadmill
+positions. These signals were generated in a similar manner using MATLAB and
+Simulink in which a Gaussian white noise block that twice integrated to obtain
+the lateral position. The signal was then high-pass filtered with a
+second-order Butterworth filter to eliminate drift and then saturated so that
+the signal remained within the 5~\si{\centi\meter} lateral range of the
+physical hardware.  The same perturbation signal was used for each of the three
+trials.
+%
+\begin{figure}
+  \centering
+  \includegraphics{figures/lateral_perturbation.pdf}
+  \caption{Commanded treadmill lateral deviation (red) and the recorded
+    deviation (blue) for the trials 6, 7, and 8.}
+  \label{fig:lateral_perturbation}
+\end{figure}
 

### 2015-03-11 paper.tex

@@
 \begin{listing}
-  \begin{minted}[gobble=4]{python}
+  \begin{minted}{pycon}
 >>> from gaitanalysis.motek import DFlowData
@@
 \begin{listing}
-  \begin{minted}[gobble=4]{python}
+  \begin{minted}{pycon}
 >>> from gaitanalysis.gait import GaitData

### 2015-03-13 paper.tex
Spelling.

@@
   \item Subject 1 walked only at a single speed with three trial repetitions.
-    This is because these trials were originally for protocol testing purposed
+    This is because these trials were originally for protocol testing purposes
     but the kinematic data is valid and the trials uniquely have lateral

### 2015-03-15 paper.tex
Added commanded data zenodo citation and source file names in captions for m files.

@@
 mean velocities of \SIlist{0.8;1.2;1.6}{\meter\per\second}, were pre-generated
-with MATLAB and Simulink (Mathworks, Natick, Massachusetts, USA). The same
-control signal was used for all trials at that given speed.
+with MATLAB and Simulink (Mathworks, Natick, Massachusetts, USA) and are
+available for download from Zenodo~\citep{Hnat2015}. The same control signal
+was used for all trials at that given speed.
 
@@
   \includegraphics{figures/input_vs_output.pdf}
-  \caption{Commanded treadmill belt speed (red) and the recorded
-    speed (blue) for average belt speeds of
-    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively.}
+  \cprotect\caption{Commanded treadmill belt speed (red) and the recorded speed
+    (blue) for average belt speeds of \SIlist{0.8;1.2;1.6}{\meter\per\second},
+    respectively. Generated by \verb|src/input_output_plot.m|.}
   \label{fig:input_output}
@@
   \includegraphics{figures/frequency_analysis.pdf}
-  \caption{Power spectral density of the commanded treadmill belt speed
-    (red) and the recorded speed (blue) for average belt speeds of
-    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively}
+  \cprotect\caption{Power spectral density of the commanded treadmill belt
+    speed (red) and the recorded speed (blue) for average belt speeds of
+    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively. Generated by
+    \verb|src/frequency_analysis.m|.}
   \label{fig:freq_analysis}
@@
   \includegraphics{figures/lateral_perturbation.pdf}
-  \caption{Commanded treadmill lateral deviation (red) and the recorded
-    deviation (blue) for the trials 6, 7, and 8.}
+  \cprotect\caption{Commanded treadmill lateral deviation (red) and the
+    recorded deviation (blue) for the trials 6, 7, and 8. Generated by
+    \verb|src/lateral_perturbation_plot.m|.}
   \label{fig:lateral_perturbation}

### 2015-03-16 paper.tex
Address's most of Ton's final comments. Other's will be addressed in subsequent commits/PRs.

@@
 normative clinical gait data collected from multiple labs and these datasets
-have influenced other studies, for example the average gait cycles from
-children used in \cite{Bogert2003}.
+have influenced other studies, for example \cite{Bogert2003} made use of the
+average gait cycles from the child subjects.
 
@@
 identification techniques to expose the feedback related relationships among
-the human's sensors and actuators. We have collected data that is richer than
-previous gait data sets and may be rich enough for control identification. The
-data can also be used for verification purposes for controllers that have been
-designed in other manners, such as those constructed from first principles
-\citep{Geyer2010}.
-
-With all of this in mind, we collected over seven and half hours of gait data
+the human's sensors and actuators. With this in mind, we have collected data
+that is richer than previous gait data sets and may be rich enough for control
+identification. The data can also be used for verification purposes for
+controllers that have been designed in other manners, such as those constructed
+from first principles \citep{Geyer2010}.
+
+With all of this in mind, we collected over seven and a half hours of gait data
 from fifteen able bodied subjects which amounts to over 25,000 gait cycles. The
@@
 collected full body marker locations and ground reaction loads from a pair of
-force plates. The protocol for the majority of the trials included two minutes
-of normal walking and eight minutes of walking under the influence of
+force plates. The final protocol for the majority of the trials included two
+minutes of normal walking and eight minutes of walking under the influence of
 pseudo-random belt speed fluctuations. The data has been organized complete
 with rich meta data and made available in the most unrestrictive form for other
-research uses following modern best practices in data sharing \citep{White2013}.
+research uses following modern best practices in data sharing
+\citep{White2013}.
 
@@
 computationally reproducible and instructions are included in the associated
-repository for doing so.
+repository~\footnote{https://github.com/csu-hmc/perturbed-data-paper} for
+reproducing the results.
 
@@
   \item A R-Mill treadmill which has dual 6 degree of freedom force plates,
-    independent belts for each foot, and lateral/pitch motion capabilities
-    (Forcelink, Culemborg, Netherlands).
+    independent belts for each foot, along with lateral translation and pitch
+    rotation capabilities (Forcelink, Culemborg, Netherlands).
   \item A 10 Osprey camera motion capture system paired with the Cortex
@@
 The Cortex software delivers high accuracy 3D marker trajectories from the
-cameras along with data from force plates and analog sensors
-(EMG/Accelerometer) through a National Instruments USB-6255 data acquisition
-unit. D-Flow is required to collect data from any digital sensors and to
-control the treadmill's motion (lateral, pitch, and belts). D-Flow can process
-the data in real time and/or export data to file.
+cameras along with data from the force plates and analog sensors (e.g.
+EMG/Accelerometer) through a National Instruments USB-6255 data acquisition
+unit. D-Flow then receives streaming data from Cortex and any other digital
+sensors. It is also responsible for controlling the treadmill's motion
+(lateral, pitch, belts). D-Flow can process the data in real time and/or export
+data to file.
 
@@
 Early on, we discovered that the factory setup of the R-Link treadmill had a
-vibration mode as low as 5\si{\hertz} that is detectable in the force measurements,
-likely due to the flexible undercarriage and pitch motion mechanism. Trials 6-8
-are affected by this vibration mode. During trials 9-15 the treadmill was
-stabilized with wooden blocks. During the remaining trials the treadmill was
-stabilized with metal supports. See the Data Limitations Section for more
-details.
+vibration mode as low as 5\si{\hertz} that was detectable in the force
+measurements, likely due to the flexible undercarriage and pitch motion
+mechanism. Trials 6-8 are affected by this vibration mode. During trials 9-15
+the treadmill was stabilized with wooden blocks. During the remaining trials
+the treadmill was stabilized with metal supports; both in-house fabricated ones
+and vendor supplied ones. These supports aimed to improve the stiffness and
+frequency response of the force plate system. See the Data Limitations Section
+for more details.
 
@@
 the manufacturer's recommended procedure. Before each subject's gait data were
-collected, the subject changed into athletic shoes, shorts, a sports bra,
-a baseball cap~\footnote{A cap was used to avoid having to shave participants'
-hair but at the expense of accuracy.}, and a rock climbing harness. All 47 markers
-were applied directly to the skin at the landmarks noted in
+collected, the subject changed into athletic shoes, shorts, a sports bra, a
+baseball cap~\footnote{A cap was used to avoid having to shave participants'
+hair but at the expense of accuracy.}, and a rock climbing harness. All 47
+markers were applied directly to the skin at the landmarks noted in
 Table~\ref{tab:marker-labels} except for the heel, toe, and head markers, which
 were placed on the respective article of clothing.~\footnote{The sacrum and
-rear pelvic markers may have been placed on the shorts for a small number of
-the subjects.} Then the subject self-reported their age, gender, and mass.
+rear pelvic markers were placed on the shorts for a small number of the
+subjects.} Then the subject self-reported their age, gender, and mass.
 Finally, their height was measured by the experimentalist and four reference
@@
 locations.
-
-After obtaining informed consent and a briefing by the experimentalist on the
-trial protocol, the subject followed the verbal instructions of the
-experimentalist and the on-screen instructions from the video display. The
-protocol for a single trial was as follows:
-%
-\begin{enumerate}
-  \item Subject stepped onto the treadmill and markers were identified with
-    Cortex.
-  \item The safety rope was attached loosely to the rock climbing harness such
-    that no undue forces were acting on the subject during walking, but that
-    the harness would prevent a full fall.
-  \item The subject started by stepping on sides of treadmill so that feet did
-    not touch the force plates and the force plate signals are zeroed. This
-    corresponds to the ``Force Plate Zeroing'' event.
-  \item Once notified by the video display, the subject stood in the
-    calibration pose: standing straight up, looking forward, arms out by their
-    sides (approximately 45 degree abduction) and the event, ``Calibration
-    Pose'', was manually recorded by the operator.
-  \item A countdown to the first normal walking phase was displayed. At the end
-    of the countdown the event ``First Normal Walking'' was recorded and the
-    treadmill ramped up to the specified speed and the subject was instructed
-    to walk normally, to focus on the ``endless'' road on the display, and not
-    to look at their feet.
-  \item After 1 minute of normal walking, the longitudinal perturbation phase
-    begun and was recorded as ``Longitudinal Perturbation''.
-  \item After 8 minutes of walking under the influence of the perturbations,
-    the second normal walking phase begun and was recorded as ``Second Normal
-    Walking''.
-  \item After 1 minute of normal walking, a countdown was shown on the display
-    and the treadmill decelerated to a stop.
-  \item The subject was instructed to step off of the force plates for 10
-    seconds and the ``Unloaded End'' event was recorded.
-  \item The subject could then take a rest break before each additional trial.
-\end{enumerate}
-
-Trials 6-8 included a calibration pose at the start of the trial but the event
-was not explicitly recorded. In those trials, the ``TreadmillPerturbation''
-event marks the beginning of longitudinal perturbations and the ``Both'' event
-marks the beginning of combined longitudinal and lateral perturbations. The
-force plate zeroing at the end was also not explicitly recorded.
-
-\subsection*{Perturbation Signals}
-%
-As previously described, the protocol included a phase of normal walking,
-followed by longitudinal belt speed perturbations, and ended with a second
-segment of normal walking. Three pseudo-random belt speed control signals, with
-mean velocities of \SIlist{0.8;1.2;1.6}{\meter\per\second}, were pre-generated
-with MATLAB and Simulink (Mathworks, Natick, Massachusetts, USA) and are
-available for download from Zenodo~\citep{Hnat2015}. The same control signal
-was used for all trials at that given speed.
-
-To create the signals, we started by generating random 100~\si{\hertz}
-acceleration signals using the Simulink discrete-time Gaussian white noise
-block followed by a saturation block set at the maximum belt acceleration of
-15~\si{\meter\per\second\squared}. The signal was then integrated to obtain
-belt speed and high-pass filtered with a second-order Butterworth filter to
-eliminate drift. One of the three mean speeds were then added to the signal and
-limited between \SIrange{0}{3.6}{\meter\per\second}. The cutoff frequencies of
-the high-pass filter, as well as the variance in the acceleration signal, were
-manually adjusted until acceptable standard deviations for each mean speed were
-obtained: \SIlist{0.06;0.12;0.21}{\meter\per\second} for the three speeds,
-respectively. These ensured that the test subjects were sufficiently perturbed
-at each speed, while remaining within the limits of our equipment and testing
-protocol.
-
-To ensure that the treadmill belts could accelerate to the desired values, the
-high performance mode in the D-Flow software was enabled. This had the side
-effect of enabling too rapid of accelerations when the belt speed changed to or
-from zero speed. To eliminate this, a suitable ramped acceleration and
-deceleration were generated for the speed transitions.
-
-The MATLAB script and Simulink model produce a comma-delimited text file of
-five desired belt speed signals indexed by the time stamp. There are slow, normal,
-and fast walking speeds and slow and fast running speeds.~\footnote{The running
-signals were not used in the experiments presented in this paper.}
-
-Figure~\ref{fig:input_output} gives an idea of the effect of the treadmill and
-controller dynamics by plotting the measured speed of the treadmill belts from
-loaded trials (76, 77, 78) against the commanded treadmill control input
-signal. The system introduces a delay and seems to act as a low pass filter.
-The standard deviations of the measured speeds do not significantly differ from
-the desired values: \SIlist{0.05;0.12;0.2}{\meter\per\second} for the three
-speeds, respectively.
-%
-\begin{figure}
-  \centering
-  \includegraphics{figures/input_vs_output.pdf}
-  \cprotect\caption{Commanded treadmill belt speed (red) and the recorded speed
-    (blue) for average belt speeds of \SIlist{0.8;1.2;1.6}{\meter\per\second},
-    respectively. Generated by \verb|src/input_output_plot.m|.}
-  \label{fig:input_output}
-\end{figure}
-
-Figure~\ref{fig:freq_analysis} gives a frequency domain view of the effects of
-the treadmill dynamics. These spectral density plots were created by averaging
-a spectrogram of a twenty second Hamming window. For all speeds, the frequency
-content of the commanded and measured time series show similarity below
-4~\si{\hertz} and attenuation in the measured spectral density above
-4~\si{\hertz}.
-%
-\begin{figure}
-  \centering
-  \includegraphics{figures/frequency_analysis.pdf}
-  \cprotect\caption{Power spectral density of the commanded treadmill belt
-    speed (red) and the recorded speed (blue) for average belt speeds of
-    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively. Generated by
-    \verb|src/frequency_analysis.m|.}
-  \label{fig:freq_analysis}
-\end{figure}
-
-When belt speed is not constant, the inertia of the rollers and motor will
-likely induce error in the force plate $x$ axis moment, and hence, the
-anterior-posterior coordinate ($z$ axis) of the center of pressure that is
-measured by the instrumentation in the treadmill. This error may or may not be
-pertinent to different analyses. If needed, this error can be partially
-compensated by a linear model as shown in \cite{Hnat2014}. The model
-coefficients can be identified from the unloaded trials given in
-Table~\ref{tab:comptrials}. The error due to inertia is random and does not
-affect the averaged joint moments presented in
-Figure~\ref{fig:angle-torque-comparison}. Compensation should, however, be done
-if joint moments from individual gait cycles are of interest rather than the
-ensemble average.
-%
-\begin{table}
-  \cprotect\caption{A list of unloaded trials collected for each speed. Each
-    loaded trial includes a compensation file listed in its meta data which
-    matches it to these unloaded trials. Generated by
-    \verb|src/subject_table.py|.}
-  \centering
-  \small
-  \input{tables/compensation_trials.tex}
-  \label{tab:comptrials}
-\end{table}
-
-In addition to the longitudinal perturbations, lateral perturbations were also
-prescribed for a duration of four minutes for one subject in earlier trials
-(6,7,8). Figure~\ref{fig:lateral_perturbation} shows the effect of the lateral
-control of the treadmill by plotting the commanded and measured treadmill
-positions. These signals were generated in a similar manner using MATLAB and
-Simulink in which a Gaussian white noise block that twice integrated to obtain
-the lateral position. The signal was then high-pass filtered with a
-second-order Butterworth filter to eliminate drift and then saturated so that
-the signal remained within the 5~\si{\centi\meter} lateral range of the
-physical hardware.  The same perturbation signal was used for each of the three
-trials.
-%
-\begin{figure}
-  \centering
-  \includegraphics{figures/lateral_perturbation.pdf}
-  \cprotect\caption{Commanded treadmill lateral deviation (red) and the
-    recorded deviation (blue) for the trials 6, 7, and 8. Generated by
-    \verb|src/lateral_perturbation_plot.m|.}
-  \label{fig:lateral_perturbation}
-\end{figure}
-
-\section*{Results}
-%
-Here we present some basic results. We first provide a detailed description of
-the raw data followed by an overview of several computed variables that give an
-idea of the characteristics of both the unperturbed and perturbed gait.
-
-\subsection*{Raw Data}
-%
-The raw data consists of a set of ASCII tab delimited text files output from
-both the ``mocap'' and ``record'' modules in D-Flow in addition to a manually
-generated YAML~\footnote{YAML is a human readable data serialization format.
-See Listing~\ref{lis:example-meta-data} for an example.} file that contains all
-of the necessary meta data for the given trial. These three files are stored
-in a hierarchy of directories with one trial per directory. The directories are
-named in the following fashion \verb+T001/+ where \verb+T+ stands for ``trial''
-and the following three digits are provide a unique trial identification
-number.
-
-\subsubsection*{mocap-xxx.txt}
-%
-The output from the D-Flow mocap module is stored in a tab separated value
-(TSV) file named \verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id
-number. The file contains a number of time series. The numerical values of the
-time series are provided in decimal fixed point notation with 6 decimals of
-precision, e.g. \verb|123456.123456|, regardless of the units. The first line
-of the file holds the header. The header includes time stamp column, frame
-number column, marker position columns, force plate force/moment columns, force
-plate center of pressure columns, and other analog columns. The columns are
-further described below:
-%
-\begin{description}
-  \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
-    receives a frame from Cortex. These are recorded approximately at
-    100~\si{\hertz} sampling rate and given in seconds.
-  \item[FrameNumber] Monotonically increasing positive integers that correspond
-    to each frame received from Cortex.
-  \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
-    or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
-    reference frame. The prefixes match the marker labels given in
-    Table~\ref{tab:marker-labels}. These values are in meters.
-  \item[Ground Reaction Loads] There are three ground reaction forces and three
-    ground reaction moments recorded by each of the two force plates in Newtons
-    and Newton-Meters, respectively. The prefix for these columns is either
-    \verb+FP1+ or \verb+FP2+ and represents either force plate 1 (left) or 2
-    (right). The suffixes are either \verb+.For[XYZ]+, \verb+.Mom[XYZ]+ for the
-    forces and moments, respectively. The force plate voltages are sampled at a
-    much higher frequency than the cameras, but delivered at the Cortex camera
-    sample rate, \~100~\si{\hertz} through the D-Flow mocap module. A
-    force/moment calibration matrix stored in Cortex converts the voltages to
-    forces and moments before sending it to D-Flow. The software also computes
-    the center of pressure from the forces, moments, and force plate
-    dimensions. These have the same prefixes for the plate number, have the
-    suffix \verb+.Cop[XYZ]+, and are given in meters.
-  \item[Analog Channels] Several analog signals are recorded under column
-    headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
-    sampled by Cortex and correspond to the 96 analog channels in the National
-    Instruments USB-6255. The first twelve are the voltages from the
-    force plate load cells. We also record the acceleration of 4 points on the
-    treadmill base in analog channels 61-72 that were in place in case inertial
-    compensation for the lateral treadmill movement was required.
-\end{description}
-
-We make use of the full body 47 marker set described in \cite{Bogert2013} and
-presented in detail in Table \ref{tab:marker-labels}. As with all camera based
-motion capture systems, the markers sometimes go missing in the recording. When
-a marker goes missing, if the data was recorded in a D-Flow version less than
-3.16.2rc4, D-Flow continues to record the last non-missing value in all
-three axes until the marker is visible again. In D-Flow versions greater than
-or equal to 3.16.2rc4, the missing markers are indicated in the TSV file as
-either \verb|0.000000| or \verb|-0.000000|. The D-Flow version must be provided
-in the meta data YAML file to be able to distinguish this detail.
 %
@@
 \end{table*}
+
+After obtaining informed consent and a briefing by the experimentalist on the
+trial protocol, the subject followed the verbal instructions of the
+experimentalist and the on-screen instructions from the video display. The
+protocol for a single trial was as follows:
+%
+\begin{enumerate}
+  \item Subject stepped onto the treadmill and markers were identified with
+    Cortex.
+  \item The safety rope was attached loosely to the rock climbing harness such
+    that no forces were acting on the subject during walking, but that the
+    harness would prevent a full fall.
+  \item The subject started by stepping on sides of treadmill so that feet did
+    not touch the force plates and the force plate signals are zeroed. This
+    corresponds to the ``Force Plate Zeroing'' event.
+  \item Once notified by the video display, the subject stood in the
+    calibration pose: standing straight up, looking forward, arms out by their
+    sides (approximately 45 degree abduction) and the event, ``Calibration
+    Pose'', was manually recorded by the operator.
+  \item A countdown to the first normal walking phase was displayed. At the end
+    of the countdown the event ``First Normal Walking'' was recorded and the
+    treadmill ramped up to the specified speed and the subject was instructed
+    to walk normally, to focus on the ``endless'' road on the display, and not
+    to look at their feet.
+  \item After 1 minute of normal walking, the longitudinal perturbation phase
+    begun and was recorded as ``Longitudinal Perturbation''.
+  \item After 8 minutes of walking under the influence of the perturbations,
+    the second normal walking phase begun and was recorded as ``Second Normal
+    Walking''.
+  \item After 1 minute of normal walking, a countdown was shown on the display
+    and the treadmill decelerated to a stop.
+  \item The subject was instructed to step off of the force plates for 10
+    seconds and the ``Unloaded End'' event was recorded.
+  \item The subject could then take a rest break before each additional trial.
+\end{enumerate}
+
+Trials 6-8 included a calibration pose at the start of the trial but the event
+was not explicitly recorded. In those trials, the ``TreadmillPerturbation''
+event marks the beginning of longitudinal perturbations and the ``Both'' event
+marks the beginning of combined longitudinal and lateral perturbations. The
+force plate zeroing at the end was also not explicitly recorded.
+
+\subsection*{Perturbation Signals}
+%
+As previously described, the protocol included a phase of normal walking,
+followed by longitudinal belt speed perturbations, and ended with a second
+segment of normal walking. Three pseudo-random belt speed control signals, with
+mean velocities of \SIlist{0.8;1.2;1.6}{\meter\per\second}, were pre-generated
+with MATLAB and Simulink (Mathworks, Natick, Massachusetts, USA) and are
+available for download from Zenodo~\citep{Hnat2015}. The same control signal
+was used for all trials at that given speed.
+
+To create the signals, we started by generating random acceleration signals,
+sampled at 100~\si{\hertz}, using the Simulink discrete-time Gaussian white
+noise block followed by a saturation block set at the maximum belt acceleration
+of 15~\si{\meter\per\second\squared}. The signal was then integrated to obtain
+belt speed and high-pass filtered with a second-order Butterworth filter to
+eliminate drift. One of the three mean speeds were then added to the signal and
+limited between \SIrange{0}{3.6}{\meter\per\second}. The cutoff frequencies of
+the high-pass filter, as well as the variance in the acceleration signal, were
+manually adjusted until acceptable standard deviations for each mean speed were
+obtained: \SIlist{0.06;0.12;0.21}{\meter\per\second} for the three speeds,
+respectively. These ensured that the test subjects were sufficiently perturbed
+at each speed, while remaining within the limits of our equipment and testing
+protocol. To ensure that the treadmill belts could accelerate to the desired
+values, the high performance mode in the D-Flow software was enabled. The
+MATLAB script and Simulink model produce a comma-delimited text file of with
+the desired belt speed signals indexed by the time stamp.
+
+Figure~\ref{fig:input_output} gives an idea of the effect of the treadmill and
+controller dynamics by plotting the measured speed of the treadmill belts from
+loaded trials (76, 77, 78) against the commanded treadmill control input
+signal. The system introduces a delay and seems to act as a low pass filter.
+The standard deviations of the measured speeds do not significantly differ from
+those of the commanded speeds: \SIlist{0.05;0.12;0.2}{\meter\per\second} for
+the three speeds, respectively.
+%
+\begin{figure}
+  \centering
+  \includegraphics{figures/input_vs_output.pdf}
+  \cprotect\caption{Commanded treadmill belt speed (red) and the recorded speed
+    (blue) for average belt speeds of \SIlist{0.8;1.2;1.6}{\meter\per\second},
+    respectively. Generated by \verb|src/input_output_plot.m|.}
+  \label{fig:input_output}
+\end{figure}
+
+Figure~\ref{fig:freq_analysis} gives a frequency domain view of the effects of
+the treadmill dynamics. These spectral density plots were created by averaging
+a spectrogram of a twenty second Hamming window. For all speeds, the frequency
+content of the commanded and measured time series show similarity below
+4~\si{\hertz} and attenuation in the measured spectral density above
+4~\si{\hertz}.
+%
+\begin{figure}
+  \centering
+  \includegraphics{figures/frequency_analysis.pdf}
+  \cprotect\caption{Power spectral density of the commanded treadmill belt
+    speed (red) and the recorded speed (blue) for average belt speeds of
+    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively. Generated by
+    \verb|src/frequency_analysis.m|.}
+  \label{fig:freq_analysis}
+\end{figure}
+
+When belt speed is not constant, the inertia of the rollers and motor will
+likely induce error in the force plate $x$ axis moment, and hence, the
+anterior-posterior coordinate ($z$ axis) of the center of pressure that is
+measured by the instrumentation in the treadmill. This error may or may not be
+pertinent to different analyses. If needed, this error can be partially
+compensated by a linear model as shown in \cite{Hnat2014}. The model
+coefficients can be identified from the unloaded trials given in
+Table~\ref{tab:comptrials}. The error due to inertia is random and does not
+affect the averaged joint moments presented in
+Figure~\ref{fig:angle-torque-comparison}. Compensation should, however, be done
+if joint moments from individual gait cycles are of interest rather than the
+ensemble average.
+%
+\begin{table}
+  \cprotect\caption{A list of unloaded trials collected for each speed. Each
+    loaded trial includes a compensation file listed in its meta data which
+    matches it to these unloaded trials. Generated by
+    \verb|src/subject_table.py|.}
+  \centering
+  \small
+  \input{tables/compensation_trials.tex}
+  \label{tab:comptrials}
+\end{table}
+
+In addition to the longitudinal perturbations, lateral perturbations were also
+prescribed for a duration of four minutes for one subject in earlier trials
+(6,7,8). Figure~\ref{fig:lateral_perturbation} shows the effect of the lateral
+control of the treadmill by plotting the commanded and measured treadmill
+positions. These signals were generated in a similar manner using MATLAB and
+Simulink in which a Gaussian white noise block that twice integrated to obtain
+the lateral position. The signal was then high-pass filtered with a
+second-order Butterworth filter to eliminate drift and then saturated so that
+the signal remained within the 5~\si{\centi\meter} lateral range of the
+physical hardware.  The same perturbation signal was used for each of the three
+trials.
+%
+\begin{figure}
+  \centering
+  \includegraphics{figures/lateral_perturbation.pdf}
+  \cprotect\caption{Commanded treadmill lateral deviation (red) and the
+    recorded deviation (blue) for the trials 6, 7, and 8. Generated by
+    \verb|src/lateral_perturbation_plot.m|.}
+  \label{fig:lateral_perturbation}
+\end{figure}
+
+\section*{Results}
+%
+Here we present some basic results. We first provide a detailed description of
+the raw data followed by an overview of several computed variables that give an
+idea of the characteristics of both the unperturbed and perturbed gait.
+
+\subsection*{Raw Data}
+%
+The raw data consists of a set of ASCII tab delimited text files output from
+both the ``mocap'' and ``record'' modules in D-Flow in addition to a manually
+generated YAML~\footnote{YAML is a human readable data serialization format.
+See Listing~\ref{lis:example-meta-data} for an example.} file that contains all
+of the necessary meta data for the given trial. These three files are stored
+in a hierarchy of directories with one trial per directory. The directories are
+named in the following fashion \verb+T001/+ where \verb+T+ stands for ``trial''
+and the following three digits are provide a unique trial identification
+number.
+
+\subsubsection*{mocap-xxx.txt}
+%
+The output from the D-Flow mocap module is stored in a tab separated value
+(TSV) file named \verb+mocap-xxx.txt+ where \verb+xxx+ represents the trial id
+number. The file contains a number of time series. The numerical values of the
+time series are provided in decimal fixed point notation with 6 decimals of
+precision, e.g. \verb|123456.123456|, regardless of the units. The first line
+of the file holds the header. The header includes time stamp column, frame
+number column, marker position columns, force plate force/moment columns, force
+plate center of pressure columns, and other analog columns. The columns are
+further described below:
+%
+\begin{description}
+  \item[TimeStamp] The monotonically increasing computer clock time when D-Flow
+    receives a frame from Cortex. These are recorded approximately at
+    100~\si{\hertz} sampling rate and given in seconds.
+  \item[FrameNumber] Monotonically increasing positive integers that correspond
+    to each frame received from Cortex.
+  \item[Marker Coordinates] Any column that ends in \verb+.PosX+, \verb+.PosY+,
+    or \verb+.PosZ+ are marker coordinates expressed in Cortex's Cartesian
+    reference frame. The prefixes match the marker labels given in
+    Table~\ref{tab:marker-labels}. These values are in meters.
+  \item[Ground Reaction Loads] There are three ground reaction forces and three
+    ground reaction moments recorded by each of the two force plates in Newtons
+    and Newton-Meters, respectively. The prefix for these columns is either
+    \verb+FP1+ or \verb+FP2+ and represents either force plate 1 (left) or 2
+    (right). The suffixes are either \verb+.For[XYZ]+, \verb+.Mom[XYZ]+ for the
+    forces and moments, respectively. The force plate voltages are sampled at a
+    much higher frequency than the cameras, but delivered at the Cortex camera
+    sample rate, \~100~\si{\hertz} through the D-Flow mocap module. A
+    force/moment calibration matrix stored in Cortex converts the voltages to
+    forces and moments before sending it to D-Flow. The software also computes
+    the center of pressure from the forces, moments, and force plate
+    dimensions. These have the same prefixes for the plate number, have the
+    suffix \verb+.Cop[XYZ]+, and are given in meters.
+  \item[Analog Channels] Several analog signals are recorded under column
+    headers \verb+Channel[1-99].Anlg+. These correspond to analog signals
+    sampled by Cortex and correspond to the 96 analog channels in the National
+    Instruments USB-6255. The first twelve are the voltages from the
+    force plate load cells. We also record the acceleration of 4 points on the
+    treadmill base in analog channels 61-72 that were in place in case inertial
+    compensation for the lateral treadmill movement was required.
+\end{description}
+
+We make use of the full body 47 marker set described in \cite{Bogert2013} and
+presented in detail in Table \ref{tab:marker-labels}. As with all camera based
+motion capture systems, the markers sometimes go missing in the recording. When
+a marker goes missing, if the data was recorded in a D-Flow version less than
+3.16.2rc4, D-Flow continues to record the last non-missing value in all
+three axes until the marker is visible again. In D-Flow versions greater than
+or equal to 3.16.2rc4, the missing markers are indicated in the TSV file as
+either \verb|0.000000| or \verb|-0.000000|. The D-Flow version must be provided
+in the meta data YAML file to be able to distinguish this detail.
 

### 2015-03-17 src/lateral_perturbation_plot.m
Started to update the lateral io plot.

@@ -1,13 +1,18 @@
 clear
 clc
 
+PATHS = get_config.m;
+
 %-------------------------------------------------------------------------
 %Loading, Parsing, and Truncating
 %-------------------------------------------------------------------------
-    perturbation_input = importdata('Lateral_Perturbation.txt');
-    perturbation_output = importdata('mocap-006.txt');
-    frame=1:1:length(perturbation_input);            
-    time_input=(frame-1)*0.0033;                    
+    perturb_sig_dir = [PATHS.raw_data_dir filesep 'perturbation-signals'];
+    lat_per_file = [perturb_sig_dir filesep 'lateral-perturbation.txt'];
+    perturbation_input = importdata(lat_per_file);
+    mocap_file = [PATHS.raw_data filesep 'T006' filesep 'mocap-006.txt'];
+    perturbation_output = importdata(mocap_file);
+    frame=1:1:length(perturbation_input);
+    time_input = (frame-1)*0.0033;
     signal_input = perturbation_input(:,2);
     time_output = perturbation_output.data(33090:57336,1);
     time_output = time_output - time_output(1,1);
@@ -15,7 +20,7 @@ clc
     signal_output = signal_output - signal_output(1,1);
     signal_output = interp1(time_output,signal_output,time_input);
 %-------------------------------------------------------------------------
-%Plotting 
+%Plotting
 %-------------------------------------------------------------------------
     %Settings to Improve Graph Appearance
         ti = get(gca,'TightInset');
@@ -26,7 +31,7 @@ clc
         set(gcf, 'PaperUnits','centimeters');
         set(gcf, 'PaperSize', [pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
         set(gcf, 'PaperPositionMode', 'manual');
-        set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]); 
+        set(gcf, 'PaperPosition',[0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
     %Plot
         plot(time_input, signal_input,'Color',[0.815 0.3294 0.3020])
         hold on

### 2015-03-17 src/subject_table.py
Sorted subject table by trial # and added rules to group pilot subjects.

@@ -94,12 +94,16 @@ if not os.path.exists(table_dir):
     os.makedirs(table_dir)
 
 table_path = os.path.join(table_dir, 'subjects.tex')
-tex = unique_subjects.sort().to_latex(na_rep='NA', index=False,
-                                      columns=new_cols,
-                                      formatters=formatters)
+tex = unique_subjects.sort('1.6 m/s', na_position='first').to_latex(
+    na_rep='NA', index=False, columns=new_cols, formatters=formatters)
+
 tex = tex.replace('dollar', '$')
 tex = tex.replace('plusminus', '\pm')
 tex = tex.replace('rlrrllll', 'rlrrrrrr')
+lines = tex.split('\n')
+lines.insert(5, r'\midrule[0.25pt]')
+lines.insert(9, r'\midrule[0.25pt]')
+tex = '\n'.join(lines)
 with open(table_path, 'w') as f:
     f.write(tex)
 print('Table at: {}'.format(table_path))

### 2015-03-17 paper.tex
Made the distinction between pilot trials and the final protocol much more apparent.

@@
 % - No tabs, just spaces!
-% - No hanging whitespace
-% - Wrap lines to 79 characters
-% - Indentation should be two spaces per indent
+% - No hanging whitespace.
+% - Wrap lines to 79 characters.
+% - Indentation should be two spaces per indent.
 % - Only one space after periods.
@@
 \begin{table}
-  \cprotect\caption{Information about the 15 study participants. The final
-    three columns provide the trial numbers associated with each nominal
-    treadmill speed. The measured mass is computed from the mean total vertical
-    ground reaction force just after the calibration pose event, if possible.
-    If the mass is reported without an accompanying standard deviation, it is
-    the subject's self-reported mass. Additional trials found in the data set
-    with a subject identification number 0 are trials with no subject, i.e.
-    unloaded trials that can be used for inertial compensation purposes, and
-    are not shown in the table. Generated by \verb|src/subject_table.py|.}
+  \cprotect\caption{Information about the 15 study participants in order of
+    collection date. The subjects are divided into those that were used for the
+    protocol pilot trials, i.e. the first three, and those used for the final
+    protocol. The final three columns provide the trial numbers associated with
+    each nominal treadmill speed. The measured mass is computed from the mean
+    total vertical ground reaction force just after the calibration pose event,
+    if possible.  If the mass is reported without an accompanying standard
+    deviation, it is the subject's self-reported mass.  Additional trials found
+    in the data set with a subject identification number 0 are trials with no
+    subject, i.e.  unloaded trials that can be used for inertial compensation
+    purposes, and are not shown in the table. Generated by
+    \verb|src/subject_table.py|.}
   \centering
@@
 measurements, likely due to the flexible undercarriage and pitch motion
-mechanism. Trials 6-8 are affected by this vibration mode. During trials 9-15
+mechanism. Trials 6--8 are affected by this vibration mode. During trials 9--15
 the treadmill was stabilized with wooden blocks. During the remaining trials
-the treadmill was stabilized with metal supports; both in-house fabricated ones
-and vendor supplied ones. These supports aimed to improve the stiffness and
-frequency response of the force plate system. See the Data Limitations Section
-for more details.
-
-The acceleration of the treadmill base was measured during each trial by four
+($>$~15) the treadmill was stabilized with metal supports; both ones we
+fabricated in-house and ones supplied by the vendor. These supports aimed to
+improve the stiffness and frequency response of the force plate system. See the
+Data Limitations Section for more details.
+
+The acceleration of the treadmill base was measured during each trial by the
 ADXL330 accelerometers placed at the four corners of the machine. These
@@
 trials greater than number 8 due to the treadmill being stabilized in those
-degrees of freedom.
+degrees of freedom by the aforementioned supports.
 %
@@
 baseball cap~\footnote{A cap was used to avoid having to shave participants'
-hair but at the expense of accuracy.}, and a rock climbing harness. All 47
-markers were applied directly to the skin at the landmarks noted in
+hair at the expense of accuracy.}, and a rock climbing harness. All 47 markers
+were applied directly to the skin at the landmarks noted in
 Table~\ref{tab:marker-labels} except for the heel, toe, and head markers, which
@@
 \begin{table*}
-  \cprotect\caption{Descriptions of the 47 markers used in this study. The
-    ``Set'' column indicates whether the marker exists in the lower and/or full
-    body marker set. The label column matches the column headers in the
+  \cprotect\caption{Descriptions of the 47 subject markers used in this study.
+    The ``Set'' column indicates whether the marker exists in the lower and/or
+    full body marker set. The label column matches the column headers in the
     \verb|mocap-xxx.txt| files and/or the marker map in the \verb|meta-xxx.yml|
@@
 experimentalist and the on-screen instructions from the video display. The
-protocol for a single trial was as follows:
+final protocol for a single trial was as follows:
 %
@@
   \item The safety rope was attached loosely to the rock climbing harness such
-    that no forces were acting on the subject during walking, but that the
+    that no forces were acting on the subject during walking, but so that the
     harness would prevent a full fall.
@@
 
-Trials 6-8 included a calibration pose at the start of the trial but the event
-was not explicitly recorded. In those trials, the ``TreadmillPerturbation''
-event marks the beginning of longitudinal perturbations and the ``Both'' event
-marks the beginning of combined longitudinal and lateral perturbations. The
-force plate zeroing at the end was also not explicitly recorded.
+\subsubsection*{Pilot Protocols}
+%
+Trials 3--15 were pilot tests for finalizing the protocol design an thus have
+some slight variations with respect to the subsequent trials. We include these
+trials due to the uniqueness of trials 6--8 and the fact that the kinematic
+data is valid. We believe there may be useful analyses that only require the
+kinematic data. Additional information needed to interpret the data in the
+pilot trials can be found in the associated meta data files and the Data
+Limitations Section of this paper.
+
+Trials 6--8 use an early experimental protocol which divided the walking period
+into three sections: no perturbation, longitudinal perturbation, and a
+combination of longitudinal and lateral perturbation. The calibration pose and
+zeroing events are present in the data but lumped into one event. These trials
+only use the lower body marker set described in Table~\ref{tab:marker-labels}.
+Additionally, there are five markers that have labels beginning with \verb|ROT|
+that were attached to the treadmill base to capture the lateral motion. Trials
+9--15 use the final protocol but have corrupt ground reaction loads due to the
+wooden treadmill base stabilizers.
 
@@
 In addition to the longitudinal perturbations, lateral perturbations were also
-prescribed for a duration of four minutes for one subject in earlier trials
-(6,7,8). Figure~\ref{fig:lateral_perturbation} shows the effect of the lateral
-control of the treadmill by plotting the commanded and measured treadmill
-positions. These signals were generated in a similar manner using MATLAB and
-Simulink in which a Gaussian white noise block that twice integrated to obtain
-the lateral position. The signal was then high-pass filtered with a
-second-order Butterworth filter to eliminate drift and then saturated so that
-the signal remained within the 5~\si{\centi\meter} lateral range of the
-physical hardware.  The same perturbation signal was used for each of the three
-trials.
+prescribed for a duration of four minutes in the pilot trials 3--8.
+Figure~\ref{fig:lateral_perturbation} shows the effect of the lateral control
+of the treadmill by plotting the commanded and measured treadmill positions.
+These signals were generated in a similar manner using MATLAB and Simulink in
+which a Gaussian white noise block that twice integrated to obtain the lateral
+position. The signal was then high-pass filtered with a second-order
+Butterworth filter to eliminate drift and then saturated so that the signal
+remained within the 5~\si{\centi\meter} lateral range of the physical hardware.
+The same perturbation signal was used for each of the three trials.
 %
@@
     much higher frequency than the cameras, but delivered at the Cortex camera
-    sample rate, \~100~\si{\hertz} through the D-Flow mocap module. A
-    force/moment calibration matrix stored in Cortex converts the voltages to
-    forces and moments before sending it to D-Flow. The software also computes
-    the center of pressure from the forces, moments, and force plate
-    dimensions. These have the same prefixes for the plate number, have the
-    suffix \verb+.Cop[XYZ]+, and are given in meters.
+    sample rate, approximately 100~\si{\hertz}, through the D-Flow mocap
+    module. A force/moment calibration matrix stored in Cortex converts the
+    voltages to forces and moments before sending it to D-Flow. The software
+    also computes the center of pressure from the forces, moments, and force
+    plate dimensions. These have the same prefixes for the plate number, have
+    the suffix \verb+.Cop[XYZ]+, and are given in meters.
   \item[Analog Channels] Several analog signals are recorded under column
@@
     force plate load cells. We also record the acceleration of 4 points on the
-    treadmill base in analog channels 61-72 that were in place in case inertial
-    compensation for the lateral treadmill movement was required.
+    treadmill base in analog channels 61--72 that were in place in case
+    inertial compensation for the lateral treadmill movement was required.
 \end{description}
@@
     the meta data to the events detected in the record module file.
-  \item Interially compensate the ground reaction loads based on whether the
+  \item Inertially compensate the ground reaction loads based on whether the
     meta data indicates there was treadmill motion.
@@
 all data there are anomalies that may affect the use and interpretation of
-results emanating from the data. The following list gives various notes and
-warnings about the data that should be taken into account when making use of
-it.
+results emanating from the data. The following lists give various notes and
+warnings about the data that should be taken into account before use.
+
+\subsubsection*{All Trials}
 %
@@
     mentioned.
-  \item The subject identification number 0 stands for "no subject" and was
-    used whenever data was collected from the system with no subject on the
+  \item The subject identification number 0 represents the ``null subject'' and
+    was used whenever data was collected from the system with no subject on the
     treadmill, for example during the trials that were intended to be used for
@@
     identification numbers were reserved for early data exploration from data
-    collected in other studies.
+    collected in other studies and are not included in this dataset.
+  \item Trials 37, 38, and 39 do not exist. The numbers were accidentally
+    skipped.
+  \item The ankle joint torques computed from subject 9's data in trials 25--27
+    are abnormal and should be used with caution or not at all. We were not
+    able to locate the source of the error, but it is likely related to the
+    force calibration.
+\end{itemize}
+
+\subsubsection*{Pilot Trials}
+%
+\begin{itemize}
   \item Subject 1 walked only at a single speed with three trial repetitions.
-    This is because these trials were originally for protocol testing purposes
-    but the kinematic data is valid and the trials uniquely have lateral
-    perturbations, so we have included them. See the next items for more
-    details about these trials.
-  \item Trials 6-8 use a only the lower body marker set. The remaining trials
-    are full body.
-  \item Trials 6-8 use an early experimental protocol which divided the
-    perturbation sections into three sections: longitudinal perturbations,
-    lateral perturbations, and a combination of each. The calibration pose and
-    zeroing are present in the data but not identified explicitly by an event.
-  \item  Trials 6-8's force measurements are affected by the treadmill
-    vibration mode mentioned in the equipment section and the force plate data
-    should not be used. We include the trials because both the kinematic data
-    is valid and it includes lateral perturbations in addition to the
-    longitudinal.
-  \item Trials 9, 10, and 11 used a slightly different event definition where the
+  \item Trials 6--8 included a calibration pose at the start of the trial but
+    the event was not explicitly recorded. In those trials, the
+    ``TreadmillPerturbation'' event marks the beginning of longitudinal
+    perturbations and the ``Both'' event marks the beginning of combined
+    longitudinal and lateral perturbations. The force plate zeroing at the end
+    was also not explicitly recorded.
+  \item Trials 6--8's force measurements are affected by the treadmill vibration
+    mode mentioned in the equipment section and the force plate data should not
+    be used.
+  \item Trials 9--11 used a slightly different event definition where the
     calibration poses were not explicitly tagged by an event, yet the protocol
@@
     determined manually.
-  \item During trials 9-15, we used wooden blocks to fix the treadmill to the
-    concrete floor to eliminate the treadmill's low vibration mode
-    (approximately 5\si{\hertz}). But these blocks seem to have corrupted the
-    force plate measurements by imposing frictional stresses on the system. The
-    force plate measurements should not be used from these trials, but the
-    marker data is fine.
-  \item We did not record unloaded compensation trials for trials 9-15.
+  \item During trials 9--15, we used wooden blocks to fix the treadmill to the
+    concrete floor to eliminate the treadmill's low vibration mode. But these
+    blocks seem to have corrupted the force plate measurements by imposing
+    frictional stresses on the system. The force plate measurements should not
+    be used from these trials.
+  \item We did not record unloaded compensation trials for trials 9--15.
     Regardless, they would likely be useless due to the corruption from the
-    wooden blocks.
-  \item The ankle joint torques computed from subject 9's data in trials 25-27
-    are abnormal and should be used with caution or not at all. We were not
-    able to locate the source of the error, but it is likely related to the
-    force calibration.
-  \item Trials 37, 38, and 39 do not exist. The numbers were accidentally
-    skipped.
+    wooden blocks and are not needed because the treadmill base is fixed.
 \end{itemize}
@@
 loads from human subjects during both normal walking and when recovering from
-longitudinal perturbations. The raw data is provided for reuse with complete
-meta data. In addition to the data, we provide software that can process the
-data for both cleaning purposes and to produce typical sagittal plane gait
-variables of interest. Among other uses, we believe the dataset is ideally
-suited for control identification purposes. Many researchers are working on
-mathematical models for control in gait and this dataset provides both a way to
-validate these models and a source for generating them.
+perturbations. The raw data is provided for reuse with complete meta data. In
+addition to the data, we provide software that can process the data for both
+cleaning purposes and to produce typical sagittal plane gait variables of
+interest. Among other uses, we believe the dataset is ideally suited for
+control identification purposes. Many researchers are working on mathematical
+models for control in gait and this dataset provides both a way to validate
+these models and a source for generating them.
 

### 2015-03-17 src/subject_table.py

@@ -84,7 +84,6 @@ units = ['', '', ' [yr]', ' [m]', ' [kg]', '', '', '']
 new_cols = [s.capitalize() + u for s, u in zip(cols, units)]
 unique_subjects.rename(columns=dict(zip(cols, new_cols)), inplace=True)
 
-
 formatters = {'Height [m]': lambda x: 'NA' if np.isnan(x)
               else '{:0.2f}'.format(x),
               'Mass [kg]': lambda x: '{:0.0f}'.format(x) if isinstance(x, float) else x}
@@ -95,14 +94,14 @@ if not os.path.exists(table_dir):
 
 table_path = os.path.join(table_dir, 'subjects.tex')
 tex = unique_subjects.sort('1.6 m/s', na_position='first').to_latex(
-    na_rep='NA', index=False, columns=new_cols, formatters=formatters)
+    na_rep='NA', index=False, index_names=False, columns=new_cols,
+    formatters=formatters)
 
 tex = tex.replace('dollar', '$')
 tex = tex.replace('plusminus', '\pm')
 tex = tex.replace('rlrrllll', 'rlrrrrrr')
 lines = tex.split('\n')
-lines.insert(5, r'\midrule[0.25pt]')
-lines.insert(9, r'\midrule[0.25pt]')
+lines.insert(7, r'\midrule[0.25pt]')
 tex = '\n'.join(lines)
 with open(table_path, 'w') as f:
     f.write(tex)

### 2015-03-17 paper.tex
6 -> 3

@@
 
-Trials 6--8 use an early experimental protocol which divided the walking period
+Trials 3--8 use an early experimental protocol which divided the walking period
 into three sections: no perturbation, longitudinal perturbation, and a

### 2015-03-17 src/lateral_perturbation_plot.m
The lateral perturb figure now runs from the Makefile.

@@ -1,46 +1,70 @@
+% This script generates a plot of the lateral deviation of the treadmill
+% base durig trial number 6.
+
 clear
 clc
 
-PATHS = get_config.m;
+%-------------------------------------------------------------------------
+% File Paths
+%-------------------------------------------------------------------------
+
+PATHS = get_config();
+
+per_sig_dir = [PATHS.raw_data_dir filesep 'perturbation-signals'];
+lat_per_file = [per_sig_dir filesep 'lateral-perturbation.txt'];
+mocap_file = [PATHS.raw_data_dir filesep 'T006' filesep 'mocap-006.txt'];
 
 %-------------------------------------------------------------------------
-%Loading, Parsing, and Truncating
+% Loading, Parsing, and Truncating
 %-------------------------------------------------------------------------
-    perturb_sig_dir = [PATHS.raw_data_dir filesep 'perturbation-signals'];
-    lat_per_file = [perturb_sig_dir filesep 'lateral-perturbation.txt'];
-    perturbation_input = importdata(lat_per_file);
-    mocap_file = [PATHS.raw_data filesep 'T006' filesep 'mocap-006.txt'];
-    perturbation_output = importdata(mocap_file);
-    frame=1:1:length(perturbation_input);
-    time_input = (frame-1)*0.0033;
-    signal_input = perturbation_input(:,2);
-    time_output = perturbation_output.data(33090:57336,1);
-    time_output = time_output - time_output(1,1);
-    signal_output = perturbation_output.data(33090:57336,3);
-    signal_output = signal_output - signal_output(1,1);
-    signal_output = interp1(time_output,signal_output,time_input);
+
+perturbation_input = importdata(lat_per_file);
+frame = 1:1:length(perturbation_input);
+time_input = (frame - 1) * 0.0033;
+signal_input = perturbation_input(:, 2);
+
+perturbation_output = importdata(mocap_file);
+time_output = perturbation_output.data(33090:57336, 1);
+time_output = time_output - time_output(1, 1);
+signal_output = perturbation_output.data(33090:57336, 3);
+signal_output = signal_output - signal_output(1, 1);
+signal_output = interp1(time_output,signal_output,time_input);
+
[…]

### 2015-03-17 src/lateral_perturbation_plot.m
Set plot settings to print correct size.

@@ -34,23 +34,20 @@ signal_output = interp1(time_output,signal_output,time_input);
 % Plotting
 %-------------------------------------------------------------------------
 
-% Settings to Improve Graph Appearance
+h = figure();
 
-% TODO : I have no idea what this does and the figure doesn't come out to
-% an adequate size. I think it may just crop the image from the default
-% size which is incorrect.
+fig_width = 5.0;
+golden_ratio = (1 + sqrt(5)) / 2;
+fig_height = fig_width / golden_ratio;
 
-ti = get(gca, 'TightInset');
-set(gca, 'Position', [ti(1) ti(2) 1-ti(3)-ti(1) 1-ti(4)-ti(2)]);
-set(gca, 'units', 'centimeters')
-pos = get(gca, 'Position');
-ti = get(gca, 'TightInset');
-set(gcf, 'PaperUnits', 'centimeters');
-set(gcf, 'PaperSize',  [pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
-set(gcf, 'PaperPositionMode',  'manual');
-set(gcf, 'PaperPosition', [0 0 pos(3)+ti(1)+ti(3) pos(4)+ti(2)+ti(4)]);
-
-% Plot
+set(gcf, ...
+    'Color', [1, 1, 1], ...
+    'PaperOrientation', 'portrait', ...
+    'PaperUnits', 'inches', ...
+    'PaperPositionMode', 'manual', ...
+    'OuterPosition', [424, 305 - 50, 518, 465], ...
+    'PaperPosition', [0, 0, fig_width, fig_height], ...
+    'PaperSize', [fig_width, fig_height])
 
 hold on
 
@@ -64,7 +61,4 @@ ylabel('Lateral Position (m)')
 title('Lateral Perturbation Signal')
 legend('Commanded', 'Measured')
 
-% Saving
-
-h = figure(1);
 saveas(h, [PATHS.figures_dir filesep 'lateral_perturbation.pdf']);

### 2015-03-17 src/lateral_perturbation_plot.m
Removed the commanded signal from the graph.

@@ -10,25 +10,22 @@ clc
 
 PATHS = get_config();
 
-per_sig_dir = [PATHS.raw_data_dir filesep 'perturbation-signals'];
-lat_per_file = [per_sig_dir filesep 'lateral-perturbation.txt'];
 mocap_file = [PATHS.raw_data_dir filesep 'T006' filesep 'mocap-006.txt'];
 
 %-------------------------------------------------------------------------
 % Loading, Parsing, and Truncating
 %-------------------------------------------------------------------------
 
-perturbation_input = importdata(lat_per_file);
-frame = 1:1:length(perturbation_input);
-time_input = (frame - 1) * 0.0033;
-signal_input = perturbation_input(:, 2);
+start = 33090;
+stop = 57336;
 
 perturbation_output = importdata(mocap_file);
-time_output = perturbation_output.data(33090:57336, 1);
+
+time_output = perturbation_output.data(start:stop, 1);
 time_output = time_output - time_output(1, 1);
-signal_output = perturbation_output.data(33090:57336, 3);
+
+signal_output = perturbation_output.data(start:stop, 3);
 signal_output = signal_output - signal_output(1, 1);
-signal_output = interp1(time_output,signal_output,time_input);
 
 %-------------------------------------------------------------------------
 % Plotting
@@ -51,14 +48,12 @@ set(gcf, ...
 
 hold on
 
-plot(time_input, signal_input, 'Color', [0.815 0.3294 0.3020])
-plot(time_input, signal_output, 'Color', [0.2157 0.4706 0.7490])
+plot(time_output, signal_output, 'Color', [0.2157 0.4706 0.7490])
 
 xlim([100 200])
 
 xlabel('Time (s)')
 ylabel('Lateral Position (m)')
 title('Lateral Perturbation Signal')
-legend('Commanded', 'Measured')
 
 saveas(h, [PATHS.figures_dir filesep 'lateral_perturbation.pdf']);

### 2015-03-17 src/lateral_perturbation_plot.m
Updated the paper to reflect the absence of the commanded lateral perturbation.

@@ -53,7 +53,7 @@ plot(time_output, signal_output, 'Color', [0.2157 0.4706 0.7490])
 xlim([100 200])
 
 xlabel('Time (s)')
-ylabel('Lateral Position (m)')
+ylabel('Lateral Deviation (m)')
 title('Lateral Perturbation Signal')
 
 saveas(h, [PATHS.figures_dir filesep 'lateral_perturbation.pdf']);

### 2015-03-17 paper.tex

@@
 prescribed for a duration of four minutes in the pilot trials 3--8.
-Figure~\ref{fig:lateral_perturbation} shows the effect of the lateral control
-of the treadmill by plotting the commanded and measured treadmill positions.
-These signals were generated in a similar manner using MATLAB and Simulink in
-which a Gaussian white noise block that twice integrated to obtain the lateral
-position. The signal was then high-pass filtered with a second-order
-Butterworth filter to eliminate drift and then saturated so that the signal
-remained within the 5~\si{\centi\meter} lateral range of the physical hardware.
-The same perturbation signal was used for each of the three trials.
+Figure~\ref{fig:lateral_perturbation} shows an example of the measured lateral
+deviation of the treadmill base. These signals were generated in a similar
+manner using MATLAB and Simulink in which a Gaussian white noise block was
+twice integrated to obtain the lateral deviation. The signal was then high-pass
+filtered with a second-order Butterworth filter to eliminate drift and then
+saturated so that the signal remained within the 5~\si{\centi\meter} lateral
+range of the physical hardware. The same perturbation signal was used for each
+of the three trials.
 %
@@
   \includegraphics{figures/lateral_perturbation.pdf}
-  \cprotect\caption{Commanded treadmill lateral deviation (red) and the
-    recorded deviation (blue) for the trials 6, 7, and 8. Generated by
-    \verb|src/lateral_perturbation_plot.m|.}
+  \cprotect\caption{The measured lateral deviation of the treadmill base from
+    trial 6. Generated by \verb|src/lateral_perturbation_plot.m|.}
   \label{fig:lateral_perturbation}

### 2015-03-17 src/lateral_perturbation_plot.m
Removed clear;clc.

@@ -1,8 +1,5 @@
 % This script generates a plot of the lateral deviation of the treadmill
-% base durig trial number 6.
-
-clear
-clc
+% base during trial number 6.
 
 %-------------------------------------------------------------------------
 % File Paths

### 2015-03-17 src/input_output_plot.m
Introduced PATHS and specified plot size.

@@ -1,67 +1,107 @@
-clear
-clc
+%-------------------------------------------------------------------------
+% File Paths
+%-------------------------------------------------------------------------
+
+PATHS = get_config();
+
+raw_dir = PATHS.raw_data_dir;
+
+per_sig_dir = [raw_dir filesep 'perturbation-signals'];
+commanded_file = [per_sig_dir filesep 'longitudinal-perturbation.txt'];
+
+measured_076 = [raw_dir filesep 'T076' filesep 'record-076.txt'];
+measured_077 = [raw_dir filesep 'T077' filesep 'record-077.txt'];
+measured_078 = [raw_dir filesep 'T078' filesep 'record-078.txt'];
 
 %-------------------------------------------------------------------------
-%Loading Files
+% Loading Files
 %-------------------------------------------------------------------------
-    input_signals = importdata('Longitudinal_Perturbation_2.4.2014.txt');
-    output_slow = importdata('record-076.txt');
-    output_normal = importdata('record-077.txt');
-    output_fast = importdata('record-078.txt');
+
+input_signals = importdata(commanded_file);
+output_slow = parse_record(measured_076);
+output_normal = parse_record(measured_077);
+output_fast = parse_record(measured_078);
+
 %-------------------------------------------------------------------------
-%Declaring Variables
+% Declaring Variables
 %-------------------------------------------------------------------------
-    %Time Variables (necessary because Record Module has varying timestamp)
-        time_input = input_signals(:,1);
-        output_slow_time = output_slow.data(:,1); 
-        output_slow_time = output_slow_time-output_slow_time(1,1);
-        output_normal_time = output_normal.data(:,1);
-        output_normal_time = output_normal_time-output_normal_time(1,1);
-        output_fast_time = output_fast.data(:,1);
-        output_fast_time  =output_fast_time-output_fast_time(1,1);
-    %Speed
-        inputs=[input_signals(:,2) input_signals(:,3) input_signals(:,4)];
-        output_slow = output_slow.data(:,2); 
[…]

### 2015-03-17 src/input_output_plot.m
box on for matlab plots

@@ -67,6 +67,7 @@ set(gcf, ...
 red = [0.815 0.3294 0.3020];
 blue = [0.2157 0.4706 0.7490];
 
+
 % Slow
 subplot(3, 1, 1)
 hold on
@@ -75,8 +76,10 @@ plot(time_input, inputs(:, 1), 'Color', red)
 plot(output_slow_time, output_slow, 'Color', blue)
 xlim([305 310]);
 
+box on
+
 leg = legend('Commanded', 'Measured');
-set(leg,'FontSize', 8);
+set(leg,'FontSize', 6);
 
 ylabel('Speed (m/s)')
 title('Slow Walking (0.8 m/s)')
@@ -89,6 +92,8 @@ plot(time_input, inputs(:, 2), 'Color', red)
 plot(output_normal_time, output_normal, 'Color', blue)
 xlim([305 310]);
 
+box on
+
 ylabel('Speed (m/s)')
 title('Normal Walking (1.2 m/s)')
 
@@ -99,6 +104,8 @@ hold on
 plot(time_input, inputs(:, 3), 'Color', red)
 plot(output_fast_time, output_fast, 'Color', blue)
 xlim([305 310])
+
+box on
 
 xlabel('Time (s)')
 ylabel('Speed (m/s)')

### 2015-03-17 src/lateral_perturbation_plot.m

@@ -47,6 +47,8 @@ hold on
 
 plot(time_output, signal_output, 'Color', [0.2157 0.4706 0.7490])
 
+box on
+
 xlim([100 200])
 
 xlabel('Time (s)')

### 2015-04-28 paper.tex
Added editor and reviewers to the acknowledgements.
Updates the draft to match the minor edits found in the proof.
Fixes #173, #172, #171, #170.

@@
 database contained kinematic data of 409 gait cycles of children from 1 to 7
-years old but the data does not seem to be publicly available. This is,
-unfortunately, typical. \cite{Tirosh2010} recognized the need for a
+years old but the data does not seem to be publicly available. This is
+unfortunately typical. \cite{Tirosh2010} recognized the need for a
 comprehensive data base for clinical gait data and created the Gaitabase. This
@@
 from 25 subjects with lower extremity marker data from multiple gait cycles and
-force plate measurements from a single gait cycle. The CMU Graphics Lab
-Motion Capture Database~\citep{CMU2015} is also a good example and contains
-full body marker kinematics for a fair number of trials with small number of
-gait cycles during both walking and running.
+force plate measurements from a single gait cycle. The CMU Graphics Lab Motion
+Capture Database~\citep{CMU2015} is also a good example and contains full body
+marker kinematics for a fair number of trials with small number of gait cycles
+during both walking and running.
 
@@
 from eleven subjects walking on an instrumented treadmill and \cite{Wang2014}
-who includes a larger set of data from ten subjects walking for five minutes
+who include a larger set of data from ten subjects walking for five minutes
 each at three different speeds but only a small set of lower extremity markers
-are present. The second is notable because it publishes the data in Dryad, a
+are present. The second is notable because they publish the data in Dryad, a
 modern citable data repository. It is also worth noting purely visual data
@@
 controllers that have been designed in other manners, such as those constructed
-from first principles \citep{Geyer2010}.
+from first principles, e.g. \citep{Geyer2010}.
 
 With all of this in mind, we collected over seven and a half hours of gait data
-from fifteen able bodied subjects which amounts to over 25,000 gait cycles. The
-subjects walked at three different speeds on an instrumented treadmill while we
-collected full body marker locations and ground reaction loads from a pair of
-force plates. The final protocol for the majority of the trials included two
-minutes of normal walking and eight minutes of walking under the influence of
-pseudo-random belt speed fluctuations. The data has been organized complete
-with rich meta data and made available in the most unrestrictive form for other
-research uses following modern best practices in data sharing
-\citep{White2013}.
+from fifteen able bodied subjects which amounts to over 25,000 gait cycles
+\citep{Moore2014}. The subjects walked at three different speeds on an
+instrumented treadmill while we collected full body marker locations and ground
+reaction loads from a pair of force plates. The final protocol for the majority
+of the trials included two minutes of normal walking and eight minutes of
+walking under the influence of pseudo-random belt speed fluctuations. The data
+has been organized complete with rich meta data and made available in the most
+unrestrictive form for other research uses following modern best practices in
+data sharing \citep{White2013}.
 
@@
 walking direction. The camera's coordinate system is aligned to an origin point
-on treadmill's surface during camera calibration. The same point is used as the
-origin of the ground reaction force measuring system.
+on the treadmill's surface during camera calibration. The same point is used as
+the origin of the ground reaction force measuring system.
 Figure~\ref{fig:treadmill} shows the layout of the equipment.
@@
 vibration mode as low as 5\si{\hertz} that was detectable in the force
-measurements, likely due to the flexible undercarriage and pitch motion
+measurements; likely due to the flexible undercarriage and pitch motion
 mechanism. Trials 6--8 are affected by this vibration mode. During trials 9--15
 the treadmill was stabilized with wooden blocks. During the remaining trials
-($>$~15) the treadmill was stabilized with metal supports; both ones we
+($>$~15) the treadmill was stabilized with metal supports; both with ones we
 fabricated in-house and ones supplied by the vendor. These supports aimed to
@@
 \begin{enumerate}
-  \item Subject stepped onto the treadmill and markers were identified with
+  \item The subject stepped onto the treadmill and markers were identified with
     Cortex.
@@
 named in the following fashion \verb+T001/+ where \verb+T+ stands for ``trial''
-and the following three digits are provide a unique trial identification
-number.
+and the following three digits provide a unique trial identification number.
 
@@
     receives a frame from Cortex. These are recorded approximately at
-    100~\si{\hertz} sampling rate and given in seconds.
+    100~\si{\hertz} sampling rate and given in seconds. Due to data buffering,
+    it is preferred to derive sample times from the FrameNumber column rather
+    than the TimeStamp.
   \item[FrameNumber] Monotonically increasing positive integers that correspond
@@
   \item[nominal-speed-units] A string providing the units of the nominal speed.
-  \item[notes] Any notes about the trial.
+  \item[notes] A string with any notes about the trial.
   \item[pitch] A boolean that indicates if the treadmill pitch degree of
@@
        cortex-version: 3.1.1.1290
+       marker-set: full
        marker-map:
@@
 The \verb|GaitData| class is then used to compute gait events (toe off and heel
-strike times), basic 2D kinematics and inverse dynamics, and to store the data
+strike times), basic 2D inverse kinematics and dynamics, and to store the data
 into a Pandas \verb|Panel| with each gait cycle on the item axis at a specified
@@
 
-A similar work flow was used to produce Figure~\ref{fig:angle-torque-comparison}
-which compares the mean and standard deviation of sagittal plane joint angles
-and torques from the perturbed gait cycles and the unperturbed gait cycles
-computed from trial 20. This gives an idea of the more highly variable dynamics
-required to walk while being longitudinally perturbed.
+A similar work flow was used to produce
+Figure~\ref{fig:angle-torque-comparison} which compares the mean and standard
+deviation of sagittal plane joint angles and torques from the perturbed gait
+cycles to the unperturbed gait cycles computed from trial 20. This gives an
+idea of the more highly variable dynamics required to walk while being
+longitudinally perturbed.
 %
@@
 larger variation in frequency and length and even larger variation in stride
-width. It is also interesting to note that the coupled nature of system's
+width. It is also interesting to note that the coupled nature of the system's
 degrees of freedom can be exploited to increase the stride width with only
@@
 %
-The data set \citep{Moore2014} is available via the Zenodo data repository.
-Two approximately 1.2GB gzipped tar balls contain the data and a \verb|README|
-file with a short description of the contents. The data is released under the
-Creative Commons CC0 license (http://creativecommons.org/about/cc0) following
-best practices for sharing scientific data.
+The data set \citep{Moore2014} is available via two Zenodo data repositories.
+The first holds two approximately 1.2GB gzipped tar balls with contain the main
+data and a \verb|README| file with a short description of the contents. The
+second \citep{Hnat2015} contains the Simulink generated data for the treadmill
+control inputs. The data is released under the Creative Commons CC0 license
+(http://creativecommons.org/about/cc0) following best practices for sharing
+scientific data.
 
@@
 the preprint and being our guinea pigs on the software/data instructions. Dan
-Simon also gave valuable feedback on the preprint.
+Simon also gave valuable feedback on the preprint. Furthermore, we thank the
+academic editor, Arti Ahluwalia, and three reviewers, Morgan Sangeux, Paul Lee,
+and Manoj Srinivasan, for their valuable feedback which helped improve the
+quality of the paper and data.
 

### 2015-06-08 paper.tex
Fixed typo R-Link > R-Mill.

@@
 
-Early on, we discovered that the factory setup of the R-Link treadmill had a
+Early on, we discovered that the factory setup of the R-Mill treadmill had a
 vibration mode as low as 5\si{\hertz} that was detectable in the force
